---
title: 'MEC analysis'
subtitle: "Association of interactions between SNPs and PFAS with HCC"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup snp and pfas, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("1_Project_Setup/!libraries.R")))
source(fs::path(here::here("1_Project_Setup/!directories.R")))
source(fs::path(here::here("1_Project_Setup/!functions.R")))
source(fs::path(here::here("1_Project_Setup/!load_data.R")))
```

# Checking SNPs
**x3_12288284, x14_94378610, x17_66214462, x19_19268740 were excluded from analysis**
```{r correlation of SNPs}
pfc_chem_summaries <-  data_hcc %>%
  dplyr::select(x1_220796686:x22_43928850) %>%
  pivot_longer(cols = x1_220796686:x22_43928850) %>% 
  group_by(name) %>% 
  summarise(pct_one_val   = mean(value == 0 | value == 2, na.rm = TRUE),
            variance = var(value, na.rm = TRUE),
            mean    = mean(value, na.rm = TRUE))

snp_name_df <-  tidylog::left_join(pfc_chem_summaries, 
                                   snp_info, by = c("name" = "snp_name")) |>
  tidylog::filter(pct_one_val < 0.75)

snp_name_scld <- c(paste0(snp_name_df$name, "_scld"))

snp_name_new <- snp_name_df$name
```

## RAF calculation

```{r}
crc_b <-data_hcc %>% filter(eth=='African American') %>% pull(sid)
crc_h <-data_hcc %>% filter(eth=='Native Hawaiian') %>% pull(sid)
crc_j <-data_hcc %>% filter(eth=='Japanese American') %>% pull(sid)
crc_l <-data_hcc %>% filter(eth=='Latino') %>% pull(sid)
crc_w <-data_hcc %>% filter(eth=='White') %>% pull(sid)

ds_correct <- data_hcc %>% select(sid, x1_220796686:x22_43928850) %>%
  column_to_rownames("sid") %>% t() %>% as.data.frame()

#get single dataset of the risk allele frequency across all race/eth
raf <- data.frame(ds_correct,         # can carry over any cols from ds_correct as needed
                  raf = rowSums(ds_correct, na.rm = TRUE)/(2*dim(data_hcc)[1]),
                  raf_b=rowSums(ds_correct %>% select(any_of(crc_b)), na.rm = TRUE)/(2*length(crc_b)),
                  raf_h=rowSums(ds_correct %>% select(any_of(crc_h)), na.rm = TRUE)/(2*length(crc_h)),
                  raf_j=rowSums(ds_correct %>% select(any_of(crc_j)), na.rm = TRUE)/(2*length(crc_j)),
                  raf_l=rowSums(ds_correct %>% select(any_of(crc_l)), na.rm = TRUE)/(2*length(crc_l)),
                  raf_w=rowSums(ds_correct %>% select(any_of(crc_w)), na.rm = TRUE)/(2*length(crc_w)))%>%
  rownames_to_column("snp_name") %>%
  tidylog::left_join(snp_info %>% select(rsid, snp_name), by = "snp_name")

write_csv(raf, fs::path(dir_result,"Supplemental Tables", "raf_hcc.csv"))

```

## summary of SNPs
```{r summary of SNPs}
# summarytools::view(
#   summarytools::dfSummary(data_hcc %>% 
#                             dplyr::select(all_of(snp_name_df$name))%>%
#                             janitor::clean_names(),
#                           style = 'grid',
#                           max.distinct.values = 10, 
#                           plain.ascii =   FALSE, 
#                           valid.col = FALSE, 
#                           headings = FALSE), method = "render")

```

## summary of SNPs after normalized 
* Finally, We used raw SNP instead of normalized SNPs in the following analysis
```{r summary of SNPs after normalized}
data_analysis_hcc_snp <- data_hcc %>%
  # rename_at(.vars = vars(colnames(data_hcc %>% select(x1_220796686:x22_43928850))),
  #           .funs = list(~str_sub(., start = 1, end = 8))) %>%
  mutate_at(.vars = vars(snp_name_df$name),
            .funs = list(scld = ~scale(.))) 

data_analysis_nafld_hcc_snp <- data_nafld_hcc %>%
  # rename_at(.vars = vars(colnames(data_hcc %>% select(x1_220796686:x22_43928850))),
  #           .funs = list(~str_sub(., start = 1, end = 8))) %>%
  mutate_at(.vars = vars(snp_name_df$name),
            .funs = list(scld = ~scale(.))) 

data_analysis_other_hcc_snp <- data_other_hcc %>%
  # rename_at(.vars = vars(colnames(data_hcc %>% select(x1_220796686:x22_43928850))),
  #           .funs = list(~str_sub(., start = 1, end = 8))) %>%
  mutate_at(.vars = vars(snp_name_df$name),
            .funs = list(scld = ~scale(.))) 

data_analysis_hcc_japan_snp <- data_hcc %>%
  filter(eth == "Japanese American") %>%
  # rename_at(.vars = vars(colnames(data_hcc %>% select(x1_220796686:x22_43928850))),
  #           .funs = list(~str_sub(., start = 1, end = 8))) %>%
  mutate_at(.vars = vars(snp_name_df$name),
            .funs = list(scld = ~scale(.))) 


data_analysis_hcc_latino_snp <- data_hcc %>%
  filter(eth == "Latino") %>%
  # rename_at(.vars = vars(colnames(data_hcc %>% select(x1_220796686:x22_43928850))),
  #           .funs = list(~str_sub(., start = 1, end = 8))) %>%
  mutate_at(.vars = vars(snp_name_df$name),
            .funs = list(scld = ~scale(.))) 


# summarytools::view(
#   summarytools::dfSummary(data_analysis_hcc_snp %>%
#                             dplyr::select(contains("scld"),
#                                           -contains("pfas")),
#                           style = 'grid',
#                           max.distinct.values = 10,
#                           plain.ascii =   FALSE,
#                           valid.col = FALSE,
#                           headings = FALSE),
#   method = "render")
```

# Distribution of SNP
```{r}
temp_data <- data_hcc %>% rename(rs5117 = x19_44915533)%>%
  mutate(rs5117 = round(rs5117, 0)) %>%
  drop_na(rs5117)%>%
  mutate(rs5117 = case_when(rs5117 == 2 ~ "2 rs5117-T alleles", 
                            rs5117 == 1 ~ "1 rs5117-T allele", 
                            rs5117 == 0 ~ "0 rs5117-T alleles"))
 
temp_counts <- temp_data %>%
  count(rs5117) 
temp_counts <- temp_counts %>%
  mutate(perc = n / sum(n) * 100)%>%
  bind_rows(data.frame(rs5117 = "Overall",
                       n = sum(temp_counts$n),
                       perc = 100)) %>%
  mutate(rs5117 = factor(rs5117, levels = c("2 rs5117-T alleles",
                                            "1 rs5117-T allele",
                                            "0 rs5117-T alleles",
                                            "Overall"
                                            ))) %>%
  mutate(group = if_else(str_detect(rs5117, "Overall"), "overall", "genetic"),
         group = factor(group, levels = c("overall", "genetic")))

# Plot
(p_dist<- ggplot(temp_counts, aes(y = rs5117, x = n)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(perc, 0), "%")), 
            hjust = -0.1) +  
  labs(x = "Count") +
  facet_grid(group ~. , scales = "free_y", space = "free_y")+
  theme(panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_blank(),
          # strip.text.x = element_blank(),
          text = element_text(size = 12),
          # axis.text.x =element_text(angle = 60, hjust = 1),
          axis.line.x = element_line(color = "black"),
          axis.title.y=element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
  xlim(0, max(temp_counts$n) * 1.2) ) 
```

# Analysis 1: PRS risk score

**Outcome: HCC vs control**

**Exposure: normalzied log2 PFAS**

**Covars: smoking, diabetes, q1_edih, v1:v6**

**Analysis: conditional logistic regression**

**Data Info: 240 cases, 240 controls**

**All races(480), White(48), African American(52), Japanese American(184), Latino(162), Native Hawaiian(34)**

**ANOVA: Model1: status ~ pfas + prs_wt_std + smoking + diabetes + q1_edih + v1...+v6**
**Model2:status ~ pfas + prs_wt_std + pfas:prs_wt_std + diabetes + q1_edih + v1...+v6**
```{r}
# Anova: if interaction is significant
interaction_models <- function(pfas) {
  # Formula
  base_covars <- paste("v1 + v2 + v3 + v4 + v5 + v6 +", 
                       paste(covars[1:3], collapse = " + "), 
                       "+ strata(setnum)")
  
  frm_no_int <- as.formula(paste("status ~ prs_wt_std+", paste(pfas, base_covars, sep = " + ")))
  frm_with_int <- as.formula(paste("status ~", paste0(pfas, "*prs_wt_std"), base_covars, sep = " + "))
  
  # Modeling
  model_no_int   <- clogit(frm_no_int, data = data_hcc)
  model_with_int <- clogit(frm_with_int, data = data_hcc)
  
  # ANOVA
  anova_res <- anova(model_no_int, model_with_int, test = "Chisq")%>% 
    tidy(conf.int = TRUE) %>%
    mutate(pfas_name = pfas)%>%
    select(pfas_name, everything())
  
  return(anova_res)
}



(outputs <- map(c(pfas_name, "pfas_t_PFBS_detected"),interaction_models)%>%
    bind_rows()%>%
    mutate(model = ifelse(grepl("\\*",term), "With Interaction", "Without Interaction"),
           pfas_name = rename_pfas(pfas_name)) %>%
    select(pfas_name, model, everything(),-term)
    )

# writexl::write_xlsx(outputs, fs::path(dir_result, "anova_prs_pfas_inter_1208.xlsx"))
```

# Table. Interaction of PRS and PFAS
```{r}
covariate_input <- paste(covars[1:3], collapse = " + ")

# create dataframe with names of dependent vars (PFAS) 
pfas_names <- tibble(exposure = c(pfas_name, "pfas_t_PFBS_detected"))

## create dataframe with names of SNP vars ----
prs_name <- tibble(prs = "prs_wt_std")

pfas_prs <- expand_grid(pfas_names, prs_name) %>%
  mutate(pfas_prs = paste(exposure,"*",prs)) 

# Get Model covariates 
covar_names <- tibble(covariates =  covariate_input)

# 2. Make all clogit formulas ----
ind_vars <- expand_grid(pfas_prs,covar_names) |>  
  mutate(formula = paste("status ~ v1 + v2 + v3 + v4 + v5 + v6",
                         "+", pfas_prs,
                         "+", covariates, 
                         "+strata(setnum)") |>
           str_replace("\\+  \\+", " + "))

ind_vars$clogit_out <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_hcc,
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

all_results_df_hcc <- ind_vars %>% 
  unnest(clogit_out) %>% 
  dplyr::select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high))

res_inter <- all_results_df_hcc %>%
  # filter(grepl(":", term))%>%
  group_by(exposure) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         sig = case_when(p_adjusted < 0.05 ~ "FDR. Sig.", 
                         p.value < 0.05 ~ "Sig.",
                         TRUE ~ "Not Sig."))

writexl::write_xlsx(res_inter, fs::path(dir_result, "pfas_prs_interaction.xlsx"))

```

# Table. Interaction of PRS (dichotomous weighted PRS) and PFAS
```{r}
model_interaction <- function(pfas){
  # Formula
  base_covars <- paste(paste(covars[-4], collapse = " + "), 
                       "+ strata(setnum)")
  prs = "prs_wt_median"
  frm_with_int <- as.formula(paste("status ~", paste0(pfas, "*", prs), base_covars, sep = " + "))
  # Model
  model_with_int <- clogit(frm_with_int, 
                           data = data_hcc, 
                           method = "efron", 
                           robust = TRUE)
  
  int <- model_with_int %>% tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")
  
  # Interaction effect at high/low PRS
  res <- emtrends(model_with_int,
                  ~ prs_wt_median,
                  var = pfas,
                  at = list(prs_wt_median = c("1", "2"),
                            smokestatus_imputed = "Never",
                            diabetes = "No",
                            q1_edih = 0,
                            prs_wt = 0,
                            v1 = 0,
                            v2 = 0,
                            v3 = 0,
                            v4 = 0,
                            v5 = 0,
                            v6 = 0),
                  transform = NULL) %>%
    summary(infer = TRUE) %>%
    as.data.frame() %>%
    select(`prs_wt_median`, ends_with(".trend"), asymp.LCL, asymp.UCL, p.value) %>%
    rename(estimate = ends_with(".trend"),
           conf_low = asymp.LCL,
           conf_high = asymp.UCL,
           p_value = p.value) %>%
    mutate(type = case_when(`prs_wt_median` == 1 ~ "Low Genetic Risk",
                            `prs_wt_median` == 2 ~ "High Gnetic Risk"
                            ))
  
  # Main effect without interaction
  res_main <- epiomics::owas_clogit(data_hcc , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars,
                                    omics = pfas,
                                    conf_int = TRUE) %>%
    mutate(type = "Overall (no interaction with PRS)")%>%
    select(-adjusted_pval)
  # Combine results
  res_combined <- bind_rows(res_main, res) %>%
    mutate(odds_ratio = exp(estimate),
           exp_ci_low = exp(conf_low),
           exp_ci_high = exp(conf_high),
           group = if_else(str_detect(type, "Overall"), "overall", "with genetic"),
           pfas = pfas)
  
  res_combined
  return(list(res_combined = res_combined,
  int = int))
}

res <- map(c(pfas_name[-3], "pfas_t_PFBS_detected"), model_interaction)%>%
  map("res_combined") %>%
    bind_rows()%>%
    mutate(pfas = str_remove(pfas,"pfas_t_"))%>%
  group_by(type) %>%
  mutate(q_value = p.adjust(p_value, "fdr"))

int <- map(c(pfas_name[-3], "pfas_t_PFBS_detected"), model_interaction)%>%
  map("int") %>%
  bind_rows()%>%
  filter(str_detect(term, ":")) %>%
  mutate(pfas = str_remove(term,"pfas_t_") %>%
           str_remove(":prs_wt_median2"))%>%
  rename(p_value = p.value) %>%
  mutate(q_value = p.adjust(p_value, "fdr"))%>%
  mutate(type = "PFAS*PRS intercation",
         odds_ratio = exp(estimate),
         exp_ci_low = exp(conf.low),
         exp_ci_high = exp(conf.high))%>%
  select(pfas, p_value, q_value, type, odds_ratio, exp_ci_low, exp_ci_high)

res <- res %>% bind_rows(int)

writexl::write_xlsx(res, fs::path(dir_result, "pfas_prs_wt_median_interaction_result_0724.xlsx"))



res %>% 
       select(pfas, type, odds_ratio, exp_ci_low, exp_ci_high, p_value)%>%
       filter(pfas == "PFNA")

```

# Plot PFAS and PRS interaction results

```{r,fig.width=15, fig.height=4}
cuts <- quantile(data_hcc %>% filter(status == 0) %>% select(prs_wt_std), probs = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9), na.rm = TRUE)

res <- map(c(pfas_name, "pfas_t_PFBS_detected"), model_interaction)%>%
    bind_rows()%>%
    mutate(pfas = str_remove(pfas,"pfas_t_"))

# writexl::write_xlsx(res, fs::path(dir_result, "pfas_prs_interaction_result_1208.xlsx"))

res %>% select(pfas, type, odds_ratio, exp_ci_low, exp_ci_high, p_value)

(p <- res  %>%
    ggplot(aes(x = factor(type, 
                          levels = c("High Genetic Risk", 
                                     "Low Genetic Risk", 
                                     "Overall (no interaction with PRS)")),
               y = exp(estimate))) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = exp(conf_low),
                      ymax = exp(conf_high)),
                  width = 0) +
    geom_hline(yintercept = 1, linetype = 2, color = "grey") +
    ylab("Odds Ratio (95% CI)") +
    facet_grid(group ~ pfas, scales = "free_y", space = "free_y")+
    theme(panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_blank(),
          # strip.text.x = element_blank(),
          text = element_text(size = 12),
          # axis.text.x =element_text(angle = 60, hjust = 1),
          axis.line.x = element_line(color = "black"),
          axis.title.y=element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
    coord_flip()
)

# ggsave(filename = fs::path(dir_figures,
#                            "Fig2. Interaction of PFAS with NAFLD PRS_All_PFAS.jpeg"),
#        width =14,
#        height = 3,
#        dpi = 600,
#        bg = "white")
```


## PFHpA: Plot PFAS and PRS interaction results
```{r}
(pfhpa <- res  %>%
    filter(pfas == "PFHPA") %>%
    ggplot(aes(x = factor(type, 
                          levels = c("High Genetic Risk", 
                                     "Low Genetic Risk", 
                                     "Overall (no interaction with PRS)")),
               y = exp(estimate))) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = exp(conf_low),
                      ymax = exp(conf_high)),
                  width = 0) +
    geom_hline(yintercept = 1, linetype = 2, color = "grey") +
    ylab("Odds Ratio (95% CI)") +
    facet_grid(group ~ pfas, scales = "free_y", space = "free_y")+
    theme(panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_blank(),
          # strip.text.x = element_blank(),
          text = element_text(size = 12),
          # axis.text.x =element_text(angle = 60, hjust = 1),
          axis.line.x = element_line(color = "black"),
          axis.title.y=element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
    coord_flip()
)

# ggsave(filename = fs::path(dir_figures,
#                            "Fig2. Interaction of PFHpA with NAFLD_0421.jpeg"),
#        width =14,
#        height = 3,
#        dpi = 600,
#        bg = "white")
```

# Analysis 2: All SNPS
**Outcome: HCC vs control**

**Exposure: normalzied log2 PFAS, raw SNP**

**Covars: smoking, diabetes, q1_edih, prs_wt_std, v1:v6**

**Analysis: conditional logistic regression**

**Data Info: 240 cases, 240 controls**

**All races(480), White(48), African American(52), Japanese American(184), Latino(162), Native Hawaiian(34)**

**Model: outcome ~ one pfas + one SNP + pfas:snp**
```{r interaction between PFAS and SNPs hcc}
# 1. Set Covariates ----
# Covariates for main exposure outcome associations
covariate_input <- paste(covars[1:4], collapse = " + ")

# create dataframe with names of dependent vars (PFAS) 
pfas_names <- tibble(exposure = c(pfas_name, "pfas_t_PFBS_detected"))

## create dataframe with names of SNP vars ----
# snp_name <- tibble(snp = snp_name_scld)
# based on Dave's suggestion use raw snp instead of normalized snp
snp_name <- tibble(snp = snp_name_new)

pfas_snp <- expand_grid(pfas_names, snp_name) %>%
  mutate(pfas_snp = paste(exposure,"*",snp)) 

# Get Model covariates 
covar_names <- tibble(covariates =  covariate_input)

# 2. Make all clogit formulas ----
ind_vars <- expand_grid(pfas_snp,covar_names) |>  
  mutate(formula = paste("status ~ v1 + v2 + v3 + v4 + v5 + v6",
                         "+", pfas_snp,
                         "+", covariates, 
                         "+strata(setnum)") |>
           str_replace("\\+  \\+", " + "))

ind_vars$clogit_out <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_analysis_hcc_snp,
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 


ind_vars$clogit_out1 <- map(
  ind_vars$formula,
  ~clogit(as.formula(.),
          data = data_analysis_hcc_japan_snp,
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>%
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit"))
# 
# 
ind_vars$clogit_out2 <- map(
  ind_vars$formula,
  ~clogit(as.formula(.),
          data = data_analysis_hcc_latino_snp,
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>%
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit"))

# Get final data -------------------
all_results_df_hcc <- ind_vars %>% 
  unnest(clogit_out) %>% 
  dplyr::select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high))

all_results_df_hcc_japan <- ind_vars %>%
  unnest(clogit_out1) %>%
  dplyr::select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig",
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high))

all_results_df_hcc_latino <- ind_vars %>%
  unnest(clogit_out2) %>%
  dplyr::select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig",
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high))

## Adding modify variable names, add type or category info for data-----

### All races
all_results_df_hcc1 <- all_results_df_hcc %>% 
  # bind_rows(intercept_data) %>% 
  filter(grepl("pfas_t_",term)|grepl("x", term)) %>%
  dplyr::select(exposure, snp, term, p.value, odds_ratio, contains("exp_")) %>%
  mutate(pfas_name = str_remove(exposure, "pfas_t_"))


all_results_df_hcc2 <- all_results_df_hcc1 |> 
  mutate(
    effect_type = case_when(str_detect(term, ":")   ~ "PFAS and SNP",
                            str_detect(term, "^x")   ~ "Genetic Risk",
                            TRUE ~ "PFAS")) |>
  filter(!is.na(effect_type)) 


all_results_df_hcc3 <- all_results_df_hcc2 %>% 
  group_by(pfas_name) %>%
  mutate(snp = str_replace_all(snp, "_scld", ""),
         p_adjusted = p.adjust(p.value, method = "fdr"),
         sig = case_when(p_adjusted < 0.05 ~ "FDR. Sig.", 
                         p.value < 0.05 ~ "Sig.",
                         TRUE ~ "Not Sig.")) %>%
  ungroup() %>%
  tidylog::left_join(snp_info%>%
                       dplyr::select(rsid, snp_name, genes), by = c("snp"="snp_name"))

## Japanese American
all_results_df_hcc_japan1 <- all_results_df_hcc_japan %>%
  # bind_rows(intercept_data) %>%
  filter(grepl("pfas_t_",term)|grepl("x", term)) %>%
  dplyr::select(exposure, snp,term, p.value, odds_ratio, contains("exp_")) %>%
  mutate(pfas_name = str_remove(exposure, "pfas_lg2_"))


all_results_df_hcc_japan2 <- all_results_df_hcc_japan1 |>
  mutate(
    effect_type = case_when(str_detect(term, ":")   ~ "PFAS and SNP",
                            str_detect(term, "^x")   ~ "Genetic Risk",
                            TRUE ~ "PFAS")) |>
  filter(!is.na(effect_type))

all_results_df_hcc_japan3 <- all_results_df_hcc_japan2 %>%
  filter(!grepl("x14", term)) |>
  group_by(pfas_name) %>%
  mutate(snp = str_replace_all(snp, "_scld", ""),
         p_adjusted = p.adjust(p.value, method = "fdr"),
         sig = case_when(p_adjusted < 0.05 ~ "FDR. Sig.",
                         p.value < 0.05 ~ "Sig.",
                         TRUE ~ "Not Sig.")) %>%
  tidylog::left_join(snp_info%>%
                       dplyr::select(rsid, snp_name, genes), by = c("snp"="snp_name")) %>%
  ungroup()
## Latino

all_results_df_hcc_latino1 <- all_results_df_hcc_latino %>%
  # bind_rows(intercept_data) %>%
  filter(grepl("pfas_t_",term)|grepl("x", term)) %>%
  dplyr::select(exposure, snp,term, p.value, odds_ratio, contains("exp_")) %>%
  mutate(pfas_name = str_remove(exposure, "pfas_lg2_"))


all_results_df_hcc_latino2 <- all_results_df_hcc_latino1 |>
  mutate(
    effect_type = case_when(str_detect(term, ":")   ~ "PFAS and SNP",
                            str_detect(term, "^x")   ~ "Genetic Risk",
                            TRUE ~ "PFAS")) |>
  filter(!is.na(effect_type))

all_results_df_hcc_latino3 <- all_results_df_hcc_latino2 %>%
  filter(!grepl("x14", term)) |>
  group_by(pfas_name) %>%
  mutate(snp = str_replace_all(snp, "_scld", ""),
         p_adjusted = p.adjust(p.value, method = "fdr"),
         sig = case_when(p_adjusted < 0.05 ~ "FDR. Sig.",
                         p.value < 0.05 ~ "Sig.",
                         TRUE ~ "Not Sig.")) %>%
  tidylog::left_join(snp_info%>%
                       dplyr::select(rsid, snp_name, genes), by = c("snp"="snp_name")) %>%
  ungroup()

all_result <- all_results_df_hcc3 %>%
  mutate(type = "All races") %>%
  bind_rows(all_results_df_hcc_japan3 %>%
              mutate (type = "Japanese American")) %>%
  bind_rows(all_results_df_hcc_latino3 %>%
              mutate (type = "Latino"))%>%
  mutate(name_plot = case_when(effect_type == "PFAS" ~ paste0(genes, "(", " PFAS Main Effect)"),
                                effect_type == "Genetic Risk" ~ paste0(genes, " Main Effect"),
                                effect_type == "PFAS and SNP" ~ paste0(genes, "*",
                                                                       "PFAS Interaction")))

all_result_new <- all_results_df_hcc3 %>%
  mutate(type = "All races") %>%
  bind_rows(all_results_df_hcc_japan3 %>%
              mutate (type = "Japanese American")) %>%
  bind_rows(all_results_df_hcc_latino3 %>%
              mutate (type = "Latino"))%>%
  mutate(name_plot  = case_when(effect_type == "PFAS" ~ paste0(snp, " (", "PFAS Main Effect)"),
                                effect_type == "Genetic Risk" ~ paste0(snp, " Main Effect"),
                                effect_type == "PFAS and SNP" ~ paste0(snp, "*",
                                                                       "PFAS Interaction")),
         gene = if_else(is.na(genes), snp, genes),
         rsid = if_else(is.na(rsid), snp, rsid) )

# Calculate adjusted p-values
intef <- all_result |> 
  tidylog::filter(effect_type == "PFAS and SNP") |>
  group_by(type) |> 
  mutate(p_adjust2 = p.adjust(p.value, method = "fdr")) |> 
  ungroup()

pfhpa_gene_interaction <- intef |> 
  tidylog::filter(effect_type == "PFAS and SNP") |>
  tidylog::filter(str_detect(exposure, "PFHPA"), 
                  type == "All races") 

min(pfhpa_gene_interaction$p_adjust2, na.rm = TRUE)
min(pfhpa_gene_interaction$p_adjusted, na.rm = TRUE)
# .12

# writexl::write_xlsx(all_result_new, fs::path(dir_result, "PFAS and SNP interaction_1208.xlsx"))

```


# Plot PFAS and SNP interaction results
## All Races
```{r, fig.width=19, fig.height=7}

plot_pfas_snp <- function(sample_type){
  i <- 1
  plot <- list()
  for(i in 1:length(pfas_name)){
  pfas = pfas_name[i]
  pfas_gene_interaction <- all_result_new |> 
    tidylog::filter(exposure == pfas&
                    # effect_type == "PFAS and SNP", 
                    type == sample_type) |> 
    mutate(effect_type = fct_relevel(effect_type, 
                                     "PFAS", "Genetic Risk", "PFAS and SNP"))

  # Plot 
  
  p <- pfas_gene_interaction |> 
    mutate(gene_snp = str_c(rsid, "\n", gene)) |>
    # tidylog::filter(snp %in% sig_int$snp) |>  
    ggplot(aes(x = effect_type, 
               y = odds_ratio, 
               color = p.value < 0.05)) +
    geom_point(position = position_dodge(width = .3)) +
    geom_errorbar(aes(ymin = exp_ci_low, ymax = exp_ci_high), width = 0) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    facet_grid(~gene_snp) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
    labs(title = paste0(sample_type, ": ", pfas, " and SNP Interaction"),
         x = "Effect",
         y = "Odds Ratio",
         color = "Significance") +
    scale_color_manual(values = c("grey60", "red"))
  plot[[i]] <- p
  i = i + 1
  }
  return(plot)
}
sample_types <- unique(all_result_new$type)

p1 <- plot_pfas_snp(sample_types[1])
p2 <- plot_pfas_snp(sample_types[2])
p3 <- plot_pfas_snp(sample_types[3])

p1[[1]]
p1[[2]]
p1[[3]]
p1[[4]]
p1[[5]]
p1[[6]]
p1[[7]]
p1[[8]]
p1[[9]]
p1[[10]]
p1[[11]]
p1[[12]]
```

## Japanese American
```{r,fig.width=19, fig.height=7}
p2[[1]]
p2[[2]]
p2[[3]]
p2[[4]]
p2[[5]]
p2[[6]]
p2[[7]]
p2[[8]]
p2[[9]]
p2[[10]]
p2[[11]]
p2[[12]]
```

## Latino
```{r,fig.width=19, fig.height=7}
p3[[1]]
p3[[2]]
p3[[3]]
p3[[4]]
p3[[5]]
p3[[6]]
p3[[7]]
p3[[8]]
p3[[9]]
p3[[10]]
p3[[11]]
p3[[12]]
```

# Interaction plot of PFNA and APOC1
```{r}
# Anova: if interaction is significant
interaction_models_snp <- function(pfas) {
  # Formula
  base_covars <- paste(paste(covars, collapse = " + "), 
                       "+ strata(setnum)")
  snp = "x19_44915533"
  frm_no_int <- as.formula(paste0("status ~", snp,"+", paste(pfas, base_covars, sep = " + ")))
  frm_with_int <- as.formula(paste("status ~", paste0(pfas, "*",snp), base_covars, sep = " + "))
  
  # Modeling
  model_no_int   <- clogit(frm_no_int, data = data_hcc)
  model_with_int <- clogit(frm_with_int, data = data_hcc)
  
  # ANOVA
  anova_res <- anova(model_no_int, model_with_int)%>% 
    tidy(conf.int = TRUE) %>%
    mutate(pfas_name = pfas)%>%
    select(pfas_name, everything())
  
  return(anova_res)
}

(outputs <- interaction_models_snp("pfas_t_PFNA")%>%
    mutate(model = ifelse(grepl("\\*",term), "With Interaction", "Without Interaction"),
           pfas_name = rename_pfas(pfas_name)) %>%
    select(pfas_name, model, everything(),-term)
    )


```

# Plot PFAS and SNP interaction results

```{r,fig.width=15, fig.height=4}
# Model with interaction of prs_wt and pfas
model_interaction <- function(pfas){
  # Formula
  base_covars <- paste(paste(covars, collapse = " + "), 
                       "+ strata(setnum)")
  snp = "x19_44915533"
  frm_with_int <- as.formula(paste("status ~", paste0(pfas, "*",snp), base_covars, sep = " + "))
  # Model
  model_with_int <- clogit(frm_with_int, 
                           data = data_hcc, 
                           method = "efron", 
                           robust = TRUE)
  # Interaction effect at high/low PRS
  res <- emtrends(model_with_int, ~x19_44915533, var = pfas, 
                  at = list(`x19_44915533` = c(0,1,2), 
                            smokestatus_imputed = "Never",
                            diabetes = "No",
                            q1_edih = 0,
                            prs_wt = 0,
                            v1 = 0,
                            v2 = 0,
                            v3 = 0,
                            v4 = 0,
                            v5 = 0,
                            v6 = 0),
                  transform = NULL) %>%
    data.frame() %>%
    select(`x19_44915533`, ends_with(".trend"), asymp.LCL, asymp.UCL) %>%
    rename(estimate = ends_with(".trend"),
           conf_low = asymp.LCL,
           conf_high = asymp.UCL) %>%
    mutate(type = case_when(`x19_44915533` == 0 ~ "0 rs5117-T alleles",
                            `x19_44915533` == 1 ~ "1 rs5117-T allele",
                            `x19_44915533` == 2 ~ "2 rs5117-T alleles"))
  
  # Main effect without interaction
  res_main <- epiomics::owas_clogit(data_hcc , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars,
                                    omics = pfas,
                                    conf_int = TRUE) %>%
    mutate(type = "Overall (no interaction with SNP)")
  # Combine results
  res_combined <- bind_rows(res_main, res) %>%
    mutate(odds_ratio = exp(estimate),
           exp_ci_low = exp(conf_low),
           exp_ci_high = exp(conf_high),
           group = if_else(str_detect(type, "Overall"), "overall", "with genetic"),
           pfas = pfas)
  
  res_combined
}

res <- map(pfas_name, model_interaction)%>%
    bind_rows()%>%
    mutate(pfas = str_remove(pfas,"pfas_t_"))

# writexl::write_xlsx(res, fs::path(dir_result, "pfas_prs_interaction_result.xlsx"))

res %>% 
       select(pfas, type, odds_ratio, exp_ci_low, exp_ci_high, p_value)%>%
       filter(pfas == "PFNA")

(p <- res  %>%
    filter(pfas == "PFNA") %>%
    ggplot(aes(x = factor(type, 
                          levels = c("2 rs5117-T alleles", 
                                     "1 rs5117-T allele", 
                                     "0 rs5117-T alleles",
                                     "Overall (no interaction with SNP)")),
               y = exp(estimate))) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = exp(conf_low),
                      ymax = exp(conf_high)),
                  width = 0) +
    geom_hline(yintercept = 1, linetype = 2, color = "grey") +
    ylab("Odds Ratio (95% CI)") +
    facet_grid(group ~. , scales = "free_y", space = "free_y")+
    theme(panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_blank(),
          # strip.text.x = element_blank(),
          text = element_text(size = 12),
          # axis.text.x =element_text(angle = 60, hjust = 1),
          axis.line.x = element_line(color = "black"),
          axis.title.y=element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
    coord_flip()
)


# ggsave(filename = fs::path(dir_figures,
#                            "Fig2. Interaction of PFNA with SNP_072426.jpeg"),
#        width =5,
#        height = 2.5,
#        dpi = 600,
#        bg = "white")

(combined_plot <- cowplot::plot_grid(
                                    NULL,
                                    NULL,
                                    p,
                                    p_dist, 
                                    labels = c("A. PFNA-HCC Risk by rs5117-T Alleles", "B. Distribution of rs5117-T Alleles", NULL, NULL), 
                                    ncol = 2,
                                    label_x = c(-0.29, -0.16),
                                    rel_widths = c(1.2, 1),
                                    rel_heights = c(0.1, 1,0.1,  1)
                                    ))

ggsave(filename = fs::path(dir_figures,
                           "Fig2. Inter PFNA SNP and Dist SNP_072426.jpeg"),
       width =10,
       height = 4,
       dpi = 600,
       bg = "white")
```


