---
title: 'MEC analysis'
subtitle: "Association of interactions between Sex and PFAS with HCC"
author: "Hongxu Wang, Jesse Goodrich"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("1_Project_Setup/!libraries.R")))
source(fs::path(here::here("1_Project_Setup/!directories.R")))
source(fs::path(here::here("1_Project_Setup/!functions.R")))
source(fs::path(here::here("1_Project_Setup/!load_data.R")))
```

## Check for missing PFAS values
```{r}
data_pfas_missing <- data_hcc %>% filter(is.na(pfas_t_PFDA))
data_hcc_c <- data_hcc %>% filter(!setnum %in% data_pfas_missing$setnum)

# Data with complete genetic data based on data with complete PFAS data (data_hcc_c)
data_gene_missing <- data_hcc_c %>% filter(is.na(prs_wt))
data_hcc_c_gene <- data_hcc_c %>% filter(!setnum %in% data_gene_missing$setnum)

rm(data_hcc_c)
```


# Analysis 1: ANOVA test for Sex x PFAS interaction
**Outcome: HCC vs control**

**Exposure: normalized log2 PFAS**

**Covars: smoking, diabetes, q1_edih, prs_wt_std, v1:v6, eth, studyarea, q1_age_cohent**

**Analysis: logistic regression (owas) - matching broken by sex stratification**

**Model1: status ~ pfas + sex + covars**

**Model2: status ~ pfas * sex + covars**

```{r anova sex interaction}
covars_sex <- c(covars, covars_matched[2:4])

# Anova: if interaction is significant
interaction_models_sex <- function(pfas) {
  # Formula
  base_covars <- paste(paste(covars_sex, collapse = " + "))

  frm_no_int <- as.formula(paste("status ~ sex +", paste(pfas, base_covars, sep = " + ")))
  frm_with_int <- as.formula(paste("status ~", paste0(pfas, "*sex"), "+", base_covars))

  # Modeling
  model_no_int   <- glm(frm_no_int, data = data_hcc, family = binomial())
  model_with_int <- glm(frm_with_int, data = data_hcc, family = binomial())

  # ANOVA
  anova_res <- anova(model_no_int, model_with_int, test = "Chisq") %>%
    tidy(conf.int = TRUE) %>%
    mutate(pfas_name = pfas) %>%
    select(pfas_name, everything())

  return(anova_res)
}

(outputs_sex <- map(c(pfas_name, "pfas_t_PFBS_detected"), interaction_models_sex) %>%
    bind_rows() %>%
    mutate(model = ifelse(grepl("\\*", term), "With Interaction", "Without Interaction"),
           pfas_name = rename_pfas(pfas_name)) %>%
    select(pfas_name, model, everything(), -term)
)

# Calculate FDR q-values for the interaction p-values
outputs_sex_inter <- outputs_sex %>%
  filter(model == "With Interaction") %>%
  mutate(q_value = p.adjust(p.value, method = "fdr"))

outputs_sex_inter %>% select(pfas_name, p.value, q_value)

writexl::write_xlsx(outputs_sex_inter, fs::path(dir_result, "anova_sex_pfas_inter.xlsx"))
```


# Analysis 2: Sex x PFAS interaction model (continuous PFAS)

**Model: status ~ pfas * sex + covars**

```{r interaction model sex}
covars_sex <- c(covars, covars_matched[2:4])
covariate_input_sex <- paste(covars_sex, collapse = " + ")

# create dataframe with names of dependent vars (PFAS)
pfas_names <- tibble(exposure = c(pfas_name, "pfas_t_PFBS_detected"))

## create dataframe with sex var
sex_var <- tibble(modifier = "sex")

pfas_sex <- expand_grid(pfas_names, sex_var) %>%
  mutate(pfas_sex = paste(exposure, "*", modifier))

# Get Model covariates
covar_names <- tibble(covariates = covariate_input_sex)

# Make all logistic regression formulas
ind_vars <- expand_grid(pfas_sex, covar_names) |>
  mutate(formula = paste("status ~",
                         pfas_sex,
                         "+", covariates) |>
           str_replace("\\+  \\+", " + "))

ind_vars$glm_out <- map(
  ind_vars$formula,
  ~glm(as.formula(.),
       data = data_hcc,
       family = binomial()
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>%
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "logistic"))

# Get final data
all_results_df_sex <- ind_vars %>%
  unnest(glm_out) %>%
  dplyr::select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig",
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high))

res_inter_sex <- all_results_df_sex %>%
  filter(grepl(":", term)) %>%
  group_by(exposure) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         sig = case_when(p_adjusted < 0.05 ~ "FDR. Sig.",
                         p.value < 0.05 ~ "Sig.",
                         TRUE ~ "Not Sig."))

writexl::write_xlsx(res_inter_sex, fs::path(dir_result, "sex_pfas_interaction_results.xlsx"))
```


### Fig. Sex x PFAS interaction terms
```{r plot interaction sex, fig.width=10, fig.height=5}
res_inter_sex_plot <- res_inter_sex %>%
  mutate(pfas_name = str_remove(exposure, "pfas_t_"),
         pfas_name = rename_pfas(pfas_name)) %>%
  ungroup()

(p_inter_sex <- res_inter_sex_plot %>%
    filter(!pfas_name %in% c("PFBS", "PFBS_detected")) %>%
    ggplot(aes(x = pfas_name, y = odds_ratio, color = sig)) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = exp_ci_low, ymax = exp_ci_high),
                  width = 0) +
    geom_hline(yintercept = 1, linetype = 2, color = "grey") +
    ylab("Odds Ratio (95% CI) for Sex x PFAS Interaction") +
    theme(panel.background = element_rect(fill="white"),
          strip.background = element_rect(fill = "white"),
          text = element_text(size = 15),
          axis.line.x = element_line(color = "black"),
          axis.title.y = element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
    coord_flip() +
    scale_color_manual(values = c("grey", "red", "purple")))

# ggsave(filename = fs::path(dir_figures,
#                            "Fig_Sex_PFAS_Interaction.jpeg"),
#        width = 10,
#        height = 5,
#        dpi = 300,
#        bg = "white")
```

