---
title: 'MEC analysis'
subtitle: "Final Analysis"
author: "Hongxu Wang, Jesse Goodrich"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup univariate analysis cat pfas, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("1_Project_Setup/!libraries.R")))
source(fs::path(here::here("1_Project_Setup/!directories.R")))
source(fs::path(here::here("1_Project_Setup/!functions.R")))
source(fs::path(here::here("1_Project_Setup/!load_data.R")))

```

## Check for missing PFAS values and rerun descriptive statistics using the complete PFAS data
* Complete PFAS means both control and case (same setnum) have PFAS values
```{r}
data_pfas_missing <- data_hcc %>% filter(is.na(pfas_t_PFDA))
data_hcc_c <- data_hcc %>% filter(!setnum %in% data_pfas_missing$setnum)
table(data_hcc_c$eth)


# Data with complete genetic data based on data with complete PFAS data (data_hcc_c)
data_gene_missing <- data_hcc_c %>% filter(is.na(prs_wt))
data_hcc_c_gene <- data_hcc_c %>% filter(!setnum %in% data_gene_missing$setnum)
```

# HEP status
```{r}
hep <- read_sas(fs::path(
  dir_cleaned_data,
  "hhear_2020_00500_hep_072426.sas7bdat"))

# for (v in names(hep)) {
#   lab <- attr(hep[[v]], "label")
#   
#   if (!is.null(lab)) {
#     Hmisc::label(hep[[v]]) <- paste0(v, ": ", lab)
#   } else {
#     Hmisc::label(hep[[v]]) <- v
#   }
# }

var_labels <- sapply(
  hep,
  function(x) attr(x, "label")
)

hep <- hep %>%
  janitor::clean_names() %>%
  mutate(across(5:11, haven::as_factor))

# restore variable labels
names(var_labels) <- names(hep)

for (v in names(var_labels)) {
  if (!is.na(var_labels[v])) {
    attr(hep[[v]], "label") <- var_labels[v]
  }
}

data_full <- data_hcc %>% 
  tidylog::left_join(hep %>%
                       select(-setnum, -status, -casetype), by = "barcode")
data_analysis <- data_hcc_c_gene %>% 
  tidylog::left_join(hep %>%
                       select(-setnum, -status, -casetype), by = "barcode")

table1::table1(~ . |status, data = hep %>% 
                 filter(setnum %in% unique(data_hcc$setnum)) %>%
                 select(-setnum,-barcode))


table1::table1(~ . |status, data = hep %>% 
                 filter(setnum %in% unique(data_hcc_c_gene$setnum)) %>%
                 select(-setnum,-barcode))
```

## Analysis1. Viral HCC analysis
**Outcome: viral HCC vs control**

**Exposure: normalzied log2 PFAS**

**Covars: smoking, diabetes, q1_edih, prs_wt_std, v1:v6**

**Analysis: conditinal logistic regression**

```{r analysis1}
viral <- data_analysis %>% filter(hep_b_c_9919 == 1)
data_hcc_viral <- data_full %>% filter(setnum %in% viral$setnum)

res_hcc_viral <- epiomics::owas_clogit(data_hcc_viral, 
                                 cc_set = "setnum", 
                                 cc_status = "status", 
                                 covars = covars,
                                 omics = c(pfas_name[-3], "pfas_t_PFBS_detected"),
                                 conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high))

# reorder PFAS by estimate
res_hcc_viral1 <- res_hcc_viral |> 
  mutate(pfas_name = str_remove(feature_name, "pfas_t_"),
         sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "FDR p value < 0.05",
                         p_value < 0.05 ~ "p value < 0.05", 
                         p_value >= 0.05 ~ "p value ≥ 0.05"),
         sig = factor(sig, levels = c("p value ≥ 0.05", "p value < 0.05", "FDR p value < 0.05"))) %>%
  mutate(covar = sub(".*~(.*?)\\+pfas.*", "\\1", formula))%>%
  mutate(group = "Viral Cases") %>%
  select(group, covar, everything())
  
```

## Analysis2. non-Viral HCC analysis
**Outcome: non-viral HCC vs control**

**Exposure: normalzied log2 PFAS**

**Covars: smoking, diabetes, q1_edih, prs_wt_std, v1:v6**

**Analysis: conditional logistic regression**


```{r analysis1}
non_viral <- data_analysis %>% 
  filter(status == 1) %>% 
  filter(hep_b_c_9919 == 0)

data_hcc_non_viral <- data_hcc %>% filter(setnum %in% non_viral$setnum)

res_hcc_non_viral <- epiomics::owas_clogit(data_hcc_non_viral, 
                                 cc_set = "setnum", 
                                 cc_status = "status", 
                                 covars = covars,
                                 omics = c(pfas_name[-3],"pfas_t_PFBS_detected"),
                                 conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high))

# reorder PFAS by estimate
res_hcc_non_viral1 <- res_hcc_non_viral |> 
  mutate(pfas_name = str_remove(feature_name, "pfas_t_"),
         sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "FDR p value < 0.05",
                         p_value < 0.05 ~ "p value < 0.05", 
                         p_value >= 0.05 ~ "p value ≥ 0.05"),
         sig = factor(sig, levels = c("p value ≥ 0.05", "p value < 0.05", "FDR p value < 0.05"))) %>%
  mutate(covar = sub(".*~(.*?)\\+pfas.*", "\\1", formula))%>%
  mutate(group = "Non-Viral Cases")%>%
  select(group,covar, everything())

res_hcc <- res_hcc_viral1 %>%
  bind_rows(res_hcc_non_viral1)
```


### Fig1a. association of pfas and hcc - HCC
```{r plot with covars only, fig.width=10, fig.height=5}
### Rename PFAS for plotting
res_hcc2 <- res_hcc %>%
  mutate(pfas_name = rename_pfas(pfas_name))%>%
  mutate(group = factor(group, levels = c("Viral Cases", "Non-Viral Cases")))


(res_hcc2  %>%
    mutate(pfas_name = fct_reorder(pfas_name, if_else(group == "Viral Cases", estimate, NA_real_)))%>%
    filter(pfas_name != "PFBS_detected")%>%
    ggplot(aes(x = pfas_name,y = exp(estimate), color = sig)) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = exp(conf_low),
                      ymax = exp(conf_high)),
                  width = 0) +
    geom_hline(yintercept = 1, linetype = 2, color = "grey") +
    ylab("Odds Ratio (95% CI)") +
    # ggtitle("Exposure Outcome Analysis for HCC Cases stratified by HEP status") +
    facet_grid(.~ factor(group), scales = "free_y")+
    # facet_grid(.~ factor(type, levels = c("Overall \n(n = 148)", "Japanese American \n(n = 176)", "Latino \n(n = 74)", "White \n(n = 46)", "African American \n(n = 50)", "Native Hawaiian \n(n = 34)")), scales = "free")+ 
    theme(panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_text(angle = 0, hjust = 0),
          # strip.text.x = element_blank(),
          text = element_text(size = 15),
          # axis.title.x=element_blank(),
          title = element_text(size = 15),
          axis.line.x = element_line(color = "black"),
          axis.title.y=element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
    coord_flip() +
    scale_color_manual(values = c("grey","red","purple")))


ggsave(filename = fs::path(dir_figures,
                           "Fig 5 Single Exposure Analysis HCC with covars_stratified_by_hep_status_0727.jpeg"),
       width = 10,
       height = 5,
       dpi = 300,
       bg = "white")

writexl::write_xlsx(res_hcc2, fs::path(dir_result, "T5.Continuous PFAS and HCC_Stratified_by_hep_status.xlsx"))
```

