---
title: 'MEC analysis'
subtitle: "Final Analysis"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
```{r setup univariate analysis cat pfas, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("!libraries.R")))
source(fs::path(here::here("!directories.R")))
source(fs::path(here::here("functions.R")))
source(fs::path(here::here("!load_data.R")))
```


## Analysis1. Conditional logistic regression - HCC
**Outcome: HCC vs control**

**Exposure: lg2 transformed and then normalzied PFAS**

**Covars: bmi, smoking status, alcohol intake, diabetes**

**Analysis: conditional logistic regression**

**Data Info: 237 cases, 237 controls**

**All races(474), White(46), African American(52), Japanese American(182), Latino(160), Native Hawaiian(34)**

```{r analysis1}
data_hcc <- data_hcc %>% 
  mutate_at(.vars = c("q1_fdgp6", "q1_fdgp7"),
            .funs = ~ifelse(. < median(., na.rm = TRUE), "Low", "High") %>%
             factor(., levels = c("Low", "High")))

res_hcc <- epiomics::owas_clogit(data_hcc , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars[-6],
                                    omics = pfas_name,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "Overall")

eth_categories <- unique(data_hcc$eth_new)

for(race in eth_categories) {
  res_hcc_temp <- epiomics::owas_clogit(data_hcc %>% filter(eth_new == race) , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars[-6],
                                    omics = pfas_name,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = race)
  res_hcc <- res_hcc %>% bind_rows(res_hcc_temp)
}


# reorder PFAS by estimate
res_hcc1 <- res_hcc |> 
  mutate(pfas_name = str_remove(feature_name, "pfas_lg2_"),
         sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "FDR p value < 0.05",
                          p_value < 0.05 ~ "p value < 0.05", 
                         p_value >= 0.05 ~ "p value ≥ 0.05"),
         sig = factor(sig, levels = c("p value ≥ 0.05", "p value < 0.05", "FDR p value < 0.05"))) 

```

### Fig1a. association of pfas and hcc - HCC
```{r plot with covars only, fig.width=10, fig.height=5}
### Create metabolites Coefficient Plots
### Rename PFAS for plotting
res_hcc2 <- res_hcc1 %>%
  mutate(pfas_name = case_when(pfas_name == "4H_PFBA" ~ "4H-PFBA",
                               pfas_name == "10_2_FToH" ~ "10:2 FTOH",
                               pfas_name == "2_aminohexafluoropropan_2_ol" ~ "2-Aminohexafluoropropan-2-ol",
                               pfas_name == "Me_PFHxA" ~ "MePFHxA",
                               TRUE ~ pfas_name),
         ) %>%
  mutate(pfas_name = fct_reorder(pfas_name, if_else(type == "Overall", estimate, NA_real_))) %>%
  mutate(type = case_when(type == "Overall" ~ "Overall \n(n = 474)",
                          type == "Japanese American" ~ "Japanese American \n(n = 182)", 
                        type == "Latino"~ "Latino \n(n = 160)",
                        type == "Others" ~ "Others \n(n = 132)"))

res <- res %>% bind_rows(res_hcc2 %>% mutate(covars = "Fish excluding shellfish (categorical)"))

res1 <- res %>% select(covars, everything())%>%
  mutate_if(is.numeric, round, digits = 3)%>%
  mutate(covars = ifelse(covars =="Shellfish", "Shellfish (Continuous)", covars)) %>%
  mutate(covars = ifelse(covars =="Fish Excluding Shellfish", "Fish Excluding Shellfish (Continuous)", covars))

writexl::write_xlsx(res1, fs::path(dir_result, "pfas_hcc_with_diff_covars.xlsx"))


(pfhpa_hcc_main <- res_hcc2  %>%
  ggplot(aes(x = pfas_name,y = exp(estimate), color = sig)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = exp(conf_low),
                    ymax = exp(conf_high)),
                width = 0) +
  geom_hline(yintercept = 1, linetype = 2, color = "grey") +
  ylab("Odds Ratio (95% CI)") +
  # ggtitle("Exposure Outcome Analysis for HCC Cases") +
  facet_grid(.~ factor(type, levels = c("Overall \n(n = 474)", "Japanese American \n(n = 182)", "Latino \n(n = 160)", "Others \n(n = 132)")), scales = "free_y")+
  theme(panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 15),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.title.y=element_blank(),
        axis.line.y = element_line(color = "black"),
        legend.position = "bottom",
        legend.title = element_blank())+
  coord_flip() +
  scale_color_manual(values = c("grey","red","purple")))


# ggsave(filename = fs::path(dir_figures,
#                            "Fig 1a Single Exposure Analysis HCC with covars_stratified_by_eth.jpeg"),
#        width = 10,
#        height = 5,
#        dpi = 300,
#        bg = "white")
# 
writexl::write_xlsx(res_hcc2, fs::path(dir_result, "T1.Continuous PFAS and HCC.xlsx"))
```


## Analysis2. Conditional logistic regression - Categorized PFAS and HCC
**Outcome: HCC vs control**

**Exposure: categorized pfas(quartile)**

**Covars: bmi, smoking status, alcohol intake, diabetes**

**Analysis: conditional logistic regression**

**Data Info: 237 cases, 237 controls**

**All races(474), White(46), African American(52), Japanese American(182), Latino(160), Native Hawaiian(34)**

```{r analysis1 cat}
covariate_input <- " alcohol_intake + scale(bmi) + diabetes + smokestatus + q1_dp_amds_e_totscore" 

## create dataframe with names of dependent vars (PFAS) ----
pfas_names <- tibble(exposure = pfas_name_cat)

# Get Model covariates 
covar_names <- tibble(
  covariates =  covariate_input)

# Make all clogit formulas
ind_vars <- expand_grid(pfas_names,  covar_names) |>  
  mutate(formula = paste("status ~ ", 
                         exposure, 
                         "+", covariates, 
                         "+strata(setnum)") |>
           str_replace("\\+  \\+", " + "))

## Analysis for all related participants
ind_vars$clogit_out <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_hcc,
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Japanese American 
ind_vars$clogit_out1 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_hcc %>% filter(eth == "Japanese American"),
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Latino
ind_vars$clogit_out2 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_hcc %>% filter(eth == "Latino"),
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Others
ind_vars$clogit_out3 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_hcc %>% filter(eth_new == "Others"),
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit"))
# Get final data -------------------

intercept_data<- tibble(exposure = pfas_name_cat, 
                        term = paste0(pfas_name_cat, "1"), 
                        odds_ratio = rep(1,length(pfas_name_cat)),
                        effect_est  = rep("1",length(pfas_name_cat)))

all_results_df_hcc <- ind_vars %>% 
  unnest(clogit_out) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out1, -clogit_out2) %>%
  filter(grepl("pfas", term)) %>%
  bind_rows(intercept_data) %>% 
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
        type = "Overall")

all_results_df_hcc_japan <- ind_vars %>% 
  unnest(clogit_out1) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out2) %>%
  filter(grepl("pfas", term))%>%
  bind_rows(intercept_data) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type= "Japanese American")

all_results_df_hcc_latino <- ind_vars %>% 
  unnest(clogit_out2) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out1) %>%
  filter(grepl("pfas", term))%>%
  bind_rows(intercept_data) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type = "Latino")

all_results_df_hcc_others <- ind_vars %>% 
  unnest(clogit_out3) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out1) %>%
  filter(grepl("pfas", term))%>%
  bind_rows(intercept_data) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type = "Others")


all_result_df <- all_results_df_hcc %>% 
  bind_rows(all_results_df_hcc_japan) %>%
  bind_rows(all_results_df_hcc_latino) %>%
  bind_rows(all_results_df_hcc_others) %>%
  select(exposure, term, odds_ratio, exp_ci_low, exp_ci_high,effect_est,p.value,p_adjusted, type) %>%
  arrange(exposure)

```

```{r does response}
data_hcc1 <- data_hcc %>% mutate_at(.vars = pfas_name_cat,
                                    .funs = ~as.numeric(.))
res_hcc <- epiomics::owas_clogit(data_hcc1 , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars,
                                    omics = pfas_name_cat,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "Overall")

eth_categories <- unique(data_hcc1$eth_new)

for(race in eth_categories) {
  res_hcc_temp <- epiomics::owas_clogit(data_hcc1 %>% filter(eth_new == race) , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars,
                                    omics = pfas_name_cat,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = race)
  res_hcc <- res_hcc %>% bind_rows(res_hcc_temp)
}

# reorder PFAS by estimate
res_hcc1 <- res_hcc |> 
  mutate(sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "Sig.(FDR < 0.05)",
                          p_value < 0.05 ~ "Sig.(p value < 0.05)", 
                         p_value >= 0.05 ~ "Not Sig.")) 

p_trend_df <- res_hcc1 %>% select(feature_name, p_value, adjusted_pval,type) %>%
  rename(exposure = feature_name,
         p_trend = p_value)

all_result_df1 <- all_result_df %>% 
  tidylog::left_join(p_trend_df, by = c("exposure", "type")) %>%
  mutate(pfas_name = str_remove(exposure, "pfas_lg2_") %>%
           str_remove("_quartile")) %>%
  select(type,pfas_name, term,everything(), -exposure)

## N value for each quartile split by status
# Assuming df is your dataframe

# Convert to long format and create a pair identifier
data_list <- list(data1 = data_hcc %>% mutate(type = "Overall"), 
                  data2 = data_hcc %>% filter(eth == "Japanese American") %>%
                    mutate(type = "Japanese American"), 
                  data3 = data_hcc %>% filter(eth == "Latino") %>%
                    mutate(type = "Latino"),
                  data4 = data_hcc %>% filter(eth_new == "Others") %>%
                    mutate(type = "Others"))

data_n <- data.frame()

for(data in data_list){
  type = unique(data$type)
  long_data_hcc <- data %>%
    select(contains("quartile"), status) %>%
    mutate(status = ifelse(status == 0, "Control", "Case")) %>%
    pivot_longer(cols = c(contains("quartile")), names_to = "variable",
                 values_to = "value") %>%
    drop_na() %>%
    mutate(pair = paste0(variable, value)) %>%
    select(-variable,-value) %>%
    group_by(pair, status) %>%
    summarise(count = n(), .groups = "drop")
  # Now, to spread back for a wide format summary if needed
  wide_summary_hcc <- pivot_wider(long_data_hcc, names_from = status,
                                  values_from = count, values_fill = list(count = 0))%>%
    mutate(type = type) %>%
    rename(term = pair)
  data_n <- data_n %>% bind_rows(wide_summary_hcc)
}

all_result_df2 <- all_result_df1 %>% 
  tidylog::left_join(data_n, by = c("term","type")) %>%
  select(type:term, Control, Case, everything())%>%
  arrange(pfas_name, type, term)

```

### Fig 1b. Dose response plot for PFHpA and HCC
```{r dose response plot}
# Plot interaction effects
all_result_df3 <- all_result_df2 %>% 
  mutate(plot_name = substr(term, 
                            start = nchar(term) - 8,
                            stop = nchar(term)) %>% toTitleCase()) %>%
  mutate(
         plot_name = paste(substr(plot_name, 1, 8), substr(plot_name, 9, 9)),
         sig = case_when(p_adjusted < 0.05 ~ "FDR. Sig.", 
                         p.value < 0.05 ~ "Sig.",
                         TRUE ~ "Not Sig."),
         label = str_c("p trend = ", format(signif(p_trend, digits = 2), scientific = FALSE))) %>%
  mutate(type = case_when(type == "Overall" ~ "Overall \n(n = 474)",
                          type == "Japanese American" ~ "Japanese American \n(n = 182)", 
                        type == "Latino"~ "Latino \n(n = 160)",
                        type == "Others" ~ "Others \n(n = 132)"))

# writexl::write_xlsx(all_result_df3, fs::path(dir_result, "single_categorical_exposure_analysis.xlsx"))

(pfhpa_hcc <-  ggplot(all_result_df3 %>% filter(pfas_name == "PFHpA"),
       aes(x = odds_ratio, y = fct_rev(plot_name))) +
  geom_errorbar(aes(xmin = exp_ci_low, xmax = exp_ci_high),
                width = 0, position = position_dodge(width = 1)) +
  geom_point(size = 1, position = position_dodge(width = 1)) + 
  geom_vline(xintercept = 1, linetype = 2, color = "grey") +
  facet_grid( .~ factor(type, levels = c("Overall \n(n = 474)", "Japanese American \n(n = 182)", "Latino \n(n = 160)", "Others \n(n = 132)")))+
  xlab("Odds Ratio (95% CI)") +
  theme(#axis.title.x = element_blank(),
        # axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), 
        legend.position = "bottom",
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black")) )
   # +
    # ylab("OR (95% CI) per\ndoubling of PFAS") + 
    # ggtitle("Interaction of PFAS and PRS") +
  # scale_color_manual(values = c("purple", "black","red")))


p1 <- pfhpa_hcc + geom_text(
  data    = all_result_df3 %>% filter(pfas_name == "PFHpA"),
  mapping = aes(x = -Inf, y = -Inf, label = label),
  color = "black",
  hjust   = -0.3,
  vjust   = -13,
  size = 4
)

# (fig <- cowplot::plot_grid(NULL, pfhpa_hcc_main, NULL, 
#                            p1,
#                            ncol = 1, 
#                            rel_heights = c(0.02, 0.2, 0.02, 0.12),
#                            labels = c("A. PFAS and HCC","", "B. Dose-response of PFHpA and HCC", ""),
#                           label_x = c(0.13, 0, 0.04, 0),
#                           label_size = 12,
#                           align = "v",
#                           axis = "lt"))
```

## Fig1. Association of single PFAS and HCC and Does response analysis
```{r combine figures}
(fig <- cowplot::plot_grid(pfhpa_hcc_main, 
                           p1,
                           ncol = 1, 
                           rel_heights = c(0.2, 0.12),
                           labels = c("A. PFAS and HCC",
                                      "B. Dose-response of PFHpA and HCC"),
                          # label_x = c(-0.08,-0.18),
                          label_x = c(-0.06, -0.14),
                          label_size = 12,
                          align = "v",
                          axis = "lt"))

ggsave(filename = fs::path(dir_figures,
                           "main_figures",
                           "Fig1. Main Effect New.jpeg"),
       width = 10,
       height = 8,
       dpi = 300,
       bg = "white")
```

