---
title: 'MEC Project Data Description'
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{css, echo=FALSE}
pre {
  max-height: 200px;
  overflow-y: auto;
}

```

```{r data setup, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE)

source(fs::path(here::here("!libraries.R")))
source(fs::path(here::here("!directories.R")))

# suppress warning messages for final rendering
old.warn <- getOption("warn")
options(warn=-1)

## Load raw data
data_raw <- readxl::read_xlsx(fs::path(dir_data %>% dirname() %>% dirname() %>% dirname(), "0_data_processing", "0_raw_data",
                                   "Exposures_phentypes",
                                   "p27_hhear_pfas_covar_230303",
                                   "p27_hhear_pfas_covar_230303.xlsx"))

## Load cleaned data
data_cleaned <- readRDS(fs::path(
          dir_cleaned_data,
          "p27_hhear_pfas_covar_imputed_V2.RDS"))

```


## Data Overview
This study included 1400 cases and 1400 controls from Multiethnic Cohort (MEC) study. Cases and controls were matched by:

1) Birth year
2) Date of blood collection (+/– 6 months)
3) Sex
4) Race/ethnicity
5) Fasting status
6) Study area
7) Control must be alive at age of case diagnosis
8) Controls with invasive cancer of any site prior to blood draw were excluded

Data files included:

1) Exposure Phentypes dataset
2) Metabolomics: Untargeted Feature dataset
3) Metabolomics: Targeted Feature dataset


## Exposure-phenotype Data Dictionary 
1) PFAS
  + PFDA, PFDS, PFHpA, PFHpS, PFHxS, PFNA, PFOA, PFOS, PFPeS, PFTrDA and PFUdA
2) Phenotype
  + status, casetype
```{r load data and coddebook, echo=TRUE}
# Load data dictionary
sas_data <- haven::read_sas(fs::path(dir_data %>% dirname() %>% dirname() %>% dirname(),
                                     "0_data_processing",
                                     "0_raw_data",
                                     "Exposures_phentypes",
                                     "p27_hhear_pfas_covar_230303",
                                     "p27_hhear_pfas_covar_230303.sas7bdat"))

data_dictionary <- labelled::generate_dictionary(sas_data) %>%
  mutate(variable = tolower(variable)) %>%
  mutate(variable = case_when(variable == "area" ~ "studyarea",
                              variable == "q1_corr_sex" ~ "sex",
                              variable == "q1_eth" ~ "eth",
                              variable == "q1_newqi_corr2" ~ "bmi",
                              variable == "q1_ethanol" ~ "alcohol_intake",
                              variable == "q1_diabet" ~ "diabetes",
                              variable == "q1_smokstat" ~ "smokestatus",
                              variable == "dx_age" ~ "diagnosis_age",
                              TRUE ~ variable )) %>%
  select(pos:col_type) %>%
  add_row(pos = 38,
          variable = "year_between", 
          label = "Years between blood collection and diagnosis",
          col_type = "dbl") %>%
  add_row(pos = 39,
          variable = "bmi_cat", 
          label = "Body mass index (<25, 25-30, ≥30)",
          col_type = "chr") %>%
  add_row(pos = 40,
          variable = "alcohol_intake_cat", 
          label = "Alcohol (g/day) (0, <12, ≥12)",
          col_type = "chr") %>%
  mutate(label = case_when(variable =="casetype" ~ "Case Type (Control, NAFLD, NAFLD-related cirrhosis, NAFLD-related HCC, Other HCC)",
                           variable == "status" ~ "Case-control status (Control, NAFLD, Cirrhosis, HCC)",
                           variable == "sex" ~ "Preferred Sex Variable (Male, Female)",
                           variable == "eth" ~ "Ethnicity (African American, Native Hawaiian, Latino, Japanese American, White)",
                           variable == "smokestatus" ~ "Smoking status (Never, Former, Current)",
                           TRUE ~ label),
         col_type = case_when(
                           variable == "status" ~ "chr",
                           variable == "smokestatus" ~ "chr",
                           TRUE ~ col_type)) 
data_dictionary1 <- data_dictionary[c(1:4, 6:12,5, 24:29,39,30,40,31:38,13:23),] %>%
  mutate(pos = 1:nrow(data_dictionary))%>%
  add_row(pos = (nrow(data_dictionary)+1):length(colnames(data_cleaned)),
          variable = colnames(data_cleaned %>% 
                                select(contains("detect"),
                                       contains("imputed"))),
          label = c("Dichotomous PFDS (detected, non-detected)",
                    "Dichotomous PFPeS (detected, non-detected)",
                    "Dichotomous PFTrDA (detected, non-detected)",
                    "Imputed PFDA concentration (ng/mL), LOD=0.1 (0 was imputed as LOD/√2 )",
                    "Imputed PFHpA concentration (ng/mL), LOD=0.1 (0 was imputed as LOD/√2 )",
                    "Imputed PFHxS concentration (ng/mL), LOD=0.1 (0 was imputed as LOD/√2 )",
                    "Imputed PFHpS concentration (ng/mL), LOD=0.1 (0 was imputed as LOD/√2 )",
                    "Imputed PFNA concentration (ng/mL), LOD=0.12 (0 was imputed as LOD/√2 )",
                    "Imputed PFOA concentration (ng/mL), LOD=0.25 (0 was imputed as LOD/√2 )",
                    "Imputed PFOS concentration (ng/mL), LOD=0.25 (0 was imputed as LOD/√2 )",
                    "Imputed PFUdA concentration (ng/mL), LOD=0.13 (0 was imputed as LOD/√2 )"
                    ),
          col_type = c(rep("chr", 3), rep("dbl", 8))) %>%
  mutate(domain = case_when(pos %in% c(1:9) ~ "IDs",
                            pos %in% c(10:11) ~ "Outcome",
                            pos %in% c(12:29) ~ "Covariates",
                            TRUE ~ "Exposures"))
  


 data_dictionary1 %>% regulartable() %>% autofit() %>% 
  width(j=~pos, width=1) %>% 
  width(j=~variable,width=1) %>% 
  width(j=~label,width=3) %>%
  width(j=~col_type,width=1)
```
## Overall Data Summary for Covariates, Phenotypes, and Exposures {.tabset}

### Data Summary for covariates
```{r data summary covariates summarytools}
covariates <- data_dictionary1 %>% filter(domain == "Covariates")
summarytools::view(dfSummary(data_cleaned %>% dplyr::select(covariates$variable), style = 'grid', plain.ascii = FALSE, valid.col = FALSE, headings = FALSE), method = "render")
```

### Data Summary for Phenotype
```{r data summary Phenotypes summarytools}
outcome <- data_dictionary1 %>% filter(domain == "Outcome")
summarytools::view(dfSummary(data_cleaned %>% select(outcome$variable), style = 'grid', plain.ascii = FALSE, valid.col = FALSE, headings = FALSE), method = "render")
```

### Data Summary for Exposures
```{r data summary exposures summarytools}
exposures <- data_dictionary1 %>% filter(domain == "Exposures")
summarytools::view(dfSummary(data_cleaned %>% select(exposures$variable), style = 'grid', plain.ascii = FALSE, valid.col = FALSE, headings = FALSE), method = "render")
```



## Data Summary for Raw Exposure and Final Cleaned Exposure {.tabset}

Data processsing steps:

1) In raw data, PFAS of participants "HI4004364" ,"LA3120895" were missing. So in the final cleaned data, PFDA, PFHxS, PFNA, PFOA, PFOS PFUdA data of these two participants were from PILOT study
2) PFAS with percentage of non-detection more than 30% were changed to dichotomous variable with categories "detect" and "non-detect". In this study, PFPeS, PFDS, PFTrDa were changed to dichotomous variable for analysis part. 
3) PFAS with concentration of 0 were imputed by the LOD divided by the square root of 2.

### Data Summary for Raw Exposure (PFAS)
```{r raw Exposure Summary, results = 'asis', warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE}
data_sample <- data_raw %>% 
  filter(CLASS2 == "MEC Participant") %>%
  janitor::clean_names()

# subset PFAS data----
pfas_sample <- data_sample %>%
  select(contains("pfas_p")) 

# non-detect info----
## Percentage of non-detection(0)-----
pfas_pct0 <- pfas_sample %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>%
  dplyr::summarize(pct0  = mean(value == 0, na.rm = TRUE))

## Number of non-detected samples(0)----
pfas_num0<- pfas_sample %>% 
  summarise_each(funs(sum(.==0, na.rm = TRUE))) %>% t() %>% 
  data.frame() %>% rownames_to_column("name") %>% rename("numof0" = ".")

## complete non_detected info of pfas-----
pfas_nondetect <- pfas_pct0 %>% tidylog::left_join(pfas_num0)

## Summary of Raw PFAS data
pfas_stats_raw <- pfas_sample%>% 
  pivot_longer(everything()) %>%
  drop_na() %>%
  group_by(name) %>%
  dplyr::summarise(
    min = min(value),
    max = max(value),
    percentile_25 = jag2::qntle_fxn(value, 0.25),
    percentile_50 = jag2::qntle_fxn(value, 0.5),
    percentile_75 = jag2::qntle_fxn(value, 0.75),
    percentile_90 = jag2::qntle_fxn(value, 0.90)
  ) %>% 
  ungroup()

## Adding non detect information to stats table----
pfas_stats_raw<- pfas_stats_raw %>% tidylog::left_join(pfas_nondetect)

pfas_stats_raw %>% regulartable() %>% autofit() %>% 
  width(j=~name, width=1) %>% 
  width(j=~min,width=1) %>% 
  width(j=~max,width=1) %>%
  width(j=~percentile_25,width=1) %>%
  width(j=~percentile_50,width=1) %>%
  width(j=~percentile_75,width=1) %>%
  width(j=~percentile_90,width=1) %>%
  width(j=~pct0,width=1) %>%
  width(j = ~numof0, width = 1)
```

### Data Summary for Final Cleaned Exposure (PFAS)
```{r final cleaned Exposure Summary, results = 'asis'}
data_pfas<- data_cleaned %>% select(contains("imputed"))

## Stats----
pfas_stats <- data_pfas %>% 
  labelled::remove_labels()%>%
  pivot_longer(cols = pfas_pfda_imputed:pfas_pfuda_imputed) %>%
  drop_na() %>%
  group_by(name) %>%
  dplyr::summarise(
    geometric_mean = jag2::fungm(value),
    sd = jag2::fungsd(value),
    min = min(value),
    max = max(value),
    percentile_25 = jag2::qntle_fxn(value, 0.25),
    percentile_50 = jag2::qntle_fxn(value, 0.5),
    percentile_75 = jag2::qntle_fxn(value, 0.75),
    percentile_90 = jag2::qntle_fxn(value, 0.90)
  ) %>% 
  ungroup()


pfas_stats %>% regulartable() %>% autofit() %>% 
  width(j=~name, width=1) %>%
  width(j = ~geometric_mean, width = 1) %>%
  width(j = ~sd, width = 1) %>% 
  width(j=~min,width=1) %>% 
  width(j=~max,width=1) %>%
  width(j=~percentile_25,width=1) %>%
  width(j=~percentile_50,width=1) %>%
  width(j=~percentile_75,width=1) %>%
  width(j=~percentile_90,width=1) 
```

## Data Summary for Covariates split by case type
```{r covariate summary, results='asis'}
table1::table1(~ blood_age + year_between +
                 sex + eth + studyarea + bmi_cat + alcohol_intake_cat + smokestatus + diabetes| casetype,
               data = data_cleaned, 
               overall = FALSE)
```

## Data Summary for Phenotype case-control status split by case type
```{r phenotype summary,attr.output='style="max-height: 100px;"',}
table1::table1(~ status| casetype,
               data = data_cleaned, 
               overall = FALSE)
```

## Data Summary for Untargeted and Targeted Metabolites Features{.tabset}
**Data Preprocessing steps for Metabolites Features:**

* Feature filtering
  + Filtering out MEC study samples only
  + Filtering out features based on missingness < 50%
* Imputation
  + Replace 0's with lowest detected/sqrt(2)
* Log2 transformation

### Summary of Untargeted Metabolites Features
```{r untargeted metabolites features,results = 'asis'}
untargeted_meta <- readRDS(
        fs::path(
          dir_cleaned_data,
          "feature_metadata_untargeted.RDS")) 
# table(untargeted_meta$mode)
# table(untargeted_meta %>% drop_na(metabolites_name) %>% select(mode))

summary_untargeted <- data.frame(mode = c("c18pos", "c18neg"),
                                 number_of_metabolomic_features_raw = c(16031,
                                                               9823),
                                 number_of_metabolomic_features_cleaned =
                                   c(table(untargeted_meta$mode)[2],
                                     table(untargeted_meta$mode)[1]),
                                 number_of_metabolomic_features_annotated = c(
                                   table(untargeted_meta %>%
                                           drop_na(metabolites_name) %>%
                                           select(mode))[2],
                                   table(untargeted_meta %>%
                                           drop_na(metabolites_name) %>%
                                           select(mode))[1]
                                 ))

summary_untargeted %>% regulartable() %>% autofit() %>% 
  width(j=~mode, width=1) %>% 
  width(j=~number_of_metabolomic_features_raw,width=1) %>% 
  width(j=~number_of_metabolomic_features_cleaned,width=1) %>%
  width(j=~number_of_metabolomic_features_annotated,width=1)
```

### Summary of Targeted Metabolites Features
```{r targeted metabolites features,results = 'asis'}
targeted_meta <- readRDS(
        fs::path(
          dir_cleaned_data,
          "feature_metadata_targeted.RDS")) 


summary_targeted <- data.frame(mode = c("c18pos", "c18neg"),
                                 number_of_metabolomics_features_raw = c(157,
                                                               153),
                                 number_of_metabolomics_features_cleaned =
                                   c(table(targeted_meta %>% select(mode, name) %>% distinct() %>% select(mode))[2],
                                     table(targeted_meta %>% select(mode, name) %>% distinct() %>% select(mode))[1])
                                 )
summary_targeted %>% regulartable() %>% autofit() %>% 
  width(j=~mode, width=1) %>% 
  width(j=~number_of_metabolomics_features_raw,width=1) %>% 
  width(j=~number_of_metabolomics_features_cleaned,width=1)

```

