---
title: 'MEC analysis'
subtitle: "Single Exposure Analysis- HCC"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("!libraries.R")))
source(fs::path(here::here("!directories.R")))
```

```{r load data prs}
data <- readRDS(fs::path(
          dir_cleaned_data,
          "p27_hhear_pfas_covar_diet_prs_V1.RDS"))

prs_name <- colnames(data %>% select(contains("prs")))
```

## Data preparation

1) HCC data: 240 cases, 240 controls
2) NAFLD-related  HCC data : 97 cases, 97 controls
3) other  HCC data : 143 cases, 143 controls
```{r hcc data prs}
sample_id1 <- data %>% filter(status == "HCC") %>% select(setnum)
sample_id2 <- data %>% filter(casetype == "NAFLD-related HCC") %>% select(setnum)
sample_id3 <- data %>% filter(casetype == "Other HCC") %>% select(setnum)
data_analysis1 <- data %>% filter(setnum %in% sample_id1$setnum)
data_analysis2 <- data %>% filter(setnum %in% sample_id2$setnum)
data_analysis3 <- data %>% filter(setnum %in% sample_id3$setnum)

data_analysis_hcc <- data_analysis1 %>% 
  # mutate_at(.vars = pfas_name[1:8],
  #           .funs = list(scld = ~scale(.),
  #                        lg2 = ~log2(.), 
  #                        median = ~ifelse(. < quantile(., 0.5, na.rm = TRUE), 0, 1),
  #                        `85pct` = ~ifelse(. < quantile(., 0.85, na.rm = TRUE), 0, 1))) %>%
  mutate(status = ifelse(status== "Control", 0, 1)) %>%
  mutate_at(.vars = c("pfas_pfds_detect", "pfas_pfpes_detect", "pfas_pftrda_detect"),
            .funs = ~factor(., levels = c("non_detected", "detected")))

data_analysis_nafld_hcc <- data_analysis2 %>% 
  # mutate_at(.vars = pfas_name[1:8],
  #           .funs = list(scld = ~scale(.),
  #                        lg2 = ~log2(.), 
  #                        median = ~ifelse(. < quantile(., 0.5, na.rm = TRUE), 0, 1),
  #                        `85pct` = ~ifelse(. < quantile(., 0.85, na.rm = TRUE), 0, 1))) %>%
  mutate(status = ifelse(status== "Control", 0, 1)) %>%
  mutate_at(.vars = c("pfas_pfds_detect", "pfas_pfpes_detect", "pfas_pftrda_detect"),
            .funs = ~factor(., levels = c("non_detected", "detected")))

data_analysis_other_hcc <- data_analysis3 %>% 
  # mutate_at(.vars = pfas_name[1:8],
  #           .funs = list(scld = ~scale(.),
  #                        lg2 = ~log2(.), 
  #                        median = ~ifelse(. < quantile(., 0.5, na.rm = TRUE), 0, 1),
  #                        `85pct` = ~ifelse(. < quantile(., 0.85, na.rm = TRUE), 0, 1))) %>%
  mutate(status = ifelse(status== "Control", 0, 1)) %>%
  mutate_at(.vars = c("pfas_pfds_detect", "pfas_pfpes_detect", "pfas_pftrda_detect"),
            .funs = ~factor(., levels = c("non_detected", "detected")))
```

## Data summary {.tabset}
### Data summary for overall HCC data
```{r pfas summary for hcc data only prs, results = 'asis'}
table1::table1(~.|status, data_analysis_hcc %>% select(status, prs_name, bmi_cat, alcohol_intake_cat,smokestatus, diabetes))
```

### Data summary for NAFLD related HCC data
```{r pfas summary for NAFLD related HCC data prs}
table1::table1(~.|status, data_analysis_nafld_hcc %>% select(status, prs_name, bmi_cat, alcohol_intake_cat,smokestatus, diabetes))

```

### Data summary for other HCC data
```{r pfas summary for other HCC prs}
table1::table1(~.|status, data_analysis_other_hcc %>% select(status, prs_name, bmi_cat, alcohol_intake_cat,smokestatus, diabetes))
```

## Missing data checking
```{r missing control data included}
missing <- data_analysis_hcc %>% filter((is.na(prs_uw) & status == 0))

match_info <- data_analysis_hcc %>% filter(setnum %in% missing$setnum & status == 1)

## match control for setnum 2064-----
matched_control1 <- data %>% filter(studyarea == match_info[1,]$studyarea &
                             sex == match_info[1,]$sex &
                             eth == match_info[1,]$eth &
                             blood_age == match_info[1,]$blood_age &
                             q1_byr == match_info[1,]$q1_byr &
                             status == "Control") %>% 
  labelled::remove_labels(status) %>%
              mutate(status = 0)

## checking if setnum in hcc data
(matched_control1$setnum[2] %in% data_analysis_hcc$setnum)
### setnum 2056 was not in hcc data. Replace the missing control of 2064 with control 2056

data_analysis_hcc1 <- data_analysis_hcc %>% filter(!(status == 0 & setnum == 2064)) %>% labelled::remove_labels(setnum) %>%
  mutate(setnum = as.numeric(setnum))%>% 
  bind_rows(data %>% 
              filter(status == "Control" & setnum == matched_control1$setnum[2]) %>%  
              labelled::remove_labels(status) %>%
              mutate(setnum = 2064,
                     status = 0)
              )

data_analysis_other_hcc1 <- data_analysis_other_hcc %>% filter(!(status == 0 & setnum == 2064)) %>% labelled::remove_labels(setnum) %>%
  mutate(setnum = as.numeric(setnum))%>% 
  bind_rows(data %>% 
              filter(status == "Control" & setnum == matched_control1$setnum[2]) %>%  
              labelled::remove_labels(status) %>%
              mutate(setnum = 2064,
                     status = 0)
              )

## match control for setnum 2283----
matched_control2 <- data %>% filter(studyarea == match_info[2,]$studyarea &
                             sex == match_info[2,]$sex &
                             eth == match_info[2,]$eth &
                             blood_age == match_info[2,]$blood_age &
                             q1_byr == match_info[2,]$q1_byr &
                             status == "Control") %>% 
  labelled::remove_labels(status) %>%
              mutate(status = 0)

## checking if setnum in hcc data
(matched_control2$setnum %in% data_analysis_hcc$setnum)
### setnum 2101 was in hcc data. So that control could not be used as the control for 2283. 


## match control for setnum 2159----
matched_control3 <- data %>% filter(studyarea == match_info[3,]$studyarea &
                             sex == match_info[3,]$sex &
                             eth == match_info[3,]$eth &
                             blood_age == match_info[3,]$blood_age &
                             q1_byr == match_info[3,]$q1_byr &
                             status == "Control") %>% 
  labelled::remove_labels(status) %>%
              mutate(status = 0)

## checking if setnum in hcc data
(matched_control3$setnum %in% data_analysis_hcc$setnum)
### setnum 2159 was in hcc data. So that control could not be used as the control for 2159. 

## match control for setnum 2188----
matched_control4 <- data %>% filter(studyarea == match_info[4,]$studyarea &
                             sex == match_info[4,]$sex &
                             eth == match_info[4,]$eth &
                             blood_age == match_info[4,]$blood_age &
                             q1_byr == match_info[4,]$q1_byr &
                             status == "Control") %>% 
  labelled::remove_labels(status) %>%
              mutate(status = 0)

## checking if setnum in hcc data
(matched_control4$setnum %in% data_analysis_hcc$setnum)
# no matching control

## match control for setnum 1035----
matched_control5 <- data %>% filter(studyarea == match_info[5,]$studyarea &
                             sex == match_info[5,]$sex &
                             eth == match_info[5,]$eth &
                             blood_age == match_info[5,]$blood_age &
                             q1_byr == match_info[5,]$q1_byr &
                             status == "Control") %>% 
  labelled::remove_labels(status) %>%
              mutate(status = 0)

## checking if setnum in hcc data
(matched_control5$setnum %in% data_analysis_hcc$setnum)
# no matching control

## match control for setnum 1899----
matched_control6 <- data %>% filter(studyarea == match_info[6,]$studyarea &
                             sex == match_info[6,]$sex &
                             eth == match_info[6,]$eth &
                             blood_age == match_info[6,]$blood_age &
                             q1_byr == match_info[6,]$q1_byr &
                             status == "Control") %>% 
  labelled::remove_labels(status) %>%
              mutate(status = 0)

## checking if setnum in hcc data
(matched_control6$setnum %in% data_analysis_hcc$setnum)
# no matching control


## match control for setnum 2193----
matched_control7 <- data %>% filter(studyarea == match_info[7,]$studyarea &
                             sex == match_info[7,]$sex &
                             eth == match_info[7,]$eth &
                             blood_age == match_info[7,]$blood_age &
                             q1_byr == match_info[7,]$q1_byr &
                             status == "Control") %>% 
  labelled::remove_labels(status) %>%
              mutate(status = 0)

## checking if setnum in hcc data
(matched_control7$setnum %in% data_analysis_hcc$setnum)
# 2283 was already in the hcc data as control for 2283 we could not use it as control of 2193


##NOTE: after checking these 7 missing control, only one of them can find alternative. For the following analysis, I excluded all the participants either with missing case or missing control. 
data_analysis_hcc1 <- data_analysis_hcc %>% drop_na(prs_name)

paired_setnum <- data_analysis_hcc1 %>% 
  group_by(setnum) %>% 
  summarise(count = n()) %>% 
  ungroup()%>%
  filter(count == 2)

data_analysis_hcc2 <- data_analysis_hcc1 %>% 
  filter( setnum %in% paired_setnum$setnum)



data_analysis_nafld_hcc1 <- data_analysis_nafld_hcc %>% 
  filter( setnum %in% paired_setnum$setnum)

data_analysis_other_hcc1 <- data_analysis_other_hcc %>%
  filter(setnum %in% paired_setnum$setnum)
```


## Analysis1. Conditional logistic regression
**Outcome: HCC vs control**

**Exposure: normalized PFAS, lg2 transformed PFAS, pfas_median(above median, below median), pfas 85pct(above 85%, below 85%))**

**Covars: bmi, smoking status, alcohol intake, diabetes**

**Analysis: conditional logistic regression**

**Data Info: 231 cases, 231 controls**
```{r analysis1 prs}
## conditional logistic regression, no covars----------------------
# create dataframe with names of dependent vars

ind_vars <- tibble(exposure = colnames(data_analysis_hcc2 %>% select(contains("prs"))))

# ind_vars <- tibble(exposure = colnames(data_analysis_hcc1 %>% select(contains("scld"), contains("detect"), contains("lg2"), contains("median"), contains("85pct"))))

ind_vars$clogit_out_no_cov <- map(
  ind_vars$exposure, 
  ~clogit(as.formula(paste("status ~", ., "+strata(setnum)")),
          data = data_analysis_hcc2,
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
           odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "No covars")) 

## Conditional logistic regression, controlling for covariates ---------
ind_vars$clogit_out_cov <- map(
  ind_vars$exposure, 
  ~clogit(as.formula(paste("status ~", ., "+bmi+ alcohol_intake + smokestatus + diabetes + 
                           # casetype + 
                           strata(setnum)")), 
          data = data_analysis_hcc2,
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    filter(!term %in% c("bmi", "alcohol_intake", "smokestatusCurrent", "smokestatusFormer", "diabetesYes"
                        # , "casetypeNAFLD-related HCC", "casetypeOther HCC"
                        )) %>% 
    mutate(
           odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "Adjusted by covars")) 

# Get final data -------------------
all_results_df_hcc <- ind_vars %>% 
  mutate(clogit_results_out = map2(clogit_out_no_cov, 
                                   clogit_out_cov, 
                                   ~bind_rows(.x, .y))) %>% 
  unnest(clogit_results_out) %>% 
  select(model_type,covariates, everything(), 
         -c(clogit_out_no_cov, 
            clogit_out_cov)) %>%
  mutate(exposure_type = case_when(
                                   str_detect(exposure, "prs") ~ "prs")
         )

all_results_df_hcc1 <- all_results_df_hcc %>%
  arrange(desc(covariates))

all_results_df_hcc1 %>% filter(grepl("prs", term)) %>% select(model_type, covariates, exposure_type,exposure,term, odds_ratio, exp_ci_low, exp_ci_high, p.value) %>% regulartable() %>% autofit() 

# writexl::write_xlsx(all_results_df_hcc, fs::path(dir_result, "1_single_exposure_analysis_result_hcc.xlsx"))
```

## Coefficient plot of analysis 1
```{r coefficient plot1 prs}
### Create metabolites Coefficient Plots
all_results_df_hcc1  %>% 
  mutate(term = term %>% str_replace("pfas_", "") %>% str_replace("_imputed", ""),
         term = factor(term, levels = rev(unique(term))),
         sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         pfas = str_split_fixed(term, "_", 2)[,1],
         covariates = factor(covariates, levels = c("No covars", "Adjusted by covars")))%>%
  ggplot(aes(x = term,y = odds_ratio, color = sig)) +
  geom_point(size = 1) +
  coord_flip()+
  geom_errorbar(aes(ymin = exp_ci_low,
                    ymax = exp_ci_high),
                width = 0) +
  geom_hline(yintercept = 1, linetype = 2, color = "grey") +
  ylab("Odds Ratio (95% CI)") +
  facet_grid( pfas ~ covariates, space = "free", scale = "free")+
  theme(panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 10),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.title.y=element_blank(),
        axis.line.y = element_line(color = "black"),
        legend.position = "right")+
  coord_flip() +
  scale_color_manual(values = c("grey","red"))

# ggsave(filename = fs::path(dir_figures,
#                            "Fig 1 Single Exposure Analysis HCC.jpeg"),
#        width = 8,
#        height = 5,
#        bg = "white")
```




## Analysis2. Conditional logistic regression (NAFLD related HCC)
**Outcome: NAFLD related HCC vs control**

**Exposure: prs**

**Analysis: conditional logistic regression**

**Data Info: 91 cases, 91 controls**
```{r analysis2 prs}
## conditional logistic regression, no covars----------------------
# create dataframe with names of dependent vars


ind_vars$clogit_out_no_cov <- map(
  ind_vars$exposure, 
  ~clogit(as.formula(paste("status ~", ., "+strata(setnum)")),
          data = data_analysis_nafld_hcc1,
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
           odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "No covars"
           ))

## Conditional logistic regression, controlling for bmi ---------
ind_vars$clogit_out_cov <- map(
  ind_vars$exposure, 
  ~clogit(as.formula(paste("status ~", ., "+bmi+ alcohol_intake + smokestatus + diabetes + strata(setnum)")), 
          data = data_analysis_nafld_hcc1,
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    filter(!term %in% c("bmi", "alcohol_intake", "smokestatusCurrent", "smokestatusFormer", "diabetesYes")) %>% 
    mutate(odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "Adjusted by covars"))

# Get final data -------------------
all_results_df_nafld_hcc <- ind_vars %>% 
  mutate(clogit_results_out = map2(clogit_out_no_cov, 
                                   clogit_out_cov, 
                                   ~bind_rows(.x, .y))) %>% 
  unnest(clogit_results_out) %>% 
  select(model_type,covariates, everything(), 
         -c(clogit_out_no_cov, 
            clogit_out_cov)) %>%
    mutate(exposure_type = case_when(str_detect(exposure,"prs") ~ "prs"))

all_results_df_nafld_hcc1 <- all_results_df_nafld_hcc %>% 
  arrange(desc(covariates))


all_results_df_nafld_hcc1 %>% filter(grepl("prs",term)) %>% select(model_type, covariates, exposure_type,exposure,term, odds_ratio, exp_ci_low, exp_ci_high, p.value) %>% regulartable() %>% autofit() 

# writexl::write_xlsx(all_results_df_nafld_hcc, fs::path(dir_result, "1_single_exposure_analysis_result_nafld_hcc.xlsx"))
```

## Coefficient plot of analysis 2
```{r coefficient plot2 prs}
### Create metabolites Coefficient Plots
all_results_df_nafld_hcc1  %>% 
  mutate(term = term%>% str_replace("pfas_", "") %>% str_replace("_imputed", ""),
         term = factor(term, levels = rev(unique(term))),
         sig = case_when(
                          p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         pfas = str_split_fixed(term, "_", 2)[,1],
         covariates = factor(covariates, levels = c("No covars", "Adjusted by covars")))%>%
  ggplot(aes(x = factor(term, levels = ),y = odds_ratio, color = sig)) +
  geom_point(size = 1) +
  coord_flip()+
  geom_errorbar(aes(ymin = exp_ci_low,
                    ymax = exp_ci_high),
                width = 0) +
  geom_hline(yintercept = 1, linetype = 2, color = "grey") +
  ylab("Odds Ratio (95% CI)") +
  facet_grid( pfas~covariates, space = "free", scale = "free")+
  theme(panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 10),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.title.y=element_blank(),
        axis.line.y = element_line(color = "black"),
        legend.position = "none")+
  coord_flip() +
  scale_color_manual(values = c("grey", "red"))

# ggsave(filename = fs::path(dir_figures,
#                            "Fig 1 Single Exposure Analysis NAFLD HCC.jpeg"),
#        width = 8,
#        height = 5,
#        bg = "white")
```


## Analysis3. Conditional logistic regression (Other HCC)
**Outcome: Other HCC vs control**

**Exposure: prs**

**Analysis: conditional logistic regression**

**Data Info: 143 cases, 143 controls**
```{r analysis3 prs}
## conditional logistic regression, no covars----------------------
# create dataframe with names of dependent vars

ind_vars$clogit_out_no_cov <- map(
  ind_vars$exposure, 
  ~clogit(as.formula(paste("status ~", ., "+strata(setnum)")),
          data = data_analysis_other_hcc1,
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
           odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "No covars")) 

## Conditional logistic regression, controlling for bmi ---------
ind_vars$clogit_out_cov <- map(
  ind_vars$exposure, 
  ~clogit(as.formula(paste("status ~", ., "+bmi+ alcohol_intake + smokestatus + diabetes + strata(setnum)")), 
          data = data_analysis_other_hcc1,
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    filter(!term %in% c("bmi", "alcohol_intake", "smokestatusCurrent", "smokestatusFormer", "diabetesYes")) %>% 
    mutate(odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "Adjusted by covars")) 

# Get final data -------------------
all_results_df_other_hcc <- ind_vars %>% 
  mutate(clogit_results_out = map2(clogit_out_no_cov, 
                                   clogit_out_cov, 
                                   ~bind_rows(.x, .y))) %>% 
  unnest(clogit_results_out) %>% 
  select(model_type,covariates, everything(), 
         -c(clogit_out_no_cov, 
            clogit_out_cov))%>%
    mutate(exposure_type = case_when(str_detect(exposure,"prs") ~ "prs"))

all_results_df_other_hcc1 <- all_results_df_other_hcc %>% 
  arrange(desc(covariates))

all_results_df_other_hcc1 %>% filter(grepl("prs",term)) %>% select(model_type, covariates, exposure_type,exposure, term, odds_ratio, exp_ci_low, exp_ci_high, p.value) %>% regulartable() %>% autofit() 

# writexl::write_xlsx(all_results_df_other_hcc, fs::path(dir_result, "1_single_exposure_analysis_result_other_hcc.xlsx"))
```

## Coefficient plot of analysis 3
```{r coefficient plot3 prs}
### Create metabolites Coefficient Plots
all_results_df_other_hcc1  %>% 
  mutate(term = term %>% str_replace("pfas_", "") %>% str_replace("_imputed", ""),
         term = factor(term, levels = rev(unique(term))),
         sig = case_when(
                          p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         pfas = str_split_fixed(term, "_", 2)[,1],
         covariates = factor(covariates, levels = c("No covars", "Adjusted by covars")))%>%
  ggplot(aes(x = term,y = odds_ratio, color = sig)) +
  geom_point(size = 1) +
  coord_flip()+
  geom_errorbar(aes(ymin = exp_ci_low,
                    ymax = exp_ci_high),
                width = 0) +
  geom_hline(yintercept = 1, linetype = 2, color = "grey") +
  ylab("Odds Ratio (95% CI)") +
  facet_grid( pfas~covariates, space = "free", scale = "free")+
  theme(panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 10),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.title.y=element_blank(),
        axis.line.y = element_line(color = "black"),
        legend.position = "none")+
  coord_flip() +
  scale_color_manual(values = c("grey", "red", "purple"))

# ggsave(filename = fs::path(dir_figures,
#                            "Fig 1 Single Exposure Analysis other HCC.jpeg"),
#        width = 8,
#        height = 5,
#        bg = "white")
```


## Analysis4. Comparison of CLR with spline and CLR
**Outcome: HCC vs control**

**Exposure: prs**

**Analysis: conditional logistic regression with or without spline term**

**Data Info: 231 cases, 231 controls**
```{r Comparison of CLR with spline and CLR}
# create dataframe with names of dependent vars
# ind_vars1 <- tibble(exposure = colnames(data_analysis1 %>% select(contains("scld"), contains("detect"), contains("lg2"), contains("median"), contains("85pct"))))

ind_vars1 <- tibble(exposure = colnames(data_analysis_hcc1 %>% select(prs_uw, prs_wt)))

## Without covs ----------------------
ind_vars1 <- ind_vars1 %>% 
   mutate(clogit_mod_full = map(exposure, 
                                ~clogit(
                                  as.formula(
                                    paste0("status~pspline(",.,
                                           ",df=2,degree=1)+
                                               strata(setnum)")),
                                  data = data_analysis_hcc2, 
                                  method = "efron"
                                )), 
          clogit_mod_null = map(exposure, 
                                ~clogit(
                                  as.formula(paste0("status~",.,"+strata(setnum)")),
                                  data = data_analysis_hcc2, 
                                  method = "efron")), 
          anova = map2(clogit_mod_full, clogit_mod_null, 
                       ~anova(.x, .y)))

ind_vars1$anova[[1]] # model comparison result: pfoa
anova(ind_vars1$clogit_mod_full[[1]]) 
anova(ind_vars1$clogit_mod_full[[2]]) 


ind_vars1$clogit_out_no_cov <- map(
  ind_vars1$clogit_mod_full, 
  ~termplot(., term=1, se=TRUE, plot = FALSE) %>%
    as.data.frame() %>% 
    janitor::clean_names() %>%
    rename_all(~case_when(str_detect(., "_x") ~ "x", 
                          str_detect(., "_y") ~ "y",
                          str_detect(., "_se") ~ "se")) %>%
    as_tibble())

# Get final data 
clogit_results_df_no_cov <- ind_vars1 %>%
  unnest(clogit_out_no_cov) %>%
  mutate(est_low = y-(1.96*se),
         est_high = y+(1.96*se)) %>%
  mutate(covariates = "none")

# Add in percentile information -------------------------
# Calculate 90th percentile
(percentile_85 <- data_analysis_hcc2 %>% 
   select(prs_uw,prs_wt) %>% 
   pivot_longer(cols = everything()) %>% 
   group_by(name) %>% 
   summarise(pct_85 = quantile(value, 0.85, na.rm = TRUE)))

#Merge percentiles with data
clogit_results_df_no_cov_final <- clogit_results_df_no_cov %>% 
  tidylog::left_join(percentile_85, by = c("exposure" = "name")) %>% 
  mutate(
    # exposure = fct_relevel(exposure, "pfos", "pfhxs", "pfoa",
    #                             "pfna", "pfuda", "pfda"), 
         sig = if_else(est_low > 0 | est_high < 0, 
                       "sig", 
                       "not_sig")) 
```

## Spline plot-Fixed y axis
```{r spline plot}
## Plot -----------------------------------
(spline_fig <- ggplot(clogit_results_df_no_cov_final,
                      aes(x = x,
                          y = y)) + 
    geom_hline(yintercept = 0, color = "grey60", linetype = 3) + 
    geom_vline(aes(xintercept = pct_85), color = "red", linetype = 3) +
    geom_line() + 
    geom_line(aes(y = est_high), 
              linetype = 2) +
    geom_line(aes(y = est_low), 
              linetype = 2) +
    theme(plot.margin = unit(c(0,0,0,0), "cm")) +
    xlab("PFAS Concentration") + 
    ylab("Log odds") +
    facet_wrap(~exposure, scales = "free_x"))
```

## Spline plot with free scales
```{r spline plot 2}

(spline_fig <- ggplot(clogit_results_df_no_cov_final,
                      aes(x = x,
                          y = y)) + 
    geom_hline(yintercept = 0, color = "grey60", linetype = 3) + 
    geom_vline(aes(xintercept = pct_85), color = "red", linetype = 3) +
    geom_line() + 
    geom_line(aes(y = est_high), 
              linetype = 2) +
    geom_line(aes(y = est_low), 
              linetype = 2) +
    theme(plot.margin = unit(c(0,0,0,0), "cm")) +
    xlab("PFAS Concentration") + 
    ylab("Log odds") +
    facet_wrap(~exposure, scales = "free"))
```

## Spline plot with <85pct data
```{r spline plot 3}
(spline_fig <- ggplot(clogit_results_df_no_cov_final %>% filter(x < pct_85),
                      aes(x = x,
                          y = y)) + 
    geom_hline(yintercept = 0, color = "grey60", linetype = 3) + 
    geom_vline(aes(xintercept = pct_85), color = "red", linetype = 3) +
    geom_line() + 
    geom_line(aes(y = est_high), 
              linetype = 2) +
    geom_line(aes(y = est_low), 
              linetype = 2) +
    theme(plot.margin = unit(c(0,0,0,0), "cm")) +
    xlab("PFAS Concentration") + 
    ylab("Log odds") +
    facet_wrap(~exposure, scales = "free_x"))

# Save Figure
# ggsave(spline_fig, filename = fs::path(dir_figures,
#                                        "1_PFAS Splines fig.jpeg"),
#        height = 6, width = 8)
```

