---
title: 'MEC analysis'
subtitle: "Single Exposure Analysis- HCC"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r mixture analysis setup, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("!libraries.R")))
source(fs::path(here::here("!directories.R")))
source(fs::path(here::here("g_prior_sel_logistic_fun.R")))
```

```{r mixture analysis load data}
data <- readRDS(fs::path(
          dir_cleaned_data,
          "p27_hhear_pfas_covar_diet_prs_V1.RDS"))

pfas_name <- c("pfas_pfda_imputed","pfas_pfos_imputed","pfas_pfhpa_imputed",
               "pfas_pfhxs_imputed","pfas_pfhps_imputed","pfas_pfna_imputed",
               "pfas_pfoa_imputed","pfas_pfuda_imputed",
               "pfas_pfds_detect", "pfas_pfpes_detect",
               "pfas_pftrda_detect")
```


## Data preparation

1) HCC data: sample size 240
2) NAFLD-related  HCC data : sample size 97 
3) other  HCC data : sample size 143 

```{r 3 hcc data preparation}
sample_id1 <- data %>% filter(status == "HCC") %>% select(setnum)
sample_id2 <- data %>% filter(casetype == "NAFLD-related HCC") %>% select(setnum)
sample_id3 <- data %>% filter(casetype == "Other HCC") %>% select(setnum)

data1 <- data %>% 
  # mutate_at(.vars = pfas_name[1:8],
  #           .funs = list(lg2 = ~log2(.), 
  #                        median = ~ifelse(. < quantile(., 0.5, na.rm = TRUE), 0, 1),
  #                        `85pct` = ~ifelse(. < quantile(., 0.85, na.rm = TRUE), 0, 1))) %>%
  mutate(status = ifelse(status== "Control", 0, 1)) %>%
  mutate_at(.vars = c("pfas_pfds_detect", "pfas_pfpes_detect", "pfas_pftrda_detect"),
            .funs = ~factor(., levels = c("non_detected", "detected")))

data_analysis1 <- data1 %>% filter(setnum %in% sample_id1$setnum)
data_analysis2 <- data1 %>% filter(setnum %in% sample_id2$setnum)
data_analysis3 <- data1 %>% filter(setnum %in% sample_id3$setnum)
```

## Missing data checking
```{r missing control data included}
missing <- data_analysis1 %>% filter((is.na(pfas_pfda)|is.na(pfas_pfos)|is.na(pfas_pfhpa)|is.na(pfas_pfhxs)|is.na(pfas_pfhps)|is.na(pfas_pfna)|is.na(pfas_pfoa)|is.na(pfas_pfuda)|is.na(pfas_pfds)|is.na(pfas_pfpes)|is.na(pfas_pftrda)) & status == 0)

match_info <- data_analysis1 %>% filter(setnum == missing$setnum & status == 1)

matched_control <- data1 %>% filter(studyarea == match_info$studyarea &
                             sex == match_info$sex &
                             eth == match_info$eth &
                             blood_age == match_info$blood_age &
                             q1_byr == match_info$q1_byr &
                             status == 0) %>% 
  labelled::remove_labels(status)%>%
  mutate(status = 0)

## checking if setnum in hcc data
(matched_control$setnum[4] %in% data_analysis1$setnum)
### setnum 1696 was not in hcc data. Replace the missing control of 1297 with control 1696

data_analysis_hcc <- data_analysis1 %>% filter(!(status == 0 & setnum == 1297)) %>% labelled::remove_labels(setnum) %>%
  mutate(setnum = as.numeric(setnum))%>% 
  bind_rows(data1 %>% 
              filter(status == 0 & setnum == matched_control$setnum[4]) %>%  
              labelled::remove_labels(status) %>%
              mutate(setnum = 1297,
                     status = 0)
              )

data_analysis_other_hcc <- data_analysis3 %>% filter(!(status == 0 & setnum == 1297)) %>% labelled::remove_labels(setnum) %>%
  mutate(setnum = as.numeric(setnum))%>% 
  bind_rows(data1 %>% 
              filter(status == 0 & setnum == matched_control$setnum[4]) %>%  
              labelled::remove_labels(status) %>%
              mutate(setnum = 1297,
                     status = 0)
              )
```

## Mixture Analysis - qgcomp
```{r qgcomp}
# qgcomp
# Set covars and name of exposure

qgcompout <- qgcomp.noboot(as.formula(paste("status~(",
                                             str_c(pfas_name[1:8], collapse = "+"),
                                             ")+blood_age+sex+eth+studyarea")),
                            expnms=pfas_name[1:8], 
                            data = data_analysis_hcc, 
                            family=binomial(), 
                            q=4)
summary(qgcompout)
```

## Plot for qgcomp results(Weighted plot and coefficient plot)
```{r plot of qgcomp}
plot(qgcompout, cex = 1)

### coefficient plot----
coef <- data.frame(qgcompout$fit$coefficients) %>% 
  rownames_to_column("Predictor") %>%
  filter(Predictor %in% pfas_name[1:8]) %>%
  rename('Log Odds Ratio' = qgcompout.fit.coefficients) %>%
  mutate(direction = ifelse(`Log Odds Ratio` > 0, "pos", "neg"),
         direction = factor(direction, levels = c("pos", "neg")))

# change the name of the predictor
coef1 <- coef %>% mutate(Predictor = case_when(
  grepl("pfda", Predictor) ~ "PFDA",
  grepl("pfos", Predictor) ~ "PFOS",
  grepl("pfna", Predictor) ~ "PFNA",
  grepl("pfhxs", Predictor) ~"PFHxS",
  grepl("pfoa", Predictor) ~ "PFOA",
  grepl("pfhps", Predictor) ~ "PFHpS",
  grepl("pfhpa", Predictor) ~ "PFHpA",
  grepl("pfuda", Predictor) ~ "PFUdA"
))

p <- ggplot(coef1,aes(fct_reorder(Predictor, `Log Odds Ratio`), `Log Odds Ratio`,
                      fill = direction)) + 
  geom_bar(stat = "identity", show.legend = FALSE) + 
  ggtitle("Coefficient plot(Mixture Analysis)") +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = 2, color = "black") +
  geom_text(aes(label = round(`Log Odds Ratio`,2)), hjust= -0.05) +
  theme(axis.text = element_text(size = 15),
        text = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        panel.background = element_rect(fill="white"),
        strip.background = element_rect(fill = "white")
        )

p
```

## Mixture analysis - Weighted quantile sum regression(gwqs) 
```{r gwqs}
results <- gWQS::gwqs(status ~ wqs + blood_age + sex + eth + studyarea, 
                     mix_name = pfas_name[1:8], 
                     data = data_analysis_hcc, 
                     q = 4, validation =  0, b = 1000, b1_pos = TRUE, 
                     b1_constr = FALSE, family = "binomial", seed = 2016)

wqsr_results_unnested <- cbind(summary(results)$coefficients, 
                                    confint(results)) %>% 
                               as_tibble(rownames = "term")%>% 
  bind_rows(.id = "exposure_group") %>% 
  janitor::clean_names()

# Modify results, get effect estimates
(gwqs_results <- wqsr_results_unnested %>% 
  filter(term=="wqs") %>% 
  mutate(odds_ci = jag2::effest_ci(eff_est = exp(estimate), 
                             lcl = exp(x2_5_percent),
                             ucl = exp(x97_5_percent))) %>% 
  select(exposure_group, odds_ci, pr_z, everything()))
```

## Mixture analysis - BKMR
```{r bkmr}
set.seed(1000)
## Drop na
data_analysis_hcc_bkmr <- data_analysis_hcc %>% drop_na(pfas_name[1:7], prs_wt) #464

## filter paired setnums
paired_setnum <- data_analysis_hcc_bkmr%>%
  group_by(setnum) %>%
  summarise(count = n()) %>%
  ungroup()%>%
  filter(count == 2)
 
data_analysis_hcc_bkmr1 <- data_analysis_hcc_bkmr %>%
  filter(setnum %in% paired_setnum$setnum) #448

## log2 transform and scaled data
data_analysis_hcc_bkmr2 <- data_analysis_hcc_bkmr1 %>% 
  mutate_at(.vars = pfas_name[1:8],
            .funs = list(scld = ~scale(.),
                         lg2 = ~log2(.))) %>%
  mutate_at(.vars = c(paste0(pfas_name[1:8], "_lg2"), "prs_wt", "prs_uw"),
            .funs = list(scld = ~scale(.))) 
## rename columns
data_analysis_hcc_bkmr3 <- data_analysis_hcc_bkmr2 %>%
  rename_with(str_replace,
              pattern= "pfas_",
              replacement = "") %>%
  rename_with(str_replace,
              pattern= "imputed_",
              replacement = "")

# Analysis
mixture<- as.matrix(data_analysis_hcc_bkmr3 %>% 
                      select(contains("lg2_scld"), prs_wt_scld)) 
  

outcome<- data_analysis_hcc_bkmr3$status

covariates<- data_analysis_hcc_bkmr3 %>% 
  select(blood_age, sex, eth, studyarea) %>% 
  fastDummies::dummy_cols(select_columns = c("sex", "eth", "studyarea"),
                          remove_first_dummy = TRUE,
                          remove_selected_columns = TRUE)

fitkm <- kmbayes(y = outcome, 
                 Z = mixture,
                 X = covariates,  
                 iter = 10000, 
                 verbose = FALSE, 
                 varsel = TRUE)

ExtractPIPs(fitkm)
```

### Result: checking the convergence of the parameters
```{r check the convergence of parameters}
TracePlot(fit = fitkm, par = "beta")
TracePlot(fit = fitkm, par = "sigsq.eps")
TracePlot(fit = fitkm, par = "r", comp = 1)
```

### Univariate dose-response
**Relationship between each exposure and outcome where all the other exposures are fixed to a particular percentile (median)**

```{r unvivariate dose-response}
pred.resp.univar <- PredictorResponseUnivar(fit=fitkm,
                                            method="approx")
                                            
ggplot(pred.resp.univar, aes(z, est, ymin = est - 1.96*se, 
                             ymax = est + 1.96*se)) + 
  geom_smooth(stat = "identity") + ylab("h(z)") + facet_wrap(~ variable, scales = "free")
```

### Bivariable Exposure-response 
```{r bivariable exposure-response}
pred.resp.bivar  <- PredictorResponseBivar(fit=fitkm,
                                           min.plot.dist = 1,
                                           method="approx")
                                           
ggplot(pred.resp.bivar, aes(z1, z2, fill = est)) + 
  geom_raster() + 
  facet_grid(variable2 ~ variable1, scales = "free") +
  scale_fill_gradientn(colours=c("#0000FFFF","#FFFFFFFF","#FF0000FF")) +
  xlab("expos1") +
  ylab("expos2") +
  ggtitle("h(expos1, expos2)")
```


### Interactions
**investigate the predictor-response function of a single predictor for the second predictor fixed at various quantiles. In this study 25th, 50th and 75th were chosen**

**Interpretation: when a change in the dose-response curve of one component at a different quartile of another component, it indicates the presence of interaction while if the dose-response curves are parallel, it suggests no interaction between the two components in the mixtures **

```{r interaction}
pred.resp.bivar.levels <-
  PredictorResponseBivarLevels(pred.resp.df= 
                                 pred.resp.bivar,
                               Z = mixture, 
                               both_pairs=TRUE,
                               qs = c(0.25, 0.5, 0.75))
                               
ggplot(pred.resp.bivar.levels, aes(z1, est)) + 
  geom_smooth(aes(col = quantile), stat = "identity") + 
 facet_grid(variable2 ~ variable1,scales = "free") +
  ggtitle("h(expos1 | quantiles of expos2)") +
 xlab("expos1")

ggplot(pred.resp.bivar.levels %>% filter(variable1 == "pfhpa_lg2_scld"& variable2 == "prs_wt_scld"), aes(z1, est))+ 
  geom_smooth(aes(col = quantile), stat = "identity") + 
 facet_grid(variable2 ~ variable1,scales = "free") +
  ggtitle("h(expos1 | quantiles of expos2)") +
 xlab("expos1")
```

### overall mixture effect
**Compare the expected change in the outcome when the component in the mixture were set at each 5th percentile from the 25th to 75th**

```{r overall mixture effect}
risks.overall <- OverallRiskSummaries(fit=fitkm, 
                                      qs=seq(0.25, 0.75, by=0.05), 
                                      q.fixed = 0.5, 
                                      method = "approx")
                                      
ggplot(risks.overall, aes(quantile, est, ymin = est - 1.96*sd, 
                          ymax = est + 1.96*sd)) +  
  geom_hline(yintercept=00, linetype="dashed", color="gray") + 
  geom_pointrange() + scale_y_continuous(name="estimate") 
```


### single varibles effects
```{r single variables effects}
risks.singvar <- SingVarRiskSummaries(fit=fitkm, 
                                      qs.diff = c(0.25, 0.75),
                                      q.fixed = c(0.25, 0.50, 0.75), 
                                      method = "approx")
                                      
ggplot(risks.singvar, aes(variable, est, ymin = est - 1.96*sd,  
                          ymax = est + 1.96*sd, col = q.fixed)) + 
       geom_hline(aes(yintercept=0), linetype="dashed", color="gray") +  
       geom_pointrange(position = position_dodge(width = 0.75)) +
       coord_flip() + theme(legend.position="none")+scale_x_discrete(name="") +
       scale_y_continuous(name="estimate") 
```

### single variable interaction terms
```{r single variable interaction terms}
risks.int <- SingVarIntSummaries(fit=fitkm, 
                                 qs.diff = c(0.25, 0.75),
                                 qs.fixed = c(0.25, 0.75))
                                 
ggplot(risks.int, aes(variable, est, ymin = est - 1.96*sd, 
                      ymax = est + 1.96*sd)) + 
  geom_pointrange(position = position_dodge(width = 0.75)) + 
  geom_hline(yintercept = 0, lty = 2, col = "brown") + coord_flip()
```

## Mixture analysis - gprior
```{r gprior}
# # Set up data ------------------------------
# # Create dummy vars for ethnicity
# data_analysis_hcc <- fastDummies::dummy_cols(data_analysis_hcc, select_columns = c("sex", "eth", "studyarea"))
# 
# # Set covars and name of exposure
# covars <- c("blood_age", "sex_Female","studyarea_California",
#             "eth_African American", "eth_Japanese American", "eth_Latino", "eth_Native Hawaiian")
# 
# # Set datasets for analysis
# # X observations
# X <- data_analysis_hcc %>%  
#   dplyr::select(all_of(
# pfas_name_scld)) 
# 
# # Outcome
# Y = data_analysis_hcc$status
# 
# # Covariates (numeric only)
# U = data_analysis_hcc %>% select(all_of(covars))
# 
# SETNUM = data_analysis_hcc$setnum
# 
# # Call Function -----------------------------------
# gprior_output <- g.prior.sel.logistic(X = X, Y=Y, U=U, SETNUM = SETNUM)
# 
# # mutate output
# gprior_results_df <- bind_rows(gprior_output, 
#                                .id = "exposures") %>% 
#   janitor::clean_names() %>%
#   mutate(odds_ratio = exp(mean), 
#          odds_ratio_lcl = exp(x2_5),
#          odds_ratio_ucl = exp(x97_5)) %>% 
#   rename(log_odds = mean, 
#          conf_low  = x2_5,
#          conf_high = x97_5) %>% 
#   remove_rownames() %>% 
#   mutate(odds_ci = effest_ci(eff_est = odds_ratio,
#                              odds_ratio_lcl,
#                              odds_ratio_ucl))
# 
# 
# 
# # Save RDS
# write_csv(gprior_results_df, 
#           fs::path(dir_results, 
#                    "g prior with selection legacy PFAS mixture.csv"))
```

