---
title: 'MEC analysis'
subtitle: "Final Analysis"
author: "Hongxu Wang, Jesse Goodrich"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup univariate analysis cat pfas, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("1_Project_Setup/!libraries.R")))
source(fs::path(here::here("1_Project_Setup/!directories.R")))
source(fs::path(here::here("1_Project_Setup/!functions.R")))
source(fs::path(here::here("1_Project_Setup/!load_data.R")))

```

## Analysis1. Conditional logistic regression - HCC
**Outcome: HCC vs control**

**Exposure: normalzied log2 PFAS**

**Covars: smoking, diabetes, q1_edih, prs_wt, v1:v6**

**Analysis: conditional logistic regression**

**Data Info: 158 cases, 158 controls**


```{r analysis1}

data_hcc <- data_hcc %>% filter(sex == "Male")

res_hcc <- epiomics::owas_clogit(data_hcc , 
                                 cc_set = "setnum", 
                                 cc_status = "status", 
                                 covars = covars,
                                 omics = c(pfas_name,"pfas_t_PFBS_detected"),
                                 conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "Male")

# eth_categories <- unique(data_hcc$eth)
# 
# for(race in eth_categories[c(1:2)]) {
#   res_hcc_temp <- epiomics::owas_clogit(data_hcc %>% filter(eth == race) , 
#                                         cc_set = "setnum", 
#                                         cc_status = "status", 
#                                         covars = covars,
#                                         omics = pfas_name,
#                                         conf_int = TRUE) %>%
#     mutate(odds_ratio = exp(estimate),
#            exp_ci_low = exp(conf_low),
#            exp_ci_high = exp(conf_high),
#            type = race)
#   res_hcc <- res_hcc %>% bind_rows(res_hcc_temp)
# }
# rm(res_hcc_temp)
# reorder PFAS by estimate
res_hcc1 <- res_hcc |> 
  mutate(pfas_name = str_remove(feature_name, "pfas_lg2_"),
         sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "FDR p value < 0.05",
                         p_value < 0.05 ~ "p value < 0.05", 
                         p_value >= 0.05 ~ "p value ≥ 0.05"),
         sig = factor(sig, levels = c("p value ≥ 0.05", "p value < 0.05", "FDR p value < 0.05"))) %>%
  mutate(covar = sub(".*~(.*?)\\+pfas.*", "\\1", formula))%>%
  select(covar, everything())
```

### Fig1a. association of pfas and hcc - HCC
```{r plot with covars only, fig.width=10, fig.height=5}
### Rename PFAS for plotting
res_hcc2 <- res_hcc1 %>%
  mutate(pfas_name = rename_pfas(pfas_name))%>%
  mutate(pfas_name = fct_reorder(pfas_name, if_else(type == "Male", estimate, NA_real_))) %>%
  mutate(type = case_when(type == "Male" ~ "Male \n(n = 316)"
                          # type == "Japanese American" ~ "Japanese American \n(n = 176)",
                          # type == "Latino"~ "Latino \n(n = 158)"
                          # type == "White" ~ "White \n(n = 46)",
                          # type == "African American" ~ "African American \n(n = 50)",
                          # type == "Native Hawaiian" ~ "Native Hawaiian \n(n = 34)"
                          ))

pfhpa_hcc_main <- res_hcc2  %>%
    # filter(!type  %in% c("White \n(n = 46)", "African American \n(n = 50)"))%>%
    ggplot(aes(x = pfas_name,y = exp(estimate), color = sig)) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = exp(conf_low),
                      ymax = exp(conf_high)),
                  width = 0) +
    geom_hline(yintercept = 1, linetype = 2, color = "grey") +
    ylab("Odds Ratio (95% CI)") +
    # ggtitle("Exposure Outcome Analysis for HCC Cases") +
    # facet_grid(.~ factor(type, levels = c("Overall \n(n = 4158)", "Japanese American \n(n = 182)", "Latino \n(n = 158)", "Others \n(n = 132)")), scales = "free_y")+
    # facet_grid(.~ factor(type, levels = c("Overall \n(n = 316)", "Japanese American \n(n = 176)", "Latino \n(n = 158)", "White \n(n = 46)", "African American \n(n = 50)", "Native Hawaiian \n(n = 34)")), scales = "free")+ 
    theme(panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_text(angle = 0, hjust = 0),
          # strip.text.x = element_blank(),
          text = element_text(size = 15),
          # axis.title.x=element_blank(),
          axis.line.x = element_line(color = "black"),
          axis.title.y=element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
    coord_flip() +
    scale_color_manual(values = c("grey","red","purple"))


# ggsave(filename = fs::path(dir_figures,
#                            "Fig 1a Single Exposure Analysis HCC with covars_stratified_by_eth_0603.jpeg"),
#        width = 10,
#        height = 5,
#        dpi = 300,
#        bg = "white")
# 
writexl::write_xlsx(res_hcc2, fs::path(dir_result, "T1.Continuous PFAS and HCC_Male.xlsx"))


# (pfhpa_hcc_main_slides <- res_hcc2  %>%
#     filter(type == "Overall \n(n = 316)")%>%
#     ggplot(aes(x = pfas_name,y = exp(estimate), color = sig)) +
#     geom_point(size = 1) +
#     geom_errorbar(aes(ymin = exp(conf_low),
#                       ymax = exp(conf_high)),
#                   width = 0) +
#     geom_hline(yintercept = 1, linetype = 2, color = "grey") +
#     ylab("Odds Ratio (95% CI)") +
#     # ggtitle("Exposure Outcome Analysis for HCC Cases") +
#     # facet_grid(.~ factor(type, levels = c("Overall \n(n = 4158)", "Japanese American \n(n = 182)", "Latino \n(n = 158)")), scales = "free_y")+
#     theme(panel.background = element_rect(fill="white"), 
#           strip.background = element_rect(fill = "white"),
#           strip.text.y = element_text(angle = 0, hjust = 0),
#           # strip.text.x = element_blank(),
#           text = element_text(size = 15),
#           # axis.title.x=element_blank(),
#           axis.line.x = element_line(color = "black"),
#           axis.title.y=element_blank(),
#           axis.line.y = element_line(color = "black"),
#           legend.position = "none",
#           legend.title = element_blank())+
#     coord_flip() +
#     scale_color_manual(values = c("black","red","purple")))


# ggsave(filename = fs::path(dir_figures,
#                            "main_figures",
#                            "Fig. PFAS main effect_Overall.jpeg"),
#        width = 7,
#        height = 5,
#        dpi = 300,
#        bg = "white")
```


## Analysis2. Conditional logistic regression - Categorized PFAS and HCC
**Outcome: HCC vs control**

**Exposure: categorized pfas(quartile)**

**Covars: smoking, diabetes, q1_edih, prs_wt, v1:v6**

**Analysis: conditional logistic regression**

**Data Info: 158 cases, 158 controls**


```{r analysis1 cat}

covariate_input <- paste(covars, collapse = " + ")

## create dataframe with names of dependent vars (PFAS) ----
pfas_names <- tibble(exposure = pfas_name_cat)

# Get Model covariates 
covar_names <- tibble(
  covariates =  covariate_input)

# Make all clogit formulas
ind_vars <- expand_grid(pfas_names,  covar_names) |>  
  mutate(formula = paste("status ~ ", 
                         exposure, 
                         "+", covariates, 
                         "+strata(setnum)") |>
           str_replace("\\+  \\+", " + "))

## Analysis for all related participants
ind_vars$clogit_out <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_hcc,
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

# ## Analysis for Japanese American 
# ind_vars$clogit_out1 <- map(
#   ind_vars$formula, 
#   ~clogit(as.formula(.),
#           data = data_hcc %>% filter(eth == "Japanese American"),
#           method = "efron",
#           robust = TRUE
#   ) %>%
#     tidy(., conf.int = TRUE, level = 0.95) %>% 
#     mutate(
#       odds_ratio = exp(estimate),
#       exp_ci_low = exp(conf.low),
#       exp_ci_high = exp(conf.high),
#       model_type = "clogit")) 

## Analysis for Latino
# ind_vars$clogit_out2 <- map(
#   ind_vars$formula, 
#   ~clogit(as.formula(.),
#           data = data_hcc %>% filter(eth == "Latino"),
#           method = "efron",
#           robust = TRUE
#   ) %>%
#     tidy(., conf.int = TRUE, level = 0.95) %>% 
#     mutate(
#       odds_ratio = exp(estimate),
#       exp_ci_low = exp(conf.low),
#       exp_ci_high = exp(conf.high),
#       model_type = "clogit")) 

## Analysis for Others
# ind_vars$clogit_out3 <- map(
#   ind_vars$formula, 
#   ~clogit(as.formula(.),
#           data = data_hcc %>% filter(eth_new == "Others"),
#           method = "efron",
#           robust = TRUE
#   ) %>%
#     tidy(., conf.int = TRUE, level = 0.95) %>% 
#     mutate(
#       odds_ratio = exp(estimate),
#       exp_ci_low = exp(conf.low),
#       exp_ci_high = exp(conf.high),
#       model_type = "clogit"))
# Get final data -------------------

intercept_data<- tibble(exposure = pfas_name_cat, 
                        term = paste0(pfas_name_cat, "1"), 
                        odds_ratio = rep(1,length(pfas_name_cat)),
                        effect_est  = rep("1",length(pfas_name_cat)))

all_results_df_hcc <- ind_vars %>% 
  unnest(clogit_out) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  # select(-clogit_out1, -clogit_out2) %>%
  filter(grepl("pfas", term)) %>%
  bind_rows(intercept_data) %>% 
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type = "Male")

# all_results_df_hcc_japan <- ind_vars %>% 
#   unnest(clogit_out1) %>% 
#   select(model_type, covariates, everything()) %>%
#   mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
#                          p.value >= 0.05 ~ "Not Sig"),
#          effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
#   select(-clogit_out, -clogit_out2) %>%
#   filter(grepl("pfas", term))%>%
#   bind_rows(intercept_data) %>%
#   mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
#          type= "Japanese American")
# 
# all_results_df_hcc_latino <- ind_vars %>% 
#   unnest(clogit_out2) %>% 
#   select(model_type, covariates, everything()) %>%
#   mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
#                          p.value >= 0.05 ~ "Not Sig"),
#          effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
#   select(-clogit_out, -clogit_out1) %>%
#   filter(grepl("pfas", term))%>%
#   bind_rows(intercept_data) %>%
#   mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
#          type = "Latino")
# 
# all_results_df_hcc_others <- ind_vars %>% 
#   unnest(clogit_out3) %>% 
#   select(model_type, covariates, everything()) %>%
#   mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
#                          p.value >= 0.05 ~ "Not Sig"),
#          effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
#   select(-clogit_out, -clogit_out1) %>%
#   filter(grepl("pfas", term))%>%
#   bind_rows(intercept_data) %>%
#   mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
#          type = "Others")


all_result_df <- all_results_df_hcc %>% 
  # bind_rows(all_results_df_hcc_japan) %>%
  # bind_rows(all_results_df_hcc_latino) %>%
  # bind_rows(all_results_df_hcc_others) %>%
  select(exposure, term, odds_ratio, exp_ci_low, exp_ci_high,effect_est,p.value,p_adjusted, type) %>%
  arrange(exposure)

```

```{r Dose response}
data_hcc1 <- data_hcc %>% mutate_at(.vars = pfas_name_cat,
                                    .funs = ~as.numeric(.))
res_hcc <- epiomics::owas_clogit(data_hcc1 , 
                                 cc_set = "setnum", 
                                 cc_status = "status", 
                                 covars = covars,
                                 omics = pfas_name_cat,
                                 conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "Male")

# eth_categories <- unique(data_hcc1$eth_new)
# 
# for(race in eth_categories) {
#   res_hcc_temp <- epiomics::owas_clogit(data_hcc1 %>% filter(eth_new == race) , 
#                                         cc_set = "setnum", 
#                                         cc_status = "status", 
#                                         covars = covars,
#                                         omics = pfas_name_cat,
#                                         conf_int = TRUE) %>%
#     mutate(odds_ratio = exp(estimate),
#            exp_ci_low = exp(conf_low),
#            exp_ci_high = exp(conf_high),
#            type = race)
#   res_hcc <- res_hcc %>% bind_rows(res_hcc_temp)
# }

# reorder PFAS by estimate
res_hcc1 <- res_hcc |> 
  mutate(sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "Sig.(FDR < 0.05)",
                         p_value < 0.05 ~ "Sig.(p value < 0.05)", 
                         p_value >= 0.05 ~ "Not Sig.")) 

p_trend_df <- res_hcc1 %>% select(feature_name, p_value, adjusted_pval,type) %>%
  rename(exposure = feature_name,
         p_trend = p_value)

all_result_df1 <- all_result_df %>% 
  tidylog::left_join(p_trend_df, by = c("exposure", "type")) %>%
  mutate(pfas_name = str_remove(exposure, "pfas_lg2_") %>%
           str_remove("_quartile")) %>%
  select(type,pfas_name, term,everything(), -exposure)

## N value for each quartile split by status
# Assuming df is your dataframe

# Convert to long format and create a pair identifier
data_list <- list(data1 = data_hcc %>% mutate(type = "Male"))

data_n <- data.frame()

for(data in data_list){
  type = unique(data$type)
  long_data_hcc <- data %>%
    select(contains("quartile"), status) %>%
    mutate(status = ifelse(status == 0, "Control", "Case")) %>%
    pivot_longer(cols = c(contains("quartile")), names_to = "variable",
                 values_to = "value") %>%
    drop_na() %>%
    mutate(pair = paste0(variable, value)) %>%
    select(-variable,-value) %>%
    group_by(pair, status) %>%
    summarise(count = n(), .groups = "drop")
  # Now, to spread back for a wide format summary if needed
  wide_summary_hcc <- pivot_wider(long_data_hcc, names_from = status,
                                  values_from = count, values_fill = list(count = 0))%>%
    mutate(type = type) %>%
    rename(term = pair)
  data_n <- data_n %>% bind_rows(wide_summary_hcc)
}

all_result_df2 <- all_result_df1 %>% 
  tidylog::left_join(data_n, by = c("term","type")) %>%
  select(type:term, Control, Case, everything())%>%
  arrange(pfas_name, type, term)

```

### Fig 1b. Dose response plot for PFHpA and HCC
```{r dose response plot}
# Plot interaction effects
all_result_df3 <- all_result_df2 %>% 
  mutate(plot_name = substr(term, 
                            start = nchar(term) - 8,
                            stop = nchar(term)) %>% toTitleCase(),
         pfas_name = str_remove(pfas_name, "pfas_t_")) %>%
  mutate(
    plot_name = paste(substr(plot_name, 1, 8), substr(plot_name, 9, 9)),
    sig = case_when(p_adjusted < 0.05 ~ "FDR. Sig.", 
                    p.value < 0.05 ~ "Sig.",
                    TRUE ~ "Not Sig."),
    label = str_c("p trend = ", format(signif(p_trend, digits = 2), scientific = FALSE))) %>%
  mutate(type = case_when(type == "Male" ~ "Male \n(n = 316)")
                          # type == "Japanese American" ~ "Japanese American \n(n = 176)", 
                          # type == "Latino"~ "Latino \n(n = 158)",
                          # type == "Others" ~ "Others \n(n = 134)")
         )

writexl::write_xlsx(all_result_df3, fs::path(dir_result, "single_categorical_exposure_analysis_Male.xlsx"))

pfhpa_hcc <-  ggplot(all_result_df3 %>% 
                        filter(pfas_name == "PFHPA"),
                        # filter(type != "Others \n(n = 134)"),
                      aes(x = odds_ratio, y = fct_rev(plot_name))) +
    geom_errorbar(aes(xmin = exp_ci_low, xmax = exp_ci_high),
                  width = 0, position = position_dodge(width = 1)) +
    geom_point(size = 1, position = position_dodge(width = 1)) + 
    geom_vline(xintercept = 1, linetype = 2, color = "grey") +
    # facet_grid( .~ factor(type, levels = c("Overall \n(n = 4158)", "Japanese American \n(n = 182)", "Latino \n(n = 158)", "Others \n(n = 132)")))+
    # facet_grid(~ factor(type, levels = c("Overall \n(n = 316)", "Japanese American \n(n = 176)", "Latino \n(n = 158)")), scales = "free")+
    xlab("Odds Ratio (95% CI)") +
    theme(#axis.title.x = element_blank(),
      # axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), 
      legend.position = "bottom",
      panel.background = element_rect(fill="white"), 
      strip.background = element_rect(fill = "white"),
      strip.text.y = element_text(angle = 0, hjust = 0),
      # strip.text.x = element_blank(),
      text = element_text(size = 15),
      axis.title.y = element_blank(),
      axis.line.x = element_line(color = "black"),
      axis.line.y = element_line(color = "black")) 
# +
# ylab("OR (95% CI) per\ndoubling of PFAS") + 
# ggtitle("Interaction of PFAS and PRS") +
# scale_color_manual(values = c("purple", "black","red")))


p1 <- pfhpa_hcc + geom_text(
  data = all_result_df3 %>% 
    filter(pfas_name == "PFHPA"),
    # filter(type != "Others \n(n = 134)"),
  mapping = aes(x = -Inf, y = -Inf, label = label),
  color = "black",
  hjust   = -0.3,
  vjust   = -13,
  size = 4
)


# (fig <- cowplot::plot_grid(NULL, pfhpa_hcc_main, NULL, 
#                            p1,
#                            ncol = 1, 
#                            rel_heights = c(0.02, 0.2, 0.02, 0.12),
#                            labels = c("A. PFAS and HCC","", "B. Dose-response of PFHpA and HCC", ""),
#                           label_x = c(0.13, 0, 0.04, 0),
#                           label_size = 12,
#                           align = "v",
#                           axis = "lt"))
```


### Fig 1b. Dose response plot for PFNA and HCC
```{r dose response plot PFNA}

pfna_hcc <-  ggplot(all_result_df3 %>% 
                        filter(pfas_name == "PFNA") ,
                        # filter(type != "Others \n(n = 134)"),
                      aes(x = odds_ratio, y = fct_rev(plot_name))) +
    geom_errorbar(aes(xmin = exp_ci_low, xmax = exp_ci_high),
                  width = 0, position = position_dodge(width = 1)) +
    geom_point(size = 1, position = position_dodge(width = 1)) + 
    geom_vline(xintercept = 1, linetype = 2, color = "grey") +
    # facet_grid( .~ factor(type, levels = c("Overall \n(n = 4158)", "Japanese American \n(n = 182)", "Latino \n(n = 158)", "Others \n(n = 132)")))+
    # facet_grid(~ factor(type, levels = c("Overall \n(n = 316)", "Japanese American \n(n = 176)", "Latino \n(n = 158)")), scales = "free")+
    xlab("Odds Ratio (95% CI)") +
    theme(#axis.title.x = element_blank(),
      # axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), 
      legend.position = "bottom",
      panel.background = element_rect(fill="white"), 
      strip.background = element_rect(fill = "white"),
      strip.text.y = element_text(angle = 0, hjust = 0),
      # strip.text.x = element_blank(),
      text = element_text(size = 15),
      axis.title.y = element_blank(),
      axis.line.x = element_line(color = "black"),
      axis.line.y = element_line(color = "black"))
# +
# ylab("OR (95% CI) per\ndoubling of PFAS") + 
# ggtitle("Interaction of PFAS and PRS") +
# scale_color_manual(values = c("purple", "black","red")))


p2 <- pfna_hcc + geom_text(
  data = all_result_df3 %>% 
    filter(pfas_name == "PFNA"),
    # filter(type != "Others \n(n = 134)"),
  mapping = aes(x = -Inf, y = -Inf, label = label),
  color = "black",
  hjust   = -0.3,
  vjust   = -13,
  size = 4
)

# (fig <- cowplot::plot_grid(NULL, pfhpa_hcc_main, NULL, 
#                            p1,
#                            ncol = 1, 
#                            rel_heights = c(0.02, 0.2, 0.02, 0.12),
#                            labels = c("A. PFAS and HCC","", "B. Dose-response of PFHpA and HCC", ""),
#                           label_x = c(0.13, 0, 0.04, 0),
#                           label_size = 12,
#                           align = "v",
#                           axis = "lt"))
```


## Fig1. Association of single PFAS and HCC and Dose response analysis
```{r combine figures, fig.width=10, fig.height=8}
(fig <- cowplot::plot_grid(NULL, pfhpa_hcc_main, NULL,
                           p1,
                           ncol = 1,
                           rel_heights = c(0.01, 0.2, 0.01, 0.15),
                           labels = c("A. PFAS and HCC","", "B. Dose-response of PFHpA and HCC", ""),
                           label_x = c(-0.06, 0, -0.14, 0),
                           label_size = 12,
                           align = "v",
                           axis = "lt"))

# ggsave(filename = fs::path(dir_figures,
#                            # "main_figures",
#                            "Fig1. Main Effect Final_0521_PFHpA.jpeg"),
#        width = 10,
#        height = 8,
#        dpi = 300,
#        bg = "white")

(fig <- cowplot::plot_grid(NULL, pfhpa_hcc_main, NULL,
                           p2,
                           ncol = 1,
                           rel_heights = c(0.01, 0.2, 0.01, 0.15),
                           labels = c("A. PFAS and HCC","", "B. Dose-response of PFNA and HCC", ""),
                           label_x = c(-0.06, 0, -0.14, 0),
                           label_size = 12,
                           align = "v",
                           axis = "lt"))

# ggsave(filename = fs::path(dir_figures,
#                            # "main_figures",
#                            "Fig1. Main Effect Final_0521_PFNA.jpeg"),
#        width = 10,
#        height = 8,
#        dpi = 300,
#        bg = "white")

(fig <- cowplot::plot_grid(NULL, pfhpa_hcc_main, NULL,
                           p1, NULL, p2,
                           ncol = 1,
                           rel_heights = c(0.01, 0.2, 0.01, 0.15, 0.01, 0.15),
                           labels = c("A. PFAS and HCC","", "B. Dose-response of PFHpA and HCC", "", "C. Dose-response of PFNA and HCC", ""),
                           label_x = c(-0.06, 0, -0.14, 0, -0.14, 0),
                           label_size = 12,
                           align = "v",
                           axis = "lt"))


# ggsave(filename = fs::path(dir_figures,
#                            # "main_figures",
#                            "Fig1. Main Effect Final.jpeg"),
#        width = 10,
#        height = 12,
#        dpi = 300,
#        bg = "white")
```

