---
title: 'MEC analysis'
subtitle: "Final Analysis"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
```{r setup univariate analysis, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("!libraries.R")))
source(fs::path(here::here("!directories.R")))
source(fs::path(here::here("functions.R")))
source(fs::path(here::here("!load_data.R")))
```

## Analysis1. Conditional logistic regression - HCC
**Outcome: HCC vs control**

**Exposure: lg2 transformed and then normalzied PFAS**

**Covars: bmi, smoking status, alcohol intake, diabetes**

**Analysis: conditional logistic regression**

**Data Info: 237 cases, 237 controls**

**All races(474), White(46), African American(52), Japanese American(182), Latino(160), Native Hawaiian(34)**

```{r analysis1}

res_hcc <- epiomics::owas_clogit(data_hcc , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars,
                                    omics = pfas_name,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "Overall")

eth_categories <- unique(data_hcc$eth_new)

for(race in eth_categories) {
  res_hcc_temp <- epiomics::owas_clogit(data_hcc %>% filter(eth_new == race) , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars,
                                    omics = pfas_name,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = race)
  res_hcc <- res_hcc %>% bind_rows(res_hcc_temp)
}


# reorder PFAS by estimate
res_hcc1 <- res_hcc |> 
  mutate(pfas_name = str_remove(feature_name, "pfas_lg2_"),
         sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "FDR p value < 0.05",
                          p_value < 0.05 ~ "p value < 0.05", 
                         p_value >= 0.05 ~ "p value ≥ 0.05"),
         sig = factor(sig, levels = c("p value ≥ 0.05", "p value < 0.05", "FDR p value < 0.05"))) 
# writexl::write_xlsx(all_results_df_hcc, fs::path(dir_result, "1_single_exposure_analysis_result_hcc.xlsx"))
```

### Fig1a. association of pfas and hcc - HCC
```{r plot with covars only, fig.width=10, fig.height=5}
### Create metabolites Coefficient Plots
### Rename PFAS for plotting
res_hcc2 <- res_hcc1 %>%
  mutate(pfas_name = case_when(pfas_name == "4H_PFBA" ~ "4H-PFBA",
                               pfas_name == "10_2_FToH" ~ "10:2 FTOH",
                               pfas_name == "2_aminohexafluoropropan_2_ol" ~ "2-Aminohexafluoropropan-2-ol",
                               pfas_name == "Me_PFHxA" ~ "MePFHxA",
                               TRUE ~ pfas_name),
         ) %>%
  mutate(pfas_name = fct_reorder(pfas_name, if_else(type == "Overall", estimate, NA_real_))) %>%
  mutate(type = case_when(type == "Overall" ~ "Overall \n(n = 474)",
                          type == "Japanese American" ~ "Japanese American \n(n = 182)", 
                        type == "Latino"~ "Latino \n(n = 160)",
                        type == "Others" ~ "Others \n(n = 132)"))



(pfhpa_hcc_main <- res_hcc2  %>%
  ggplot(aes(x = pfas_name,y = exp(estimate), color = sig)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = exp(conf_low),
                    ymax = exp(conf_high)),
                width = 0) +
  geom_hline(yintercept = 1, linetype = 2, color = "grey") +
  ylab("Odds Ratio (95% CI)") +
  # ggtitle("Exposure Outcome Analysis for HCC Cases") +
  facet_grid(.~ factor(type, levels = c("Overall \n(n = 474)", "Japanese American \n(n = 182)", "Latino \n(n = 160)", "Others \n(n = 132)")), scales = "free_y")+
  theme(panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 15),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.title.y=element_blank(),
        axis.line.y = element_line(color = "black"),
        legend.position = "bottom",
        legend.title = element_blank())+
  coord_flip() +
  scale_color_manual(values = c("grey","red","purple")))


ggsave(filename = fs::path(dir_figures,
                           "Fig 1a Single Exposure Analysis HCC with covars_stratified_by_eth.jpeg"),
       width = 10,
       height = 5,
       dpi = 300,
       bg = "white")

writexl::write_xlsx(res_hcc2, fs::path(dir_result, "T1.Continuous PFAS and HCC.xlsx"))
```


## Sensitivity Analysis1. Conditional logistic regression - HCC
**Outcome: HCC vs control**

**Exposure: lg2 transformed and then normalzied PFAS**

**Covars: bmi, smoking status, alcohol intake, diabetes and PCs**

**Analysis: conditional logistic regression**

**Data Info: 237 cases, 237 controls**

**All races(474), White(46), African American(52), Japanese American(182), Latino(160), Native Hawaiian(34)**

```{r analysis1}
# data_hcc <- data_hcc %>%
#   mutate(eth = as.character(eth)) %>%
#   mutate(eth_new = ifelse(eth %in% c("White", "African American", "Native Hawaiian"), "Others", eth))
pc_name <- c("v1", "v2", "v3", "v4", "v5", "v6")

res_hcc <- epiomics::owas_clogit(data_hcc , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = c(covars, pc_name),
                                    omics = pfas_name,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "Overall")

eth_categories <- unique(data_hcc$eth)

for(race in eth_categories[c(1,2)]) {
  res_hcc_temp <- epiomics::owas_clogit(data_hcc %>% filter(eth == race) , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = c(covars, pc_name),
                                    omics = pfas_name,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = race)
  res_hcc <- res_hcc %>% bind_rows(res_hcc_temp)
}


# reorder PFAS by estimate
res_hcc1 <- res_hcc |> 
  mutate(pfas_name = str_remove(feature_name, "pfas_lg2_"),
         sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "FDR p value < 0.05",
                          p_value < 0.05 ~ "p value < 0.05", 
                         p_value >= 0.05 ~ "p value ≥ 0.05"),
         sig = factor(sig, levels = c("p value ≥ 0.05", "p value < 0.05", "FDR p value < 0.05"))) 
# writexl::write_xlsx(all_results_df_hcc, fs::path(dir_result, "1_single_exposure_analysis_result_hcc.xlsx"))
```

### Fig1a. association of pfas and hcc - HCC
```{r plot with covars only, fig.width=10, fig.height=5}
### Create metabolites Coefficient Plots
### Rename PFAS for plotting
res_hcc2 <- res_hcc1 %>%
  mutate(pfas_name = case_when(pfas_name == "4H_PFBA" ~ "4H-PFBA",
                               pfas_name == "10_2_FToH" ~ "10:2 FTOH",
                               pfas_name == "2_aminohexafluoropropan_2_ol" ~ "2-Aminohexafluoropropan-2-ol",
                               pfas_name == "Me_PFHxA" ~ "MePFHxA",
                               TRUE ~ pfas_name),
         ) %>%
  mutate(pfas_name = fct_reorder(pfas_name, if_else(type == "Overall", estimate, NA_real_))) %>%
  mutate(type = case_when(type == "Overall" ~ "Overall \n(n = 474)",
                          type == "Japanese American" ~ "Japanese American \n(n = 182)", 
                        type == "Latino"~ "Latino \n(n = 160)"))



(pfhpa_hcc_main <- res_hcc2  %>%
  ggplot(aes(x = pfas_name,y = exp(estimate), color = sig)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = exp(conf_low),
                    ymax = exp(conf_high)),
                width = 0) +
  geom_hline(yintercept = 1, linetype = 2, color = "grey") +
  ylab("Odds Ratio (95% CI)") +
  # ggtitle("Exposure Outcome Analysis for HCC Cases") +
  facet_grid(.~ factor(type, levels = c("Overall \n(n = 474)", "Japanese American \n(n = 182)", "Latino \n(n = 160)")), scales = "free_y")+
  theme(panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 15),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.title.y=element_blank(),
        axis.line.y = element_line(color = "black"),
        legend.position = "bottom",
        legend.title = element_blank())+
  coord_flip() +
  scale_color_manual(values = c("grey","red","purple")))


ggsave(filename = fs::path(dir_figures,
                           "figures",
                           "Fig 1a Single Exposure Analysis HCC with covars_stratified_by_eth_with_covars_pcs.jpeg"),
       width = 10,
       height = 5,
       dpi = 300,
       bg = "white")

writexl::write_xlsx(res_hcc2, fs::path(dir_result, "T1.Continuous PFAS and HCC.xlsx"))
```


## Analysis2. Conditional logistic regression (NAFLD related HCC)
**Outcome: NAFLD related HCC vs control**

**Exposure: lg2 transformed and then normalized pfas**

**Analysis: conditional logistic regression**

**Data Info: 96 cases, 96 controls**

**All races(192), White(28), African American(16), Japanese American(78), Latino(52), Native Hawaiian(18)**
```{r analysis2}
res_nafld_hcc <- epiomics::owas_clogit(data_nafld_hcc, 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars,
                                    omics = pfas_name,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "All Races")


for(race in eth_categories[3:4]) {
  res_nafld_hcc_temp <- epiomics::owas_clogit(data_nafld_hcc %>% filter(eth == race) , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars,
                                    omics = pfas_name,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = race)
  res_nafld_hcc <- res_nafld_hcc %>% bind_rows(res_nafld_hcc_temp)
}
# reorder PFAS by estimate
res_nafld_hcc1 <- res_nafld_hcc |> 
  mutate(pfas_name = str_remove(feature_name, "pfas_lg2_"), 
         # pfas_name = fct_reorder(pfas_name, estimate),
         pfas_name = factor(pfas_name, levels = levels_name),
         sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "Sig.(FDR < 0.05)",
                          p_value < 0.05 ~ "Sig.(p value < 0.05)", 
                         p_value >= 0.05 ~ "Not Sig."))

# writexl::write_xlsx(all_results_df_nafld_hcc, fs::path(dir_result, "1_single_exposure_analysis_result_nafld_hcc.xlsx"))
```

### Fig 1b. association of pfas and hcc - NAFLD HCC
```{r plot with covars only NAFLD HCC, fig.height=5, fig.width=10}
### Create metabolites Coefficient Plots
### Create metabolites Coefficient Plots
res_nafld_hcc1  %>%
  ggplot(aes(x = pfas_name,y = estimate, color = sig)) +
  geom_point(size = 1) +
  coord_flip()+
  geom_errorbar(aes(ymin = conf_low,
                    ymax = conf_high),
                width = 0) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey") +
  ylab("Log Odds Ratio (95% CI)") +
  ggtitle("Exposure Outcome Analysis for NAFLD-Related HCC Cases") +
  facet_grid( .~ type, space = "fixed", scales = "free")+
  theme(panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 15),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.title.y=element_blank(),
        axis.line.y = element_line(color = "black"),
        legend.position = "right")+
  scale_color_manual(values = c("grey","red"))

ggsave(filename = fs::path(dir_figures,
                           "Fig 1b Single Exposure Analysis NAFLD HCC with covars by eth.jpeg"),
       width = 10,
       height = 5,
       bg = "white")
```

## Analysis3. Conditional logistic regression (Other HCC)
**Outcome: Other HCC vs control**

**Exposure: lg2 transformed and then normalized PFAS**

**Analysis: conditional logistic regression**

**Data Info: 141 cases, 141 controls**

**All races(282), White(18), African American(36), Japanese American(104), Latino(108), Native Hawaiian(16)**
```{r analysis3}
res_other_hcc <- epiomics::owas_clogit(data_other_hcc, 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars,
                                    omics = pfas_name,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "All Races")

for(race in eth_categories[3:4]) {
  res_other_hcc_temp <- epiomics::owas_clogit(data_other_hcc %>% filter(eth == race) , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = covars,
                                    omics = pfas_name,
                                    conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = race)
  res_other_hcc <- res_other_hcc %>% bind_rows(res_other_hcc_temp)
}
# reorder PFAS by estimate
res_other_hcc1 <- res_other_hcc |> 
  mutate(pfas_name = str_remove(feature_name, "pfas_lg2_"), 
         # pfas_name = fct_reorder(pfas_name, estimate),
         pfas_name = factor(pfas_name, levels = levels_name),
         sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "Sig.(FDR < 0.05)",
                          p_value < 0.05 ~ "Sig.(p value < 0.05)", 
                         p_value >= 0.05 ~ "Not Sig."))

# writexl::write_xlsx(all_results_df_other_hcc, fs::path(dir_result, "1_single_exposure_analysis_result_other_hcc.xlsx"))
```

### Fig 1c. association of pfas and hcc - other HCC
```{r plot with covars only other HCC, fig.width=10, fig.height=5}
### Create metabolites Coefficient Plots
res_other_hcc1  %>%
  ggplot(aes(x = pfas_name,y = estimate, color = sig)) +
  geom_point(size = 1) +
  coord_flip()+
  geom_errorbar(aes(ymin = conf_low,
                    ymax = conf_high),
                width = 0) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey") +
  ylab("Log Odds Ratio (95% CI)") +
  ggtitle("Exposure Outcome Analysis for NAFLD-Related HCC Cases") +
  facet_grid( . ~ type,  scale = "free")+
  theme(panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 15),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.title.y=element_blank(),
        axis.line.y = element_line(color = "black"),
        legend.position = "right")+
  coord_flip() +
  scale_color_manual(values = c("grey","purple","red"))


ggsave(filename = fs::path(dir_figures,
                           "Fig 1c Single Exposure Analysis other HCC with covars by eth.jpeg"),
       width = 10,
       height = 5,
       bg = "white")
```


