---
title: 'MEC analysis'
subtitle: "Final Analysis"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
```{r setup PRS, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("!libraries.R")))
source(fs::path(here::here("!directories.R")))
source(fs::path(here::here("functions.R")))
source(fs::path(here::here("!load_data.R")))
```


## Analysis1: Categorical PRS and HCC
* conditional logistic regression
* unweighted PRS
```{r prs and hcc}
res <- clogit(as.formula(paste("status ~ prs_uw_tertile +scale(bmi)+ alcohol_intake + smokestatus + diabetes + strata(setnum) + v1 + v2 + v3 + v4 + v5 + v6")), 
          data = data_hcc,
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    filter(!term %in% c("scale(bmi)", "alcohol_intake", "smokestatusCurrent", "smokestatusFormer", "diabetesYes", "v1", "v2", "v3", "v4", "v5", "v6"
                        # , "casetypeNAFLD-related HCC", "casetypeOther HCC"
                        )) %>% 
    mutate(
           odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "Adjusted by covars",
           type = "All Races")

eth_categories <- levels(data_hcc$eth)

for(race in eth_categories[3:4]) {
  res_temp <- clogit(as.formula(paste("status ~ prs_uw_tertile +scale(bmi)+ alcohol_intake + smokestatus + diabetes + strata(setnum) + v1 + v2 + v3 + v4 + v5 + v6")), 
          data = data_hcc %>% filter(eth == race),
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    filter(!term %in% c("scale(bmi)", "alcohol_intake", "smokestatusCurrent", "smokestatusFormer", "diabetesYes", "v1", "v2", "v3", "v4", "v5", "v6")) %>% 
    mutate(
           odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "Adjusted by covars",
           type = race)
  res <- res %>% bind_rows(res_temp)
}

```

### Fig 2a. prs and hcc- HCC
```{r prs and hcc plot}
intercept_data<- tibble(term = rep("prs_uw_tertile1", 3),
                        odds_ratio = rep(1, 3),
                        type = c("All Races", "Japanese American", "Latino"))

all_results_df_hcc2 <- res |> 
  bind_rows(intercept_data)%>%
  mutate(
         genetic_risk = case_when(str_detect(term, "tertile3") ~ "High", 
                                  str_detect(term, "tertile2") ~ "Medium",
                                  TRUE ~ "Low") |>
           fct_relevel(c("Low", "Medium", "High")), 
         sig = if_else(p.value < 0.05, "Sig.", "Not Sig.")) |>
  mutate(color = case_when( p.value < 0.05 ~ "red", 
                            TRUE ~ "grey60"))

(grs_plot <- ggplot(all_results_df_hcc2, 
       aes(x = genetic_risk, y = odds_ratio, color = color)) +
  geom_errorbar(aes(ymin = exp_ci_low, ymax = exp_ci_high),
                width = 0, position = position_dodge(width = 1)) +
  geom_point(size = 1, position = position_dodge(width = 1)) + 
  geom_hline(yintercept = 1, linetype = 1, color = "grey50") +
  facet_wrap(.~type, scales = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), 
        legend.position = "none",
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 10),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black")) +
      ylab("Odds Ratio (95% CI)") + 
  xlab("NAFLD PRS Tertile") +
  # scale_y_continuous(breaks = c(1,3,5,7)) +
  scale_color_manual(values = c("black", "red")))

ggsave(grs_plot, filename = fs::path(dir_figures,
                           "Fig 2a PRS and HCC.jpeg"),
       width = 4,
       height = 2,
       bg = "white")

```


## Analysis2: Categorical PRS and NAFLD related HCC
* conditional logistic regression
* unweighted PRS
```{r prs and nafld hcc}
res_nafld <- clogit(as.formula(paste("status ~ prs_uw_tertile +scale(bmi)+ alcohol_intake + smokestatus + diabetes + strata(setnum) + v1 + v2 + v3 + v4 + v5 + v6")), 
          data = data_nafld_hcc,
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    filter(!term %in% c("scale(bmi)", "alcohol_intake", "smokestatusCurrent", "smokestatusFormer", "diabetesYes", "v1", "v2", "v3", "v4", "v5", "v6"
                        # , "casetypeNAFLD-related HCC", "casetypeOther HCC"
                        )) %>% 
    mutate(
           odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "Adjusted by covars",
           type = "All Races")

# for(race in eth_categories[3:4]) {
#   res_temp <- clogit(as.formula(paste("status ~ prs_uw_tertile +scale(bmi)+ alcohol_intake + smokestatus + diabetes + strata(setnum) + v1 + v2 + v3 + v4 + v5 + v6")), 
#           data = data_nafld_hcc %>% filter(eth == race),
#           method = "efron", robust = TRUE) %>%
#     tidy(., conf.int = TRUE, level = 0.95) %>% 
#     filter(!term %in% c("scale(bmi)", "alcohol_intake", "smokestatusCurrent", "smokestatusFormer", "diabetesYes", "v1", "v2", "v3", "v4", "v5", "v6")) %>% 
#     mutate(
#            odds_ratio = exp(estimate),
#            exp_ci_low = exp(conf.low),
#            exp_ci_high = exp(conf.high),
#            model_type = "clogit", 
#            covariates = "Adjusted by covars",
#            type = race)
#   res_nafld <- res_nafld %>% bind_rows(res_temp)
# }
# This part doesn't work because of sample size
```

### Fig 2b. prs and hcc- NAFLD realted HCC
```{r prs and nafld related hcc plot}
intercept_data<- tibble(term = rep("prs_uw_tertile1", 3),
                        odds_ratio = rep(1, 3),
                        type = c("All Races", "Japanese American", "Latino"))

all_results_df_nafld_hcc2 <- res_nafld |> 
  bind_rows(intercept_data) %>%
  mutate(
         genetic_risk = case_when(str_detect(term, "tertile3") ~ "High", 
                                  str_detect(term, "tertile2") ~ "Medium",
                                  TRUE ~ "Low") |>
           fct_relevel(c("Low", "Medium", "High")), 
         sig = if_else(p.value < 0.05, "Sig.", "Not Sig.")) |>
  mutate(color = case_when( p.value < 0.05 ~ "red", 
                            TRUE ~ "grey60"))

(grs_plot <- ggplot(all_results_df_nafld_hcc2, 
       aes(x = genetic_risk, y = odds_ratio, color = color)) +
  geom_errorbar(aes(ymin = exp_ci_low, ymax = exp_ci_high),
                width = 0, position = position_dodge(width = 1)) +
  geom_point(size = 1, position = position_dodge(width = 1)) + 
  geom_hline(yintercept = 1, linetype = 1, color = "grey50") +
  # facet_wrap(.~type, scales = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), 
        legend.position = "none",
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 10),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black")) +
      ylab("Odds Ratio (95% CI)") + 
  xlab("NAFLD PRS Tertile") +
  scale_color_manual(values = c("black", "red")))

ggsave(grs_plot, filename = fs::path(dir_figures,
                           "Fig 2b PRS and NAFLD related HCC.jpeg"),
       width = 2.5,
       height = 2.5,
       bg = "white")

```

## Analysis3: Categorical PRS and other HCC
* conditional logistic regression
* unweighted tertiled PRS
```{r prs and other hcc}
res_other <- clogit(as.formula(paste("status ~ prs_uw_tertile +scale(bmi)+ alcohol_intake + smokestatus + diabetes + strata(setnum) + v1 + v2 + v3 + v4 + v5 + v6")), 
          data = data_other_hcc,
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    filter(!term %in% c("scale(bmi)", "alcohol_intake", "smokestatusCurrent", "smokestatusFormer", "diabetesYes", "v1", "v2", "v3", "v4", "v5", "v6"
                        # , "casetypeNAFLD-related HCC", "casetypeOther HCC"
                        )) %>% 
    mutate(
           odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "Adjusted by covars",
           type = "All Races")

for(race in eth_categories[3:4]) {
  res_temp <- clogit(as.formula(paste("status ~ prs_uw_tertile +scale(bmi)+ alcohol_intake + smokestatus + diabetes + strata(setnum) + v1 + v2 + v3 + v4 + v5 + v6")), 
          data = data_other_hcc %>% filter(eth == race),
          method = "efron", robust = TRUE) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    filter(!term %in% c("scale(bmi)", "alcohol_intake", "smokestatusCurrent", "smokestatusFormer", "diabetesYes", "v1", "v2", "v3", "v4", "v5", "v6")) %>% 
    mutate(
           odds_ratio = exp(estimate),
           exp_ci_low = exp(conf.low),
           exp_ci_high = exp(conf.high),
           model_type = "clogit", 
           covariates = "Adjusted by covars",
           type = race)
  res_other <- res_other %>% bind_rows(res_temp)
}
```

### Fig 2c. prs and hcc- other HCC
```{r prs and other hcc plot}

intercept_data<- tibble(term = rep("prs_uw_tertile1", 3),
                        odds_ratio = rep(1, 3),
                        type = c("All Races", "Japanese American", "Latino"))

all_results_df_other_hcc2 <- res_other |> 
  bind_rows(intercept_data) |>
  mutate(
         genetic_risk = case_when(str_detect(term, "tertile3") ~ "High", 
                                  str_detect(term, "tertile2") ~ "Medium",
                                  TRUE ~ "Low") |>
           fct_relevel(c("Low", "Medium", "High")), 
         sig = if_else(p.value < 0.05, "Sig.", "Not Sig.")) |>
  mutate(color = case_when( p.value < 0.05 ~ "red", 
                            TRUE ~ "grey60"))

(grs_plot <- ggplot(all_results_df_other_hcc2, 
       aes(x = genetic_risk, y = odds_ratio, color = color)) +
  geom_errorbar(aes(ymin = exp_ci_low, ymax = exp_ci_high),
                width = 0, position = position_dodge(width = 1)) +
  geom_point(size = 1, position = position_dodge(width = 1)) + 
  geom_hline(yintercept = 1, linetype = 1, color = "grey50") +
  facet_wrap(.~type, scales = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), 
        legend.position = "none",
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_text(angle = 0, hjust = 0),
        # strip.text.x = element_blank(),
        text = element_text(size = 10),
        # axis.title.x=element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black")) +
      ylab("Odds Ratio (95% CI)") + 
  xlab("NAFLD PRS Tertile") +
  # scale_y_continuous(breaks = c(1,3,5,7)) +
  scale_color_manual(values = c("black", "red")))

ggsave(grs_plot, filename = fs::path(dir_figures,
                           "Fig 2c PRS and other HCC.jpeg"),
       width = 5,
       height = 3,
       bg = "white")

```
