---
title: 'MEC analysis'
subtitle: "Final Analysis"
author: "Hongxu Wang, Jesse Goodrich"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup univariate analysis cat pfas, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("1_Project_Setup/!libraries.R")))
source(fs::path(here::here("1_Project_Setup/!directories.R")))
source(fs::path(here::here("1_Project_Setup/!functions.R")))
source(fs::path(here::here("1_Project_Setup/!load_data.R")))

```

## Analysis1. Conditional logistic regression - HCC
**Outcome: HCC vs control**

**Exposure: normalzied log2 PFAS**

**Covars: smoking, diabetes, q1_edih, prs_wt_std, v1:v6**

**Analysis: conditional logistic regression**

**Data Info: 223 cases, 223 controls**

```{r analysis1}
res_hcc <- epiomics::owas_clogit(data_analysis, 
                                 cc_set = "setnum", 
                                 cc_status = "status", 
                                 covars = covars,
                                 omics = c(pfas_name, "pfas_t_PFBS_detected"),
                                 conf_int = TRUE,
                                 method = "exact") %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "Overall")

eth_categories <- unique(data_analysis$eth_new)

for(race in eth_categories) {
  res_hcc_temp <- epiomics::owas_clogit(data_analysis %>% filter(eth_new == race) , 
                                        cc_set = "setnum", 
                                        cc_status = "status", 
                                        covars = covars,
                                        omics = c(pfas_name, "pfas_t_PFBS_detected"),
                                        conf_int = TRUE,
                                        method = "exact") %>%
    mutate(odds_ratio = exp(estimate),
           exp_ci_low = exp(conf_low),
           exp_ci_high = exp(conf_high),
           type = race)
  res_hcc <- res_hcc %>% bind_rows(res_hcc_temp)
}
rm(res_hcc_temp)

# reorder PFAS by estimate
res_hcc1 <- res_hcc |> 
  mutate(pfas_name = str_remove(feature_name, "pfas_t_"),
         sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "FDR p value < 0.05",
                         p_value < 0.05 ~ "p value < 0.05", 
                         p_value >= 0.05 ~ "p value ≥ 0.05"),
         sig = factor(sig, levels = c("p value ≥ 0.05", "p value < 0.05", "FDR p value < 0.05"))) %>%
  mutate(covar = sub(".*~(.*?)\\+pfas.*", "\\1", formula))%>%
  select(covar, everything())
```

### Fig1a. association of pfas and hcc - HCC
```{r plot with covars only, fig.width=10, fig.height=5}
### Compute sample sizes from data
n_overall <- nrow(data_analysis)
n_japanese <- nrow(data_analysis %>% filter(eth == "Japanese American"))
n_latino <- nrow(data_analysis %>% filter(eth == "Latino"))
n_white <- nrow(data_analysis %>% filter(eth == "White"))
n_aa <- nrow(data_analysis %>% filter(eth == "African American"))
n_nh <- nrow(data_analysis %>% filter(eth == "Native Hawaiian"))
n_others <- nrow(data_analysis %>% filter(eth_new == "Others"))

label_overall <- paste0("Overall \n(n cases: ", n_overall/2, "\nn controls: ", n_overall/2, ")")
label_japanese <- paste0("Japanese American \n(n cases: ", n_japanese/2, "\nn controls: ", n_japanese/2, ")")
label_latino <- paste0("Latino \n(n cases: ", n_latino/2, "\nn controls: ", n_latino/2, ")")
label_white <- paste0("White \n(n cases: ", n_white/2, "\nn controls: ", n_white/2, ")")
label_aa <- paste0("African American \n(n cases: ", n_aa/2, "\nn controls: ", n_aa/2, ")")
label_nh <- paste0("Native Hawaiian \n(n cases: ", n_nh/2, "\nn controls: ", n_nh/2, ")")
label_others <- paste0("Others \n(n cases: ", n_others/2, "\nn controls: ", n_others/2, ")")

### Rename PFAS for plotting
res_hcc2 <- res_hcc1 %>%
  mutate(pfas_name = rename_pfas(pfas_name))%>%
  mutate(pfas_name = fct_reorder(pfas_name, if_else(type == "Overall", estimate, NA_real_))) %>%
  mutate(type = case_when(type == "Overall" ~ label_overall,
                          type == "Japanese American" ~ label_japanese,
                          type == "Latino"~ label_latino,
                          type == "Others" ~ label_others
                          # type == "White" ~ label_white,
                          # type == "African American" ~ label_aa,
                          # type == "Native Hawaiian" ~ label_nh
                          ))

(pfhpa_hcc_main <- res_hcc2  %>%
    filter(!pfas_name %in% c("PFBS", "PFBS_detected"))%>%
    ggplot(aes(x = pfas_name,y = exp(estimate), color = sig)) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = exp(conf_low),
                      ymax = exp(conf_high)),
                  width = 0) +
    geom_hline(yintercept = 1, linetype = 2, color = "grey") +
    ylab("Odds Ratio (95% CI)") +
    # ggtitle("Exposure Outcome Analysis for HCC Cases") +
    # facet_grid(.~ factor(type, levels = c("Overall \n(n = 474)", "Japanese American \n(n = 182)", "Latino \n(n = 158)", "Others \n(n = 132)")), scales = "free_y")+
    facet_grid(.~ factor(type, levels = c(label_overall, label_japanese, label_latino, label_others)), scales = "free")+
    theme(panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_text(angle = 0, hjust = 0),
          # strip.text.x = element_blank(),
          text = element_text(size = 15),
          # axis.title.x=element_blank(),
          axis.line.x = element_line(color = "black"),
          axis.title.y=element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
    coord_flip() +
    scale_color_manual(values = c("grey","red","purple")))


writexl::write_xlsx(res_hcc2, fs::path(dir_result, "T1.Continuous PFAS and HCC_081026.xlsx"))
```


## Analysis2. Conditional logistic regression - Categorized PFAS and HCC
**Outcome: HCC vs control**

**Exposure: categorized pfas(quartile)**

**Covars: smoking, diabetes, q1_edih, prs_wt_std, v1:v6**

**Analysis: conditional logistic regression**


```{r analysis1 cat}

covariate_input <- paste(covars, collapse = " + ")

## create dataframe with names of dependent vars (PFAS) ----
pfas_names <- tibble(exposure = pfas_name_cat)

# Get Model covariates 
covar_names <- tibble(
  covariates =  covariate_input)

# Make all clogit formulas
ind_vars <- expand_grid(pfas_names,  covar_names) |>  
  mutate(formula = paste("status ~ ", 
                         exposure, 
                         "+", covariates, 
                         "+strata(setnum)") |>
           str_replace("\\+  \\+", " + "))

## Analysis for all related participants
ind_vars$clogit_out <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_analysis,
          method = "exact"
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Japanese American 
ind_vars$clogit_out1 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_analysis %>% filter(eth == "Japanese American"),
          method = "exact"
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Latino
ind_vars$clogit_out2 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_analysis %>% filter(eth == "Latino"),
          method = "exact"
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Others
ind_vars$clogit_out3 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_analysis %>% filter(eth_new == "Others"),
          method = "exact"
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit"))
# Get final data -------------------

intercept_data<- tibble(exposure = pfas_name_cat, 
                        term = paste0(pfas_name_cat, "1"), 
                        odds_ratio = rep(1,length(pfas_name_cat)),
                        effect_est  = rep("1",length(pfas_name_cat)))

all_results_df_hcc <- ind_vars %>% 
  unnest(clogit_out) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out1, -clogit_out2) %>%
  filter(grepl("pfas", term)) %>%
  bind_rows(intercept_data) %>% 
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type = "Overall")

all_results_df_hcc_japan <- ind_vars %>% 
  unnest(clogit_out1) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out2) %>%
  filter(grepl("pfas", term))%>%
  bind_rows(intercept_data) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type= "Japanese American")

all_results_df_hcc_latino <- ind_vars %>% 
  unnest(clogit_out2) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out1) %>%
  filter(grepl("pfas", term))%>%
  bind_rows(intercept_data) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type = "Latino")

all_results_df_hcc_others <- ind_vars %>% 
  unnest(clogit_out3) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out1) %>%
  filter(grepl("pfas", term))%>%
  bind_rows(intercept_data) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type = "Others")


all_result_df <- all_results_df_hcc %>% 
  bind_rows(all_results_df_hcc_japan) %>%
  bind_rows(all_results_df_hcc_latino) %>%
  bind_rows(all_results_df_hcc_others) %>%
  select(exposure, term, odds_ratio, exp_ci_low, exp_ci_high,effect_est,p.value,p_adjusted, type) %>%
  arrange(exposure)

```


```{r Dose response}
data_analysis1 <- data_analysis %>% 
  mutate_at(.vars = pfas_name_cat,
            .funs = ~as.numeric(.))

res_hcc <- epiomics::owas_clogit(data_analysis1 , 
                                 cc_set = "setnum", 
                                 cc_status = "status", 
                                 covars = covars,
                                 omics = pfas_name_cat,
                                 conf_int = TRUE,
                                 method = "exact") %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "Overall")

eth_categories <- unique(data_analysis1$eth_new)

for(race in eth_categories) {
  res_hcc_temp <- epiomics::owas_clogit(data_analysis1 %>% filter(eth_new == race) , 
                                        cc_set = "setnum", 
                                        cc_status = "status", 
                                        covars = covars,
                                        omics = c(pfas_name_cat),
                                        conf_int = TRUE,
                                        method = "exact") %>%
    mutate(odds_ratio = exp(estimate),
           exp_ci_low = exp(conf_low),
           exp_ci_high = exp(conf_high),
           type = race)
  res_hcc <- res_hcc %>% bind_rows(res_hcc_temp)
}

# reorder PFAS by estimate
res_hcc1 <- res_hcc |> 
  mutate(sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "Sig.(FDR < 0.05)",
                         p_value < 0.05 ~ "Sig.(p value < 0.05)", 
                         p_value >= 0.05 ~ "Not Sig.")) 

p_trend_df <- res_hcc1 %>% select(feature_name, p_value, adjusted_pval,type) %>%
  rename(exposure = feature_name,
         p_trend = p_value)

all_result_df1 <- all_result_df %>% 
  tidylog::left_join(p_trend_df, by = c("exposure", "type")) %>%
  mutate(pfas_name = str_remove(exposure, "pfas_lg2_") %>%
           str_remove("_quartile")) %>%
  select(type,pfas_name, term,everything(), -exposure)

## N value for each quartile split by status

# Convert to long format and create a pair identifier
data_list <- list(data1 = data_analysis %>% mutate(type = "Overall"), 
                  data2 = data_analysis %>% filter(eth == "Japanese American") %>%
                    mutate(type = "Japanese American"), 
                  data3 = data_analysis %>% filter(eth == "Latino") %>%
                    mutate(type = "Latino"),
                  data4 = data_analysis %>% filter(eth_new == "Others") %>%
                    mutate(type = "Others"))

data_n <- data.frame()

for(data in data_list){
  type = unique(data$type)
  long_data_analysis <- data %>%
    select(contains("quartile"), status) %>%
    mutate(status = ifelse(status == 0, "Control", "Case")) %>%
    pivot_longer(cols = c(contains("quartile")), names_to = "variable",
                 values_to = "value") %>%
    drop_na() %>%
    mutate(pair = paste0(variable, value)) %>%
    select(-variable,-value) %>%
    group_by(pair, status) %>%
    summarise(count = n(), .groups = "drop")
  # Now, to spread back for a wide format summary
  wide_summary_hcc <- pivot_wider(long_data_analysis, names_from = status,
                                  values_from = count, values_fill = list(count = 0))%>%
    mutate(type = type) %>%
    rename(term = pair)
  data_n <- data_n %>% bind_rows(wide_summary_hcc)
}

all_result_df2 <- all_result_df1 %>% 
  tidylog::left_join(data_n, by = c("term","type")) %>%
  select(type:term, Control, Case, everything())%>%
  arrange(pfas_name, type, term)

all_result_df3 <- all_result_df2 %>% 
  mutate(plot_name = substr(term, 
                            start = nchar(term) - 8,
                            stop = nchar(term)) %>% toTitleCase(),
         pfas_name = str_remove(pfas_name, "pfas_t_")) %>%
  mutate(
    plot_name = paste(substr(plot_name, 1, 8), substr(plot_name, 9, 9)),
    sig = case_when(p_adjusted < 0.05 ~ "FDR. Sig.", 
                    p.value < 0.05 ~ "Sig.",
                    TRUE ~ "Not Sig."),
    label = str_c("p trend = ", format(signif(p_trend, digits = 2), scientific = FALSE))) %>%
  mutate(type = case_when(type == "Overall" ~ label_overall,
                          type == "Japanese American" ~ label_japanese,
                          type == "Latino"~ label_latino,
                          type == "Others" ~ label_others))

writexl::write_xlsx(all_result_df3, fs::path(dir_result, "single_categorical_exposure_analysis_081026.xlsx"))

```

### Fig1. Association of PFAS and HCC - Overall only (continuous)
```{r plot overall only continuous, fig.width=5, fig.height=5}
(pfhpa_hcc_overall <- res_hcc2 %>%
    filter(!pfas_name %in% c("PFBS", "PFBS_detected")) %>%
    filter(type == label_overall) %>%
    ggplot(aes(x = pfas_name, y = exp(estimate), color = sig)) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = exp(conf_low),
                      ymax = exp(conf_high)),
                  width = 0) +
    geom_hline(yintercept = 1, linetype = 2, color = "grey") +
    ylab("Odds Ratio (95% CI)") +
    theme(panel.background = element_rect(fill="white"),
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_text(angle = 0, hjust = 0),
          text = element_text(size = 15),
          axis.line.x = element_line(color = "black"),
          axis.title.y=element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
    coord_flip() +
    scale_color_manual(values = c("grey","red","purple")))
```

### Fig. Dose response plot for PFHpA - Overall only
```{r dose response PFHpA overall only}
quartile_cuts_df <- readxl::read_xlsx(fs::path(dir_result, "quartile_cuts_df.xlsx"))%>%
  mutate(quartile = str_replace(quartile, "n", "r"))%>%
  mutate(term = paste0(pfas_name, "_",
                       tolower(quartile)))
  
all_result_df3_quartile <- all_result_df3 %>%
  tidylog::left_join(quartile_cuts_df %>% select(term, value), 
                     by = "term")%>%
  mutate(value = ifelse(
    str_detect(value, "≤|>"),
    paste0("(", value, ")"),
    value
  ))%>%
  mutate(value = str_replace(value, ", ", "-") %>%
           str_replace(., "]", ")")) %>%
  mutate(value = gsub("\\)", " ng/mL)", value))%>%
  mutate(plot_name = paste0(str_remove(plot_name, "uartile "), " ", value))

pfhpa_hcc_overall_dr <- ggplot(all_result_df3_quartile %>%
                        filter(pfas_name == "PFHPA") %>%
                        filter(type == label_overall),
                      aes(x = odds_ratio, y = fct_rev(plot_name))) +
    geom_errorbar(aes(xmin = exp_ci_low, xmax = exp_ci_high),
                  width = 0, position = position_dodge(width = 1)) +
    geom_point(size = 1, position = position_dodge(width = 1)) +
    geom_vline(xintercept = 1, linetype = 2, color = "grey") +
    xlab("Odds Ratio (95% CI)") +
    theme(legend.position = "bottom",
      panel.background = element_rect(fill="white"),
      strip.background = element_rect(fill = "white"),
      strip.text.y = element_text(angle = 0, hjust = 0),
      text = element_text(size = 15),
      axis.title.y = element_blank(),
      axis.line.x = element_line(color = "black"),
      axis.line.y = element_line(color = "black"))

(p1_overall <- pfhpa_hcc_overall_dr + geom_text(
  data = all_result_df3 %>%
    filter(pfas_name == "PFHPA") %>%
    filter(type == label_overall),
  mapping = aes(x = Inf, y = Inf, label = label),
  color = "black",
  hjust   = 1,
  vjust   = 5,
  size = 4
))
```

### Fig. Dose response plot for PFNA - Overall only
```{r dose response PFNA overall only}
pfna_hcc_overall_dr <- ggplot(all_result_df3_quartile %>%
                        filter(pfas_name == "PFNA") %>%
                        filter(type == label_overall),
                      aes(x = odds_ratio, y = fct_rev(plot_name))) +
    geom_errorbar(aes(xmin = exp_ci_low, xmax = exp_ci_high),
                  width = 0, position = position_dodge(width = 1)) +
    geom_point(size = 1, position = position_dodge(width = 1)) +
    geom_vline(xintercept = 1, linetype = 2, color = "grey") +
    xlab("Odds Ratio (95% CI)") +
    theme(legend.position = "bottom",
      panel.background = element_rect(fill="white"),
      strip.background = element_rect(fill = "white"),
      strip.text.y = element_text(angle = 0, hjust = 0),
      text = element_text(size = 15),
      axis.title.y = element_blank(),
      axis.line.x = element_line(color = "black"),
      axis.line.y = element_line(color = "black"))

p2_overall <- pfna_hcc_overall_dr + geom_text(
  data = all_result_df3 %>%
    filter(pfas_name == "PFNA") %>%
    filter(type == label_overall),
  mapping = aes(x = Inf, y = Inf, label = label),
  color = "black",
  hjust   = 1,
  vjust   = 5,
  size = 4
)
```

### Fig. Combined Overall only (one row)
```{r combine figures overall only one row, fig.width=15, fig.height=5}
(fig_overall_row <- cowplot::plot_grid(NULL, NULL, NULL, pfhpa_hcc_overall, p1_overall, p2_overall,
                                       rel_heights = c(0.01, 0.2),
                                       rel_widths = c(1.6,1.8,1.8),
                           ncol = 3,
                           labels = c("A. PFAS and HCC",
                                      "B. Dose-response of PFHpA and HCC",
                                      "C. Dose-response of PFNA and HCC",
                                      "", "", ""),
                           label_x = c(0, -0.12,-0.12, 0, 0,0),
                           label_size = 12,
                           align = "h",
                           axis = "tb"))

ggsave(filename = fs::path(dir_figures,
                           "Fig1.Main_Effect_Overall_one_row_081026.jpeg"),
       width = 15,
       height = 5,
       dpi = 300,
       bg = "white")
```

## Meta-analysis - Race
```{r meta analysis}

meta_df <- res_hcc2 %>%
  filter(type != label_overall) %>%
  select(pfas_name, type, estimate, se_est)%>%
  mutate(vi = se_est^2) %>%
  select(-se_est)

# write_csv(meta_df, fs::path(dir_result, "meta_df.csv"))

res_df <- tibble()

# Function to fit model and test heterogeneity for each PFAS
analyze_pfas <- function(pfas, data) {
  # Subset data for the current PFAS
  
  subset_data <- subset(data, pfas_name == pfas)%>%
    select(-pfas_name)
  
  # Fit the random-effects model
  model <- rma(yi = estimate, vi = vi, data = subset_data, method = "REML")
  
  # Calculate and print the I^2 statistic
  # I2 <- model$I2
  model_summary <- summary(model)
  res_df <- data.frame(
    pfas = pfas,
    tau2 = model_summary$tau2,
    I2 = model_summary$I2,
    H2 = model_summary$H2,
    Q = model_summary$QE,
    Q_df = model_summary$k - 1,
    Q_pval = model_summary$QEp,
    pooled_lg_or = model_summary$b,
    se = model_summary$se,
    pval_model = model_summary$pval,
    pooled_ci.lb = model_summary$ci.lb,
    pooled_ci.ub = model_summary$ci.ub
  )
  return(res_df)
}

# Apply the function to each PFAS
for (pfas in unique(meta_df$pfas_name)) {
  print(pfas)
  temp_df <- analyze_pfas(pfas, meta_df)
  res_df <- res_df %>% bind_rows(temp_df)
}

res_df1 <- res_df %>% 
  mutate(or_ci = paste0(round(exp(pooled_lg_or),2), " [",
                        round(exp(pooled_ci.lb),2), ", ",
                        round(exp(pooled_ci.ub),2), "]"))%>%
  rename(pfas_name = pfas) %>%
  mutate(pfas_name = rename_pfas(pfas_name))

writexl::write_xlsx(res_df1, fs::path(dir_result, "meta_analysis_res_race_081026.xlsx"))
```

