---
title: 'MEC analysis'
subtitle: "Final Analysis"
author: "Hongxu Wang, Jesse Goodrich"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup univariate analysis cat pfas, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("1_Project_Setup/!libraries.R")))
source(fs::path(here::here("1_Project_Setup/!directories.R")))
source(fs::path(here::here("1_Project_Setup/!functions.R")))
source(fs::path(here::here("1_Project_Setup/!load_data.R")))

data_hcc <- data_hcc %>%
  mutate(bmi_imputed = bmi_imputed %>% scale(.)%>%as.vector())
```


## Analysis1. Conditional logistic regression - HCC
**Outcome: HCC vs control**

**Exposure: alcohol**

**Covars: smoking, diabetes, q1_edih, prs_wt_std, v1:v6**

**Analysis: conditional logistic regression**

**Data Info: 240 cases, 240 controls**

**All races(470), White(47), African American(51), Japanese American(178), Latino(160), Native Hawaiian(34)**

```{r analysis1}
exposure <- c("bmi_imputed")
res_hcc <- epiomics::owas_clogit(data_hcc, 
                                 cc_set = "setnum", 
                                 cc_status = "status", 
                                 covars = covars,
                                 omics = "bmi_imputed",
                                 conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "Overall")

eth_categories <- unique(data_hcc$eth)

for(race in eth_categories[c(1:2)]) {
  res_hcc_temp <- epiomics::owas_clogit(data_hcc %>% filter(eth == race) , 
                                        cc_set = "setnum", 
                                        cc_status = "status", 
                                        covars = covars,
                                        omics = exposure,
                                        conf_int = TRUE) %>%
    mutate(odds_ratio = exp(estimate),
           exp_ci_low = exp(conf_low),
           exp_ci_high = exp(conf_high),
           type = race)
  res_hcc <- res_hcc %>% bind_rows(res_hcc_temp)
}
rm(res_hcc_temp)
# reorder PFAS by estimate
res_hcc1 <- res_hcc |> 
  mutate(exp_name = "BMI",
         sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "FDR p value < 0.05",
                         p_value < 0.05 ~ "p value < 0.05", 
                         p_value >= 0.05 ~ "p value ≥ 0.05"),
         sig = factor(sig, levels = c("p value ≥ 0.05", "p value < 0.05", "FDR p value < 0.05"))) %>%
  mutate(covar = sub(".*~(.*?)\\+pfas.*", "\\1", formula))%>%
  select(covar, everything())
```

### Fig1a. association of BMI and hcc - HCC
```{r plot with covars only, fig.width=10, fig.height=5}
### Rename PFAS for plotting
res_hcc2 <- res_hcc1 %>%
  # mutate(pfas_name = rename_pfas(pfas_name))%>%
  # mutate(pfas_name = fct_reorder(pfas_name, if_else(type == "Overall", estimate, NA_real_))) %>%
  mutate(type = case_when(type == "Overall" ~ "Overall \n(n = 464)",
                          type == "Japanese American" ~ "Japanese American \n(n = 176)",
                          type == "Latino"~ "Latino \n(n = 158)"
                          # type == "White" ~ "White \n(n = 46)",
                          # type == "African American" ~ "African American \n(n = 50)",
                          # type == "Native Hawaiian" ~ "Native Hawaiian \n(n = 34)"
                          ))

(bmi_hcc_main <- res_hcc2  %>%
    # filter(!pfas_name %in% c("PFBS", "PFBS_detected"))%>%
    # filter(!type  %in% c("White \n(n = 46)", "African American \n(n = 50)"))%>%
    ggplot(aes(x = factor(type, levels = c("Native Hawaiian \n(n = 34)","African American \n(n = 50)","White \n(n = 46)","Latino \n(n = 158)","Japanese American \n(n = 176)","Overall \n(n = 464)")),y = exp(estimate), color = sig)) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = exp(conf_low),
                      ymax = exp(conf_high)),
                  width = 0) +
    geom_hline(yintercept = 1, linetype = 2, color = "grey") +
    ylab("Odds Ratio (95% CI)") +
    # ggtitle("Exposure Outcome Analysis for HCC Cases") +
    # facet_grid(.~ factor(type, levels = c("Overall \n(n = 474)", "Japanese American \n(n = 182)", "Latino \n(n = 158)", "Others \n(n = 132)")), scales = "free_y")+
    # facet_grid(.~ factor(type, levels = c("Overall \n(n = 464)", "Japanese American \n(n = 176)", "Latino \n(n = 158)", "White \n(n = 46)", "African American \n(n = 50)", "Native Hawaiian \n(n = 34)")), scales = "free")+ 
    theme(panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_text(angle = 0, hjust = 0),
          # strip.text.x = element_blank(),
          text = element_text(size = 15),
          # axis.title.x=element_blank(),
          axis.line.x = element_line(color = "black"),
          axis.title.y=element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
    coord_flip() +
    scale_color_manual(values = c("grey","red","purple")))


writexl::write_xlsx(res_hcc2, fs::path(dir_result, "T1.Continuous BMI and HCC_1217.xlsx"))
```


## Analysis2. Conditional logistic regression - Categorized BMI and HCC
**Outcome: HCC vs control**

**Exposure: categorized BMI**

**Covars: smoking, diabetes, q1_edih, prs_wt_std, v1:v6**

**Analysis: conditional logistic regression**

**Data Info: 240 cases, 240 controls**

**All races(470), White(47), African American(51), Japanese American(178), Latino(160), Native Hawaiian(34)**

```{r analysis1 cat}

covariate_input <- paste(covars, collapse = " + ")

## create dataframe with names of dependent vars (PFAS) ----
exp_names <- tibble(exposure = c("bmi_cat_imputed"))

# Get Model covariates 
covar_names <- tibble(
  covariates =  covariate_input)

# Make all clogit formulas
ind_vars <- expand_grid(exp_names,  covar_names) |>  
  mutate(formula = paste("status ~ ", 
                         exposure, 
                         "+", covariates, 
                         "+strata(setnum)") |>
           str_replace("\\+  \\+", " + "))

## Analysis for all related participants
ind_vars$clogit_out <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_hcc,
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Japanese American 
ind_vars$clogit_out1 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_hcc %>% filter(eth == "Japanese American"),
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Latino
ind_vars$clogit_out2 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_hcc %>% filter(eth == "Latino"),
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Others
ind_vars$clogit_out3 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_hcc %>% filter(eth_new == "Others"),
          method = "efron",
          robust = TRUE
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit"))
# Get final data -------------------

intercept_data<- tibble(exposure = c("bmi_cat_imputed"), 
                        term = paste0("bmi_cat_imputed", "<25"), 
                        odds_ratio = 1,
                        effect_est  = "1")

all_results_df_hcc <- ind_vars %>% 
  unnest(clogit_out) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out1, -clogit_out2) %>%
  filter(grepl("bmi", term)) %>%
  bind_rows(intercept_data) %>% 
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type = "Overall")

all_results_df_hcc_japan <- ind_vars %>% 
  unnest(clogit_out1) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out2) %>%
  filter(grepl("bmi", term))%>%
  bind_rows(intercept_data) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type= "Japanese American")

all_results_df_hcc_latino <- ind_vars %>% 
  unnest(clogit_out2) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out1) %>%
  filter(grepl("bmi", term))%>%
  bind_rows(intercept_data) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type = "Latino")

all_results_df_hcc_others <- ind_vars %>% 
  unnest(clogit_out3) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out1) %>%
  filter(grepl("bmi", term))%>%
  bind_rows(intercept_data) %>%
  mutate(p_adjusted = p.adjust(p.value, method = "fdr"),
         type = "Others")


all_result_df <- all_results_df_hcc %>% 
  bind_rows(all_results_df_hcc_japan) %>%
  bind_rows(all_results_df_hcc_latino) %>%
  bind_rows(all_results_df_hcc_others) %>%
  select(exposure, term, odds_ratio, exp_ci_low, exp_ci_high,effect_est,p.value,p_adjusted, type) %>%
  arrange(exposure)

```

```{r Dose response}
data_hcc1 <- data_hcc %>% 
  mutate_at(.vars = "bmi_cat_imputed",
            .funs = ~as.numeric(.))

res_hcc <- epiomics::owas_clogit(data_hcc1 , 
                                 cc_set = "setnum", 
                                 cc_status = "status", 
                                 covars = covars,
                                 omics = "bmi_cat_imputed",
                                 conf_int = TRUE) %>%
  mutate(odds_ratio = exp(estimate),
         exp_ci_low = exp(conf_low),
         exp_ci_high = exp(conf_high),
         type = "Overall")

eth_categories <- unique(data_hcc1$eth_new)

for(race in eth_categories) {
  res_hcc_temp <- epiomics::owas_clogit(data_hcc1 %>% filter(eth_new == race) , 
                                        cc_set = "setnum", 
                                        cc_status = "status", 
                                        covars = covars,
                                        omics = "bmi_cat_imputed",
                                        conf_int = TRUE) %>%
    mutate(odds_ratio = exp(estimate),
           exp_ci_low = exp(conf_low),
           exp_ci_high = exp(conf_high),
           type = race)
  res_hcc <- res_hcc %>% bind_rows(res_hcc_temp)
}

# reorder PFAS by estimate
res_hcc1 <- res_hcc |> 
  mutate(sig = case_when((p_value < 0.05 & adjusted_pval < 0.05) ~
                           "Sig.(FDR < 0.05)",
                         p_value < 0.05 ~ "Sig.(p value < 0.05)", 
                         p_value >= 0.05 ~ "Not Sig.")) 

p_trend_df <- res_hcc1 %>% select(feature_name, p_value, adjusted_pval,type) %>%
  rename(exposure = feature_name,
         p_trend = p_value)

all_result_df1 <- all_result_df %>% 
  tidylog::left_join(p_trend_df, by = c("exposure", "type")) %>%
  mutate(exp_name = "BMI" %>%
           str_remove("_quartile")) %>%
  select(type,exp_name,term,everything(), -exposure)

## N value for each quartile split by status
# Assuming df is your dataframe

# Convert to long format and create a pair identifier
data_list <- list(data1 = data_hcc %>% mutate(type = "Overall"), 
                  data2 = data_hcc %>% filter(eth == "Japanese American") %>%
                    mutate(type = "Japanese American"), 
                  data3 = data_hcc %>% filter(eth == "Latino") %>%
                    mutate(type = "Latino"),
                  data4 = data_hcc %>% filter(eth_new == "Others") %>%
                    mutate(type = "Others"))

data_n <- data.frame()

for(data in data_list){
  type = unique(data$type)
  long_data_hcc <- data %>%
    select("bmi_cat_imputed", status) %>%
    mutate(status = ifelse(status == 0, "Control", "Case")) %>%
    pivot_longer(cols = c(contains("bmi_cat_imputed")), names_to = "variable",
                 values_to = "value") %>%
    drop_na() %>%
    mutate(pair = paste0(variable, value)) %>%
    select(-variable,-value) %>%
    group_by(pair, status) %>%
    summarise(count = n(), .groups = "drop")
  # Now, to spread back for a wide format summary
  wide_summary_hcc <- pivot_wider(long_data_hcc, names_from = status,
                                  values_from = count, values_fill = list(count = 0))%>%
    mutate(type = type) %>%
    rename(term = pair)
  data_n <- data_n %>% bind_rows(wide_summary_hcc)
}

all_result_df2 <- all_result_df1 %>% 
  tidylog::left_join(data_n, by = c("term","type")) %>%
  select(type:term, Control, Case, everything())%>%
  arrange(exp_name, type, term)
```

### Fig 1b. Dose response plot for PFHpA and HCC
```{r dose response plot}
# Plot interaction effects
all_result_df3 <- all_result_df2 %>% 
  # mutate(plot_name = substr(term, 
  #                           start = nchar(term) - 8,
  #                           stop = nchar(term)) %>% toTitleCase(),
  #        pfas_name = str_remove(pfas_name, "pfas_t_")) %>%
  mutate(plot_name = case_when(grepl("<25", term) ~ "BMI < 25",
                               grepl("25-30", term) ~ "BMI 25-30",
                               grepl("≥30", term) ~ "BMI ≥30"))%>%
  mutate(plot_name = factor(plot_name, levels = c("BMI ≥30","BMI 25-30","BMI < 25")))%>%
  mutate(
    sig = case_when(p_adjusted < 0.05 ~ "FDR. Sig.", 
                    p.value < 0.05 ~ "Sig.",
                    TRUE ~ "Not Sig."),
    label = str_c("p trend = ", format(signif(p_trend, digits = 2), scientific = FALSE))) %>%
  mutate(type = case_when(type == "Overall" ~ "Overall \n(n = 464)",
                          type == "Japanese American" ~ "Japanese American \n(n = 176)", 
                          type == "Latino"~ "Latino \n(n = 158)",
                          type == "Others" ~ "Others \n(n = 134)"))

writexl::write_xlsx(all_result_df3, fs::path(dir_result, "single_categorical_bmi_analysis_1217.xlsx"))

(alcohol_hcc <-  ggplot(all_result_df3 %>% 
                        # filter(pfas_name == "PFHPA") %>%
                        filter(type != "Others \n(n = 134)"),
                      aes(x = odds_ratio, y = fct_rev(plot_name))) +
    geom_errorbar(aes(xmin = exp_ci_low, xmax = exp_ci_high),
                  width = 0, position = position_dodge(width = 1)) +
    geom_point(size = 1, position = position_dodge(width = 1)) + 
    geom_vline(xintercept = 1, linetype = 2, color = "grey") +
    # facet_grid( .~ factor(type, levels = c("Overall \n(n = 474)", "Japanese American \n(n = 182)", "Latino \n(n = 158)", "Others \n(n = 132)")))+
    facet_grid(~ factor(type, levels = c("Overall \n(n = 464)", "Japanese American \n(n = 176)", "Latino \n(n = 158)")), scales = "free")+
    xlab("Odds Ratio (95% CI)") +
    theme(#axis.title.x = element_blank(),
      # axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), 
      legend.position = "bottom",
      panel.background = element_rect(fill="white"), 
      strip.background = element_rect(fill = "white"),
      strip.text.y = element_text(angle = 0, hjust = 0),
      # strip.text.x = element_blank(),
      text = element_text(size = 15),
      axis.title.y = element_blank(),
      axis.line.x = element_line(color = "black"),
      axis.line.y = element_line(color = "black")) )
# +
# ylab("OR (95% CI) per\ndoubling of PFAS") + 
# ggtitle("Interaction of PFAS and PRS") +
# scale_color_manual(values = c("purple", "black","red")))


p1 <- alcohol_hcc + geom_text(
  data = all_result_df3%>%
    # filter(pfas_name == "PFHPA") %>%
    filter(type != "Others \n(n = 134)"),
  mapping = aes(x = -Inf, y = -Inf, label = label),
  color = "black",
  hjust   = -0.3,
  vjust   = -13,
  size = 4
)


(fig <- cowplot::plot_grid(NULL, NULL,bmi_hcc_main, 
                           p1,
                           ncol = 2,
                           rel_heights = c(0.02,0.2),
                           rel_widths = c(1,2),
                           # labels = c("A. Alcohol and HCC","B. Dose-response of Alcohol and HCC","",""),
                          label_x = c(0.13,0.04, 0, 0),
                          label_size = 12,
                          align = "v",
                          axis = "lt"))

ggsave(filename = fs::path(dir_figures,
                           "Fig1. BMI Main Effect Final_1217.jpeg"),
       width = 10,
       height = 5,
       dpi = 300,
       bg = "white")
```


