---
title: 'MEC analysis'
subtitle: "Association of interactions between SNPs and PFAS with HCC"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup snp and pfas, include=FALSE, echo=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message = FALSE, cache=FALSE,collapse=FALSE)

source(fs::path(here::here("1_Project_Setup/!libraries.R")))
source(fs::path(here::here("1_Project_Setup/!directories.R")))
source(fs::path(here::here("1_Project_Setup/!functions.R")))
source(fs::path(here::here("1_Project_Setup/!load_data.R")))
```

# Checking SNPs
**x3_12288284, x14_94378610, x17_66214462, x19_19268740 were excluded from analysis**
```{r correlation of SNPs}
pfc_chem_summaries <-  data_hcc %>%
  dplyr::select(x1_220796686:x22_43928850) %>%
  pivot_longer(cols = x1_220796686:x22_43928850) %>% 
  group_by(name) %>% 
  summarise(pct_one_val   = mean(value == 0 | value == 2, na.rm = TRUE),
            variance = var(value, na.rm = TRUE),
            mean    = mean(value, na.rm = TRUE))

snp_name_df <-  tidylog::left_join(pfc_chem_summaries, 
                                   snp_info, by = c("name" = "snp_name")) |>
  tidylog::filter(pct_one_val < 0.75)

snp_name_new <- snp_name_df$name
```

## RAF calculation
```{r}
crc_b <-data_hcc %>% filter(eth=='African American') %>% pull(sid)
crc_h <-data_hcc %>% filter(eth=='Native Hawaiian') %>% pull(sid)
crc_j <-data_hcc %>% filter(eth=='Japanese American') %>% pull(sid)
crc_l <-data_hcc %>% filter(eth=='Latino') %>% pull(sid)
crc_w <-data_hcc %>% filter(eth=='White') %>% pull(sid)

ds_correct <- data_hcc %>% select(sid, x1_220796686:x22_43928850) %>%
  column_to_rownames("sid") %>% t() %>% as.data.frame()

#get single dataset of the risk allele frequency across all race/eth
raf <- data.frame(ds_correct,         # can carry over any cols from ds_correct as needed
                  raf = rowSums(ds_correct, na.rm = TRUE)/(2*dim(data_hcc)[1]),
                  raf_b=rowSums(ds_correct %>% select(any_of(crc_b)), na.rm = TRUE)/(2*length(crc_b)),
                  raf_h=rowSums(ds_correct %>% select(any_of(crc_h)), na.rm = TRUE)/(2*length(crc_h)),
                  raf_j=rowSums(ds_correct %>% select(any_of(crc_j)), na.rm = TRUE)/(2*length(crc_j)),
                  raf_l=rowSums(ds_correct %>% select(any_of(crc_l)), na.rm = TRUE)/(2*length(crc_l)),
                  raf_w=rowSums(ds_correct %>% select(any_of(crc_w)), na.rm = TRUE)/(2*length(crc_w)))%>%
  rownames_to_column("snp_name") %>%
  tidylog::left_join(snp_info %>% select(rsid, snp_name), by = "snp_name")

write_csv(raf, fs::path(dir_result,"Supplemental Tables", "raf_hcc.csv"))

```

# Distribution of SNP
```{r}
temp_data <- data_hcc %>% rename(rs5117 = x19_44915533)%>%
  mutate(rs5117 = round(rs5117, 0)) %>%
  drop_na(rs5117)%>%
  mutate(rs5117 = case_when(rs5117 == 2 ~ "2 rs5117-T alleles", 
                            rs5117 == 1 ~ "1 rs5117-T allele", 
                            rs5117 == 0 ~ "0 rs5117-T alleles"))
 
temp_counts <- temp_data %>%
  count(rs5117) 
temp_counts <- temp_counts %>%
  mutate(perc = n / sum(n) * 100)%>%
  bind_rows(data.frame(rs5117 = "Overall",
                       n = sum(temp_counts$n),
                       perc = 100)) %>%
  mutate(rs5117 = factor(rs5117, levels = c("2 rs5117-T alleles",
                                            "1 rs5117-T allele",
                                            "0 rs5117-T alleles",
                                            "Overall"
                                            ))) %>%
  mutate(group = if_else(str_detect(rs5117, "Overall"), "overall", "genetic"),
         group = factor(group, levels = c("overall", "genetic")))

# Plot
(p_dist<- ggplot(temp_counts, aes(y = rs5117, x = n)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(perc, 0), "%")), 
            hjust = -0.1) +  
  labs(x = "Count") +
  facet_grid(group ~. , scales = "free_y", space = "free_y")+
  theme(panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_blank(),
          # strip.text.x = element_blank(),
          text = element_text(size = 12),
          # axis.text.x =element_text(angle = 60, hjust = 1),
          axis.line.x = element_line(color = "black"),
          axis.title.y=element_blank(),
          axis.line.y = element_line(color = "black"),
          legend.position = "bottom",
          legend.title = element_blank())+
  xlim(0, max(temp_counts$n) * 1.2) ) 
```

# Analysis 1: PRS risk score

**Outcome: HCC vs control**

**Exposure: normalzied log2 PFAS**

**Covars: smoking, diabetes, q1_edih, v1:v6**

**Analysis: conditional logistic regression**

**Data Info: 223 cases, 223 controls**

**Model2:status ~ pfas + prs_wt_std + pfas:prs_wt_std + diabetes + q1_edih + v1...+v6**

## Interaction of PRS (dichotomous weighted PRS) and PFAS
```{r}
model_interaction <- function(pfas){
  
  data_analysis$status_surv <- survival::Surv(rep(1, nrow(data_analysis)), data_analysis$status)
  
  # Formula
  base_covars <- paste(paste(covars[-4], collapse = " + "), 
                       "+ strata(setnum)")
  prs = "prs_wt_median"
  
  frm_with_int <- as.formula(paste("status_surv ~", paste0(pfas, "*", prs), base_covars, sep = " + "))
  # Model
  model_with_int <- coxph(frm_with_int, 
                           data = data_analysis, 
                           method = "exact")
  
  int <- model_with_int %>% tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")

  # Interaction effect at high/low PRS
  res <- emtrends(model_with_int,
                  ~ prs_wt_median,
                  var = pfas,
                  at = list(prs_wt_median = c("1", "2"),
                            smokestatus_imputed = "Never",
                            diabetes = "No",
                            q1_edih = 0,
                            prs_wt = 0,
                            v1 = 0,
                            v2 = 0,
                            v3 = 0,
                            v4 = 0,
                            v5 = 0,
                            v6 = 0)) %>%
    summary(infer = TRUE) %>%
    as.data.frame() %>%
    select(`prs_wt_median`, ends_with(".trend"), asymp.LCL, asymp.UCL, p.value) %>%
    rename(estimate = ends_with(".trend"),
           conf_low = asymp.LCL,
           conf_high = asymp.UCL,
           p_value = p.value) %>%
    mutate(type = case_when(`prs_wt_median` == 1 ~ "Low Genetic Risk",
                            `prs_wt_median` == 2 ~ "High Genetic Risk"
                            ))
  
  # Main effect without interaction
  res_main <- epiomics::owas_clogit(data_analysis%>% select(-status_surv) , 
                                    cc_set = "setnum", 
                                    cc_status = "status", 
                                    covars = c(covars[-4], "prs_wt_median"),
                                    omics = pfas,
                                    conf_int = TRUE,
                                    method = "exact") %>%
    mutate(type = "Overall (no interaction with PRS)")%>%
    select(-adjusted_pval)
  # Combine results
  res_combined <- bind_rows(res_main, res) %>%
    mutate(odds_ratio = exp(estimate),
           exp_ci_low = exp(conf_low),
           exp_ci_high = exp(conf_high),
           group = if_else(str_detect(type, "Overall"), "overall", "with genetic"),
           pfas = pfas)
  
  res_combined
  return(list(res_combined = res_combined,
  int = int))
}

res <- map(c(pfas_name[-3], "pfas_t_PFBS_detected"), model_interaction)%>%
  map("res_combined") %>%
    bind_rows()%>%
    mutate(pfas = str_remove(pfas,"pfas_t_"))%>%
  group_by(type) %>%
  mutate(q_value = p.adjust(p_value, "fdr"))%>%
  select(-feature_name)

int <- map(c(pfas_name[-3], "pfas_t_PFBS_detected"), model_interaction)%>%
  map("int") %>%
  bind_rows()%>%
  filter(str_detect(term, ":")) %>%
  mutate(pfas = str_remove(term,"pfas_t_") %>%
           str_remove(":prs_wt_median2"))%>%
  rename(p_value = p.value) %>%
  mutate(q_value = p.adjust(p_value, "fdr"))%>%
  mutate(type = "PFAS*PRS interaction",
         odds_ratio = exp(estimate),
         exp_ci_low = exp(conf.low),
         exp_ci_high = exp(conf.high))%>%
  select(pfas, p_value, q_value, type, odds_ratio, exp_ci_low, exp_ci_high)

res <- int %>% bind_rows(res)



writexl::write_xlsx(res, fs::path(dir_result, "pfas_prs_wt_median_interaction_result_081126.xlsx"))
```

## Interaction of PRS (dichotomous weighted PRS) and PFAS (dichotomous weighted PRS)
```{r}

pfas_name_median <- c(colnames(data_analysis %>% select(contains("median"), -contains("prs"))), "pfas_t_PFBS_detected")

run_interactionR <- function(pfas, data = data_analysis) {

  frm <- as.formula(paste(
    "status ~", pfas, "* prs_wt_median +",
    "smokestatus_imputed + diabetes + q1_edih +",
    "v1 + v2 + v3 + v4 + v5 + v6 + strata(setnum)"
  ))

  mod <- clogit(frm, data = data, method = "exact")

  out <- interactionR::interactionR(mod,
                                     exposure_names = c("prs_wt_median",pfas),
                                     ci.type = "mover",
                                     ci.level = 0.95,
                                     em = TRUE,
                                     recode = FALSE)

  out$dframe %>%
    mutate(pfas = pfas)
}


int_df <- map(pfas_name_median[-3], run_interactionR) %>%
  bind_rows()%>%
  mutate(pfas = str_remove(pfas, "pfas_t_") %>% str_remove("_median"))%>%
  select(pfas, everything())%>%
  janitor::clean_names()

int_df1 <- int_df %>%
  filter(!str_detect(measures, "OR0|OR1"))%>%
  mutate(type = case_when(str_detect(measures, "prs_wt_median2==0") ~ "Low Genetic Risk",
                           str_detect(measures, "prs_wt_median2==1") ~ "High Genetic Risk",
                           TRUE ~ measures))%>%
  select(-measures)%>%
  rename(odds_ratio = estimates)%>%
  group_by(type) %>%
  rename(p_value = p) %>%
  mutate(
    q_value = p.adjust(p_value, "fdr"))

writexl::write_xlsx(int_df1, fs::path(dir_result, "pfas_median_prs_wt_median_interaction_result_081126.xlsx"))
```

## PFHpA: Plot PFAS and PRS interaction results
```{r pfhpa prs interaction plot, fig.width=7, fig.height=4}
prs_df <- int_df1 %>%
      ungroup() %>%
      filter(pfas %in% c("PFHPA","PFNA")) %>%
      # mutate(type = ifelse(type == "Multiplicative scale", "Multiplicative Scale", type)) %>%
      rename(exp_ci_low = ci_ll, exp_ci_high = ci_ul) %>%
      select(pfas, type, odds_ratio, exp_ci_low, exp_ci_high) %>%
  mutate(type = factor(type, levels = c("Overall (no interaction with PRS)",
                                        "Low Genetic Risk",
                                        "High Genetic Risk",
                                        "Multiplicative scale",
                                        "RERI"))) 

(prs_plot <- ggplot(prs_df %>% 
                            filter(!type %in% c("Multiplicative scale",
                                        "RERI", "Overall (no interaction with PRS)")), 
                          aes(x = odds_ratio, y = fct_rev(type))) +
    geom_errorbar(aes(xmin = exp_ci_low, xmax = exp_ci_high), width = 0) +
    geom_point(size = 2) +
    geom_vline(xintercept = 1, linetype = 2, color = "grey") +
    xlab("Odds Ratio (95% CI)") +
    ggtitle("PFAS-HCC Risk by PRS") +
    facet_grid(~pfas, space = "free") +
    theme(panel.background = element_blank(),
          # strip.text = element_blank(),
          strip.background = element_blank(),
          plot.title = element_text(face = "plain"),
          text = element_text(size = 12),
          axis.title.y = element_blank(),
          axis.line.x = element_line(color = "black"),
          axis.line.y = element_line(color = "black")))

ggsave(filename = fs::path(dir_figures,
                           "Fig2.PFAS and PRS interaction_0812.jpeg"),
       width = 6,
       height = 3,
       dpi = 300,
       bg = "white")
```


# Analysis 2: All SNPS
**Outcome: HCC vs control**

**Exposure: normalzied log2 PFAS, raw SNP**

**Covars: smoking, diabetes, q1_edih, prs_wt_std, v1:v6**

**Analysis: conditional logistic regression**

**Model: outcome ~ one pfas + one SNP + pfas:snp**


```{r interaction between PFAS and SNPs hcc}

# Covariates for main exposure outcome associations
covariate_input <- paste(covars[1:4], collapse = " + ")

# create dataframe with names of dependent vars (PFAS) 
pfas_names <- tibble(exposure = c(pfas_name, "pfas_t_PFBS_detected"))

## create dataframe with names of SNP vars ----
# snp_name <- tibble(snp = snp_name_scld)
# based on Dave's suggestion use raw snp instead of normalized snp
snp_name <- tibble(snp = snp_name_new)

pfas_snp <- expand_grid(pfas_names, snp_name) %>%
  mutate(pfas_snp = paste(exposure,"*",snp)) 

# Get Model covariates 
covar_names <- tibble(covariates =  covariate_input)

# 2. Make all clogit formulas ----
ind_vars <- expand_grid(pfas_snp,covar_names) |>  
  mutate(formula = paste("status ~ v1 + v2 + v3 + v4 + v5 + v6",
                         "+", pfas_snp,
                         "+", covariates, 
                         "+strata(setnum)") |>
           str_replace("\\+  \\+", " + "))

## Analysis for all related participants
ind_vars$clogit_out <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_analysis,
          method = "exact"
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Japanese American 
ind_vars$clogit_out1 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_analysis %>% filter(eth == "Japanese American"),
          method = "exact"
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Latino
ind_vars$clogit_out2 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_analysis %>% filter(eth == "Latino"),
          method = "exact"
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit")) 

## Analysis for Others
ind_vars$clogit_out3 <- map(
  ind_vars$formula, 
  ~clogit(as.formula(.),
          data = data_analysis %>% filter(eth_new == "Others"),
          method = "exact"
  ) %>%
    tidy(., conf.int = TRUE, level = 0.95) %>% 
    mutate(
      odds_ratio = exp(estimate),
      exp_ci_low = exp(conf.low),
      exp_ci_high = exp(conf.high),
      model_type = "clogit"))
# Get final data -------------------

# intercept_data<- tibble(exposure = pfas_name_cat, 
#                         term = paste0(pfas_name_cat, "1"), 
#                         odds_ratio = rep(1,length(pfas_name_cat)),
#                         effect_est  = rep("1",length(pfas_name_cat)))

all_results_df_hcc <- ind_vars %>% 
  unnest(clogit_out) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out1, -clogit_out2,-clogit_out3)

all_results_df_hcc_japan <- ind_vars %>% 
  unnest(clogit_out1) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out2, -clogit_out3)

all_results_df_hcc_latino <- ind_vars %>% 
  unnest(clogit_out2) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out1, -clogit_out3)

all_results_df_hcc_others <- ind_vars %>% 
  unnest(clogit_out3) %>% 
  select(model_type, covariates, everything()) %>%
  mutate(sig = case_when(p.value < 0.05 ~ "Sig", 
                         p.value >= 0.05 ~ "Not Sig"),
         effect_est = jag2::effest_ci(odds_ratio, exp_ci_low, exp_ci_high)) %>%
  select(-clogit_out, -clogit_out1, clogit_out2)


### All races
res_fun <- function(data){
  data <- data %>%
    filter(grepl("pfas",term)|grepl("x",term))%>%
    mutate(pfas_name = str_remove(exposure, "pfas_t_")) %>%
    mutate(effect_type = case_when(str_detect(term, ":")   ~ "PFAS and SNP",
                              str_detect(term, "^x")   ~ "Genetic Risk",
                              TRUE ~ "PFAS")) %>%
    group_by(effect_type) %>%
    mutate(
           p_adjusted = p.adjust(p.value, method = "fdr"),
           sig = case_when(p_adjusted < 0.05 ~ "FDR. Sig.", 
                           p.value < 0.05 ~ "Sig.",
                           TRUE ~ "Not Sig.")) %>%
    # ungroup() %>%
    dplyr::select(exposure, pfas_name, snp, term, effect_type, p.value, p_adjusted, sig, odds_ratio, contains("exp_")) %>%
    tidylog::left_join(snp_info%>%
                         dplyr::select(rsid, snp_name, genes), by = c("snp"="snp_name"))
  
    return(data)
}


overall <- res_fun(all_results_df_hcc)
japanese <- res_fun(all_results_df_hcc_japan)
latino <- res_fun(all_results_df_hcc_latino)
others <- res_fun(all_results_df_hcc_others)


all_result <- overall %>%
  mutate(type = "All races") %>%
  bind_rows(japanese %>%
              mutate (type = "Japanese American")) %>%
  bind_rows(latino %>%
              mutate (type = "Latino"))%>% 
  bind_rows(others %>%
              mutate (type = "Others"))%>%
  mutate(name_plot = case_when(effect_type == "PFAS" ~ paste0(genes, "(", " PFAS Main Effect)"),
                                effect_type == "Genetic Risk" ~ paste0(genes, " Main Effect"),
                                effect_type == "PFAS and SNP" ~ paste0(genes, "*",
                                                                       "PFAS Interaction")),
         gene = if_else(is.na(genes), snp, genes),
         rsid = if_else(is.na(rsid), snp, rsid))

writexl::write_xlsx(all_result, fs::path(dir_result, "PFAS and SNP interaction_081026.xlsx"))
```
