---
title: "Gene Ontology enrichment analysis"
subtitle: "Genes associated with PFAS–HCC candidate variants"
author: "MEC PFAS Liver Cancer Study"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: false
    code_folding: hide
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(clusterProfiler)
library(org.Hs.eg.db)
library(dplyr)
library(stringr)
library(flextable)

metadata_file <- fs::path_abs(fs::path(
  here::here(),
  "..", "..", "..", "..",
  "1_analysis_ready_data",
  "metadata",
  "snp_info_new_082626.csv"
))

if (!file.exists(metadata_file)) {
  stop("SNP metadata file not found: ", metadata_file)
}
```

## Methods

Gene symbols associated with the candidate variants were deduplicated and tested
for over-representation in Gene Ontology (GO) biological process, molecular
function, and cellular component terms. P values were adjusted using the
Benjamini--Hochberg method. The table is restricted to GO terms containing at
least three genes from the input list.

```{r enrich-go}
snp_metadata <- read.csv(metadata_file, check.names = FALSE)

cleaned_genes <- snp_metadata$Genes |>
  str_split(";\\s*") |>
  unlist(use.names = FALSE) |>
  trimws() |>
  (\(x) x[!is.na(x) & x != "" & tolower(x) != "none"])() |>
  unique()

# The only enrichment analysis run in this report.
ora_go_results <- enrichGO(
  gene          = cleaned_genes,
  OrgDb         = org.Hs.eg.db,
  keyType       = "SYMBOL",
  ont           = "ALL",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  readable      = TRUE
)

res_more_than_two_gene <- as.data.frame(ora_go_results) |>
  filter(Count > 2)
```

## Results

The analysis included `r length(cleaned_genes)` unique candidate genes. The
table reports `r nrow(res_more_than_two_gene)` enriched GO terms represented by
at least three candidate genes.

```{r publication-table}
if (nrow(res_more_than_two_gene) == 0L) {
  cat("No GO terms met the prespecified significance and gene-count criteria.")
} else {
  format_p <- function(x) {
    ifelse(x < 0.001, "<0.001", sprintf("%.3f", x))
  }

  go_table_data <- res_more_than_two_gene |>
    arrange(ONTOLOGY, p.adjust) |>
    transmute(
      Ontology = recode(
        ONTOLOGY,
        BP = "Biological process",
        MF = "Molecular function",
        CC = "Cellular component"
      ),
      `GO ID` = ID,
      `GO term` = Description,
      `Gene ratio` = GeneRatio,
      `Candidate genes` = str_replace_all(geneID, "/", ", "),
      `No. genes` = Count,
      `P value` = format_p(pvalue),
      `FDR-adjusted P value` = format_p(p.adjust)
    )

  go_table <- flextable(go_table_data) |>
    set_caption(
      caption = "Table 1. Gene Ontology terms enriched among PFAS–HCC candidate genes"
    ) |>
    theme_booktabs() |>
    bold(part = "header") |>
    bg(part = "header", bg = "#D9EAF2") |>
    color(part = "header", color = "#17365D") |>
    align(j = c("Ontology", "GO ID", "Gene ratio", "No. genes",
                "P value", "FDR-adjusted P value"), align = "center",
          part = "all") |>
    valign(valign = "top", part = "body") |>
    font(fontname = "Arial", part = "all") |>
    fontsize(size = 9, part = "body") |>
    fontsize(size = 9, part = "header") |>
    padding(padding = 4, part = "all") |>
    autofit() |>
    add_footer_lines(
      values = paste0(
        "Abbreviations: FDR, false discovery rate; GO, Gene Ontology. ",
        "Gene ratio is the number of candidate genes annotated to a term ",
        "divided by the number of mapped candidate genes. FDR adjustment ",
        "used the Benjamini–Hochberg method."
      )
    ) |>
    fontsize(size = 8, part = "footer") |>
    italic(part = "footer") |>
    set_table_properties(layout = "autofit", width = 1)

  go_table
}
```

## Number of enriched pathways per gene

```{r gene-pathway-counts, include=FALSE}
gene_pathway_counts <- res_more_than_two_gene |>
  select(ID, geneID) |>
  mutate(Gene = str_split(geneID, "/")) |>
  tidyr::unnest_longer(Gene) |>
  distinct(ID, Gene) |>
  count(Gene, name = "Number of pathways") |>
  arrange(desc(`Number of pathways`), Gene)
```

The following table shows how many of the retained enriched GO terms include
each candidate gene, ordered from the largest to the smallest count.

```{r gene-pathway-table}
if (nrow(gene_pathway_counts) == 0L) {
  cat("No genes were represented in the retained GO terms.")
} else {
  gene_pathway_table <- flextable(gene_pathway_counts) |>
    set_caption(
      caption = "Table 2. Number of enriched Gene Ontology terms containing each candidate gene"
    ) |>
    theme_booktabs() |>
    bold(part = "header") |>
    bg(part = "header", bg = "#D9EAF2") |>
    color(part = "header", color = "#17365D") |>
    align(j = "Gene", align = "left", part = "all") |>
    align(j = "Number of pathways", align = "center", part = "all") |>
    valign(valign = "top", part = "body") |>
    font(fontname = "Arial", part = "all") |>
    fontsize(size = 9, part = "all") |>
    padding(padding = 4, part = "all") |>
    autofit() |>
    add_footer_lines(
      values = paste0(
        "Counts are based on enriched GO terms with FDR-adjusted P < 0.05 ",
        "and at least three candidate genes. Each gene is counted no more ",
        "than once per GO term."
      )
    ) |>
    fontsize(size = 8, part = "footer") |>
    italic(part = "footer") |>
    set_table_properties(layout = "autofit", width = 0.6)

  gene_pathway_table
}
```
